### 2021-08-27 paper/paper.tex
Add paper base

%%
%% This is file `sample-sigplan.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigplan')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigplan.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigplan,screen]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{10.1145/1122445.1122456}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
  June 03--05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Verification of OpenSSL's SHA-3 Hash Computations}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Benjamin Winters}
\authornotemark[1]
\affiliation{%
  \institution{Brigham Young University}
  \city{Provo}
  \state{Utah}
  \country{USA}
  \postcode{84601}
}

\author{Parker Hanson}
\authornotemark[2]
\affiliation{%
  \institution{Brigham Young University}
  \city{Provo}
  \state{Utah}
  \country{USA}
  \postcode{84601}
}

\author{Eric Mercer}
\authornotemark[3]
\email{egm@cs.byu.edu}
\affiliation{%
  \institution{Brigham Young University}
  \city{Provo}
  \state{Utah}
  \country{USA}
  \postcode{84601}
}

\author{Brett Decker}
\authornotemark[4]
\email{bdecker@cs.byu.edu}
\affiliation{%
  \institution{Brigham Young University}
  \city{Provo}
  \state{Utah}
  \country{USA}
  \postcode{84601}
}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
%%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  A clear and well-documented \LaTeX\ document is presented as an
  article formatted for publication by ACM in a conference proceedings
  or journal publication. Based on the ``acmart'' document class, this
  article presents and explains many of the common variations, as well
  as many of the formatting elements an author may use in the
  preparation of the documentation of their work.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
  <ccs2012>
     <concept>
         <concept_id>10002978.10002986.10002990</concept_id>
         <concept_desc>Security and privacy~Logic and verification</concept_desc>
         <concept_significance>500</concept_significance>
         </concept>
   </ccs2012>
  \end{CCSXML}

  \ccsdesc[500]{Security and privacy~Logic and verification}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Verification, SHA-3, OpenSSL, Cryptol, SAW, Symbolic Execution}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
ACM's consolidated article template, introduced in 2017, provides a
consistent \LaTeX\ style for use across ACM publications, and
incorporates accessibility and metadata-extraction functionality
necessary for future Digital Library endeavors. Numerous ACM and
SIG-specific \LaTeX\ templates have been examined, and their unique
features incorporated into this single new template.

If you are new to publishing with ACM, this document is a valuable
guide to the process of preparing your work for publication. If you
have published with ACM before, this document provides insight and
instruction into more recent changes to the article template.

The ``\verb|acmart|'' document class can be used to prepare articles
for any ACM publication --- conference or journal, and for any stage
of publication, from review to final ``camera-ready'' copy, to the
author's own version, with {\itshape very} few changes to the source.

\section{Template Overview}
As noted in the introduction, the ``\verb|acmart|'' document class can
be used to prepare many different kinds of documentation --- a
double-blind initial submission of a full-length technical paper, a
two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
journal article, a SIGCHI Extended Abstract, and more --- all by
selecting the appropriate {\itshape template style} and {\itshape
  template parameters}.

This document will explain the major features of the document
class. For further information, the {\itshape \LaTeX\ User's Guide} is
available from
\url{https://www.acm.org/publications/proceedings-template}.

\subsection{Template Styles}

The primary parameter given to the ``\verb|acmart|'' document class is
the {\itshape template style} which corresponds to the kind of publication
or SIG publishing the work. This parameter is enclosed in square
brackets and is a part of the {\verb|documentclass|} command:
\begin{verbatim}
  \documentclass[STYLE]{acmart}
\end{verbatim}

Journals use one of three template styles. All but three ACM journals
use the {\verb|acmsmall|} template style:
\begin{itemize}
\item {\verb|acmsmall|}: The default journal template style.
\item {\verb|acmlarge|}: Used by JOCCH and TAP.
\item {\verb|acmtog|}: Used by TOG.
\end{itemize}

The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
\begin{itemize}
\item {\verb|acmconf|}: The default proceedings template style.
\item{\verb|sigchi|}: Used for SIGCHI conference articles.
\item{\verb|sigchi-a|}: Used for SIGCHI ``Extended Abstract'' articles.
\item{\verb|sigplan|}: Used for SIGPLAN conference articles.
\end{itemize}

\subsection{Template Parameters}

In addition to specifying the {\itshape template style} to be used in
formatting your work, there are a number of {\itshape template parameters}
which modify some part of the applied template style. A complete list
of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}

Frequently-used parameters, or combinations of parameters, include:
\begin{itemize}
\item {\verb|anonymous,review|}: Suitable for a ``double-blind''
  conference submission. Anonymizes the work and includes line
  numbers. Use with the \verb|\acmSubmissionID| command to print the
  submission's unique ID on each page of the work.
\item{\verb|authorversion|}: Produces a version of the work suitable
  for posting by the author.
\item{\verb|screen|}: Produces colored hyperlinks.
\end{itemize}

This document uses the following string as the first command in the
source file:
\begin{verbatim}
\documentclass[sigplan,screen]{acmart}
\end{verbatim}

\section{Modifications}

Modifying the template --- including but not limited to: adjusting
margins, typeface sizes, line spacing, paragraph and list definitions,
and the use of the \verb|\vspace| command to manually adjust the
vertical spacing between elements of your work --- is not allowed.

{\bfseries Your document will be returned to you for revision if
  modifications are discovered.}

\section{Typefaces}

The ``\verb|acmart|'' document class requires the use of the
``Libertine'' typeface family. Your \TeX\ installation should include
this set of packages. Please do not substitute other typefaces. The
``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
as they will override the built-in typeface families.

\section{Title Information}

The title of your work should use capital letters appropriately -
\url{https://capitalizemytitle.com/} has useful rules for
capitalization. Use the {\verb|title|} command to define the title of
your work. If your work has a subtitle, define it with the
{\verb|subtitle|} command.  Do not insert line breaks in your title.

If your title is lengthy, you must define a short version to be used
in the page headers, to prevent overlapping text. The \verb|title|
command has a ``short title'' parameter:
\begin{verbatim}
  \title[short title]{full title}
\end{verbatim}

\section{Authors and Affiliations}

Each author must be defined separately for accurate metadata
identification. Multiple authors may share one affiliation. Authors'
names should not be abbreviated; use full first names wherever
possible. Include authors' e-mail addresses whenever possible.

Grouping authors' names or e-mail addresses, or providing an ``e-mail
alias,'' as shown below, is not acceptable:
\begin{verbatim}
  \author{Brooke Aster, David Mehldau}
  \email{dave,judy,steve@university.edu}
  \email{firstname.lastname@phillips.org}
\end{verbatim}

The \verb|authornote| and \verb|authornotemark| commands allow a note
to apply to multiple authors --- for example, if the first two authors
of an article contributed equally to the work.

If your author list is lengthy, you must define a shortened version of
the list of authors to be used in the page headers, to prevent
overlapping text. The following command should be placed just after
the last \verb|\author{}| definition:
\begin{verbatim}
  \renewcommand{\shortauthors}{McCartney, et al.}
\end{verbatim}
Omitting this command will force the use of a concatenated list of all
of the authors' names, which may result in overlapping text in the
page headers.

The article template's documentation, available at
\url{https://www.acm.org/publications/proceedings-template}, has a
complete explanation of these commands and tips for their effective
use.

Note that authors' addresses are mandatory for journal articles.

\section{Rights Information}

Authors of any work published by ACM will need to complete a rights
form. Depending on the kind of work, and the rights management choice
made by the author, this may be copyright transfer, permission,
license, or an OA (open access) agreement.

Regardless of the rights management choice, the author will receive a
copy of the completed rights form once it has been submitted. This
form contains \LaTeX\ commands that must be copied into the source
document. When the document source is compiled, these commands and
their parameters add formatted text to several areas of the final
document:
\begin{itemize}
\item the ``ACM Reference Format'' text on the first page.
\item the ``rights management'' text on the first page.
\item the conference information in the page header(s).
\end{itemize}

Rights information is unique to the work; if you are preparing several
works for an event, make sure to use the correct set of commands with
each of the works.

The ACM Reference Format text is required for all articles over one
page in length, and is optional for one-page articles (abstracts).

\section{CCS Concepts and User-Defined Keywords}

Two elements of the ``acmart'' document class provide powerful
taxonomic tools for you to help readers find your work in an online
search.

The ACM Computing Classification System ---
\url{https://www.acm.org/publications/class-2012} --- is a set of
classifiers and concepts that describe the computing
discipline. Authors can select entries from this classification
system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
commands to be included in the \LaTeX\ source.

User-defined keywords are a comma-separated list of words and phrases
of the authors' choosing, providing a more flexible way of describing
the research being presented.

CCS concepts and user-defined keywords are required for for all
articles over two pages in length, and are optional for one- and
two-page articles (or abstracts).

\section{Sectioning Commands}

Your work should use standard \LaTeX\ sectioning commands:
\verb|section|, \verb|subsection|, \verb|subsubsection|, and
\verb|paragraph|. They should be numbered; do not remove the numbering
from the commands.

Simulating a sectioning command by setting the first word or words of
a paragraph in boldface or italicized text is {\bfseries not allowed.}

\section{Tables}

The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
high-quality tables.

Table captions are placed {\itshape above} the table.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and the
table caption.  The contents of the table itself must go in the
\textbf{tabular} environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.  Again,
detailed instructions on \textbf{tabular} material are found in the
\textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed output of
this document.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \textbf{table*} to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}

Always use midrule to separate table header rows from data rows, and
use it only for this purpose. This enables assistive technologies to
recognise table headers and support their users in navigating tables
more easily.

\section{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsection{Inline (In-text) Equations}
A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \textbf{math} environment,
which can be invoked with the usual
\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \textbf{equation}
environment. An unnumbered display equation is produced by the
\textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols and
structures available in \LaTeX\@; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\section{Figures}

The ``\verb|figure|'' environment should be used for figures. One or
more images can be placed within a figure. If your figure contains
third-party material, you must clearly identify it as such, as shown
in the example below.
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{sample-franklin}
  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
    Ewing, Inc. [Public domain], via Wikimedia
    Commons. (\url{https://goo.gl/VLCRBB}).}
  \Description{A woman and a girl in white dresses sit in an open car.}
\end{figure}

Your figures should contain a caption which describes the figure to
the reader.

Figure captions are placed {\itshape below} the figure.

Every figure should also have a figure description unless it is purely
decorative. These descriptions convey what’s in the image to someone
who cannot see it. They are also used by search engine crawlers for
indexing images, and when images cannot be loaded.

A figure description must be unformatted plain text less than 2000
characters long (including spaces).  {\bfseries Figure descriptions
  should not repeat the figure caption – their purpose is to capture
  important information that is not already provided in the caption or
  the main text of the paper.} For figures that convey important and
complex new information, a short text description may not be
adequate. More complex alternative descriptions can be placed in an
appendix and referenced in a short figure description. For example,
provide a data table capturing the information in a bar chart, or a
structured list representing a graph.  For additional information
regarding how best to write figure descriptions and why doing this is
so important, please see
\url{https://www.acm.org/publications/taps/describing-figures/}.

\subsection{The ``Teaser Figure''}

A ``teaser figure'' is an image, or set of images in one figure, that
are placed after all author and affiliation information, and before
the body of the article, spanning the page. If you wish to have such a
figure in your article, place the command immediately before the
\verb|\maketitle| command:
\begin{verbatim}
  \begin{teaserfigure}
    \includegraphics[width=\textwidth]{sampleteaser}
    \caption{figure caption}
    \Description{figure description}
  \end{teaserfigure}
\end{verbatim}

\section{Citations and Bibliographies}

The use of \BibTeX\ for the preparation and formatting of one's
references is strongly recommended. Authors' names should be complete
--- use full first names (``Donald E. Knuth'') not initials
(``D. E. Knuth'') --- and the salient identifying features of a
reference should be included: title, year, volume, number, pages,
article DOI, etc.

The bibliography is included in your source document with these two
commands, placed just before the \verb|\end{document}| command:
\begin{verbatim}
  \bibliographystyle{ACM-Reference-Format}
  \bibliography{bibfile}
\end{verbatim}
where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
suffix, of the \BibTeX\ file.

Citations and references are numbered by default. A small number of
ACM publications have citations and references formatted in the
``author year'' style; for these exceptions, please include this
command in the {\bfseries preamble} (before the command
``\verb|\begin{document}|'') of your \LaTeX\ source:
\begin{verbatim}
  \citestyle{acmauthoryear}
\end{verbatim}

  Some examples.  A paginated journal article \cite{Abril07}, an
  enumerated journal article \cite{Cohen07}, a reference to an entire
  issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
  monograph/whole book in a series (see 2a in spec. document)
  \cite{Harel79}, a divisible-book such as an anthology or compilation
  \cite{Editor00} followed by the same example, however we only output
  the series if the volume number is given \cite{Editor00a} (so
  Editor00a's series should NOT be present since it has no vol. no.),
  a chapter in a divisible book \cite{Spector90}, a chapter in a
  divisible book in a series \cite{Douglass98}, a multi-volume work as
  book \cite{Knuth97}, a couple of articles in a proceedings (of a
  conference, symposium, workshop for example) (paginated proceedings
  article) \cite{Andler79, Hagerup1993}, a proceedings article with
  all possible elements \cite{Smith10}, an example of an enumerated
  proceedings article \cite{VanGundy07}, an informally published work
  \cite{Harel78}, a couple of preprints \cite{Bornmann2019,
    AnzarootPBM14}, a doctoral dissertation \cite{Clarkson85}, a
  master's thesis: \cite{anisi03}, an online document / world wide web
  resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game
  (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05}
  and (Case 3) a patent \cite{JoeScientist001}, work accepted for
  publication \cite{rous08}, 'YYYYb'-test for prolific author
  \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
  contain 'duplicate' DOI and URLs (some SIAM articles)
  \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
  multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
  couple of citations with DOIs:
  \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
  citations: \cite{TUGInstmem, Thornburg01, CTANacmart}. Artifacts:
  \cite{R} and \cite{UMassCitations}.

\section{Acknowledgments}

Identification of funding sources and other support, and thanks to
individuals and groups that assisted in the research and the
preparation of the work should be included in an acknowledgment
section, which is placed just before the reference section in your
document.

This section has a special environment:
\begin{verbatim}
  \begin{acks}
  ...
  \end{acks}
\end{verbatim}
so that the information contained therein can be more easily collected
during the article metadata extraction phase, and to ensure
consistency in the spelling of the section heading.

Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.

\section{Appendices}

If your work needs an appendix, add it before the
``\verb|\end{document}|'' command at the conclusion of your source
document.

Start the appendix with the ``\verb|appendix|'' command:
\begin{verbatim}
  \appendix
\end{verbatim}
and note that in the appendix, sections are lettered, not
numbered. This document has two appendices, demonstrating the section
and subsection identification method.

\section{SIGCHI Extended Abstracts}

The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
not in Word) produces a landscape-orientation formatted article, with
a wide left margin. Three environments are available for use with the
``\verb|sigchi-a|'' template style, and produce formatted output in
the margin:
\begin{itemize}
\item {\verb|sidebar|}:  Place formatted text in the margin.
\item {\verb|marginfigure|}: Place a figure in the margin.
\item {\verb|margintable|}: Place a table in the margin.
\end{itemize}

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
To Robert, for the bagels and explaining CMYK and color spaces.
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Research Methods}

\subsection{Part One}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
lacinia dolor. Integer ultricies commodo sem nec semper.

\subsection{Part Two}

Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
eros. Vivamus non purus placerat, scelerisque diam eu, cursus
ante. Etiam aliquam tortor auctor efficitur mattis.

\section{Online Resources}

Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
enim maximus. Vestibulum gravida massa ut felis suscipit
congue. Quisque mattis elit a risus ultrices commodo venenatis eget
dui. Etiam sagittis eleifend elementum.

Nam interdum magna at lectus dignissim, ac dignissim lorem
rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
massa et mattis lacinia.

\end{document}
\endinput
%%
%% End of file `sample-sigplan.tex'.

### 2021-08-27 paper/paper.tex
Add paper template and ignare list

@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigplan,screen]{acmart}
+\documentclass[sigplan, 10pt, anonymous, review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Verification of OpenSSL's SHA-3 Hash Computations}
+\title{The Name of the Title is Hope}
 
@@
 \author{Benjamin Winters}
-\authornotemark[1]
+\author{Parker Hanson}
+\author{Eric Mercer}
+\email{egm@cs.byu.edu}
+\orcid{0000-0002-2264-2958}
+\author{Brett Decker}
+\email{bdecker@cs.byu.edu}
 \affiliation{%
   \institution{Brigham Young University}
+  \streetaddress{3365 TMCB}
   \city{Provo}
@@
   \country{USA}
-  \postcode{84601}
-}
-
-\author{Parker Hanson}
-\authornotemark[2]
-\affiliation{%
-  \institution{Brigham Young University}
-  \city{Provo}
-  \state{Utah}
-  \country{USA}
-  \postcode{84601}
-}
-
-\author{Eric Mercer}
-\authornotemark[3]
-\email{egm@cs.byu.edu}
-\affiliation{%
-  \institution{Brigham Young University}
-  \city{Provo}
-  \state{Utah}
-  \country{USA}
-  \postcode{84601}
-}
-
-\author{Brett Decker}
-\authornotemark[4]
-\email{bdecker@cs.byu.edu}
-\affiliation{%
-  \institution{Brigham Young University}
-  \city{Provo}
-  \state{Utah}
-  \country{USA}
-  \postcode{84601}
+  \postcode{84602}
 }
@@
 %% of authors' names for this purpose.
-%%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
+\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
 
@@
  <concept>
-         <concept_id>10002978.10002986.10002990</concept_id>
-         <concept_desc>Security and privacy~Logic and verification</concept_desc>
+  <concept_id>10010520.10010553.10010562</concept_id>
+  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
   <concept_significance>500</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010575.10010755</concept_id>
+  <concept_desc>Computer systems organization~Redundancy</concept_desc>
+  <concept_significance>300</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010553.10010554</concept_id>
+  <concept_desc>Computer systems organization~Robotics</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10003033.10003083.10003095</concept_id>
+  <concept_desc>Networks~Network reliability</concept_desc>
+  <concept_significance>100</concept_significance>
  </concept>
@@
 
-  \ccsdesc[500]{Security and privacy~Logic and verification}
+\ccsdesc[500]{Computer systems organization~Embedded systems}
+\ccsdesc[300]{Computer systems organization~Redundancy}
+\ccsdesc{Computer systems organization~Robotics}
+\ccsdesc[100]{Networks~Network reliability}
 
@@
 %% the work being presented. Separate the keywords with commas.
-\keywords{Verification, SHA-3, OpenSSL, Cryptol, SAW, Symbolic Execution}
+\keywords{datasets, neural networks, gaze detection, text tagging}
+
+%% A "teaser" image appears between the author and affiliation
+%% information and the body of the document, and typically spans the
+%% page.
+% \begin{teaserfigure}
+%   \includegraphics[width=\textwidth]{sampleteaser}
+%   \caption{Seattle Mariners at Spring Training, 2010.}
+%   \Description{Enjoying the baseball game from the third-base
+%   seats. Ichiro Suzuki preparing to bat.}
+%   \label{fig:teaser}
+% \end{teaserfigure}
 
@@
 \section{Introduction}
-ACM's consolidated article template, introduced in 2017, provides a
-consistent \LaTeX\ style for use across ACM publications, and
-incorporates accessibility and metadata-extraction functionality
-necessary for future Digital Library endeavors. Numerous ACM and
-SIG-specific \LaTeX\ templates have been examined, and their unique
-features incorporated into this single new template.
-
-If you are new to publishing with ACM, this document is a valuable
-guide to the process of preparing your work for publication. If you
-have published with ACM before, this document provides insight and
-instruction into more recent changes to the article template.
-
-The ``\verb|acmart|'' document class can be used to prepare articles
-for any ACM publication --- conference or journal, and for any stage
-of publication, from review to final ``camera-ready'' copy, to the
-author's own version, with {\itshape very} few changes to the source.
-
-\section{Template Overview}
-As noted in the introduction, the ``\verb|acmart|'' document class can
-be used to prepare many different kinds of documentation --- a
-double-blind initial submission of a full-length technical paper, a
-two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
-journal article, a SIGCHI Extended Abstract, and more --- all by
-selecting the appropriate {\itshape template style} and {\itshape
-  template parameters}.
-
-This document will explain the major features of the document
-class. For further information, the {\itshape \LaTeX\ User's Guide} is
-available from
-\url{https://www.acm.org/publications/proceedings-template}.
-
-\subsection{Template Styles}
-
-The primary parameter given to the ``\verb|acmart|'' document class is
-the {\itshape template style} which corresponds to the kind of publication
-or SIG publishing the work. This parameter is enclosed in square
-brackets and is a part of the {\verb|documentclass|} command:
-\begin{verbatim}
-  \documentclass[STYLE]{acmart}
-\end{verbatim}
-
-Journals use one of three template styles. All but three ACM journals
-use the {\verb|acmsmall|} template style:
-\begin{itemize}
-\item {\verb|acmsmall|}: The default journal template style.
-\item {\verb|acmlarge|}: Used by JOCCH and TAP.
-\item {\verb|acmtog|}: Used by TOG.
-\end{itemize}
-
-The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
-\begin{itemize}
-\item {\verb|acmconf|}: The default proceedings template style.
-\item{\verb|sigchi|}: Used for SIGCHI conference articles.
-\item{\verb|sigchi-a|}: Used for SIGCHI ``Extended Abstract'' articles.
-\item{\verb|sigplan|}: Used for SIGPLAN conference articles.
-\end{itemize}
-
-\subsection{Template Parameters}
-
-In addition to specifying the {\itshape template style} to be used in
-formatting your work, there are a number of {\itshape template parameters}
-which modify some part of the applied template style. A complete list
-of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}
-
-Frequently-used parameters, or combinations of parameters, include:
-\begin{itemize}
-\item {\verb|anonymous,review|}: Suitable for a ``double-blind''
-  conference submission. Anonymizes the work and includes line
-  numbers. Use with the \verb|\acmSubmissionID| command to print the
-  submission's unique ID on each page of the work.
-\item{\verb|authorversion|}: Produces a version of the work suitable
-  for posting by the author.
-\item{\verb|screen|}: Produces colored hyperlinks.
-\end{itemize}
-
-This document uses the following string as the first command in the
-source file:
-\begin{verbatim}
-\documentclass[sigplan,screen]{acmart}
-\end{verbatim}
-
-\section{Modifications}
-
-Modifying the template --- including but not limited to: adjusting
-margins, typeface sizes, line spacing, paragraph and list definitions,
-and the use of the \verb|\vspace| command to manually adjust the
-vertical spacing between elements of your work --- is not allowed.
-
-{\bfseries Your document will be returned to you for revision if
-  modifications are discovered.}
-
-\section{Typefaces}
-
-The ``\verb|acmart|'' document class requires the use of the
-``Libertine'' typeface family. Your \TeX\ installation should include
-this set of packages. Please do not substitute other typefaces. The
-``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
-as they will override the built-in typeface families.
-
-\section{Title Information}
-
-The title of your work should use capital letters appropriately -
-\url{https://capitalizemytitle.com/} has useful rules for
-capitalization. Use the {\verb|title|} command to define the title of
-your work. If your work has a subtitle, define it with the
-{\verb|subtitle|} command.  Do not insert line breaks in your title.
-
-If your title is lengthy, you must define a short version to be used
-in the page headers, to prevent overlapping text. The \verb|title|
-command has a ``short title'' parameter:
-\begin{verbatim}
-  \title[short title]{full title}
-\end{verbatim}
-
-\section{Authors and Affiliations}
-
-Each author must be defined separately for accurate metadata
-identification. Multiple authors may share one affiliation. Authors'
-names should not be abbreviated; use full first names wherever
-possible. Include authors' e-mail addresses whenever possible.
-
-Grouping authors' names or e-mail addresses, or providing an ``e-mail
-alias,'' as shown below, is not acceptable:
-\begin{verbatim}
-  \author{Brooke Aster, David Mehldau}
-  \email{dave,judy,steve@university.edu}
-  \email{firstname.lastname@phillips.org}
-\end{verbatim}
-
-The \verb|authornote| and \verb|authornotemark| commands allow a note
-to apply to multiple authors --- for example, if the first two authors
-of an article contributed equally to the work.
-
-If your author list is lengthy, you must define a shortened version of
-the list of authors to be used in the page headers, to prevent
-overlapping text. The following command should be placed just after
-the last \verb|\author{}| definition:
-\begin{verbatim}
-  \renewcommand{\shortauthors}{McCartney, et al.}
-\end{verbatim}
-Omitting this command will force the use of a concatenated list of all
-of the authors' names, which may result in overlapping text in the
-page headers.
-
-The article template's documentation, available at
-\url{https://www.acm.org/publications/proceedings-template}, has a
-complete explanation of these commands and tips for their effective
-use.
-
-Note that authors' addresses are mandatory for journal articles.
-
-\section{Rights Information}
-
-Authors of any work published by ACM will need to complete a rights
-form. Depending on the kind of work, and the rights management choice
-made by the author, this may be copyright transfer, permission,
-license, or an OA (open access) agreement.
-
-Regardless of the rights management choice, the author will receive a
-copy of the completed rights form once it has been submitted. This
-form contains \LaTeX\ commands that must be copied into the source
-document. When the document source is compiled, these commands and
-their parameters add formatted text to several areas of the final
-document:
-\begin{itemize}
-\item the ``ACM Reference Format'' text on the first page.
-\item the ``rights management'' text on the first page.
-\item the conference information in the page header(s).
-\end{itemize}
-
-Rights information is unique to the work; if you are preparing several
-works for an event, make sure to use the correct set of commands with
-each of the works.
-
-The ACM Reference Format text is required for all articles over one
-page in length, and is optional for one-page articles (abstracts).
-
-\section{CCS Concepts and User-Defined Keywords}
-
-Two elements of the ``acmart'' document class provide powerful
-taxonomic tools for you to help readers find your work in an online
-search.
-
-The ACM Computing Classification System ---
-\url{https://www.acm.org/publications/class-2012} --- is a set of
-classifiers and concepts that describe the computing
-discipline. Authors can select entries from this classification
-system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
-commands to be included in the \LaTeX\ source.
-
-User-defined keywords are a comma-separated list of words and phrases
-of the authors' choosing, providing a more flexible way of describing
-the research being presented.
-
-CCS concepts and user-defined keywords are required for for all
-articles over two pages in length, and are optional for one- and
-two-page articles (or abstracts).
-
-\section{Sectioning Commands}
-
-Your work should use standard \LaTeX\ sectioning commands:
-\verb|section|, \verb|subsection|, \verb|subsubsection|, and
-\verb|paragraph|. They should be numbered; do not remove the numbering
-from the commands.
-
-Simulating a sectioning command by setting the first word or words of
-a paragraph in boldface or italicized text is {\bfseries not allowed.}
-
-\section{Tables}
-
-The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
-package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
-high-quality tables.
-
-Table captions are placed {\itshape above} the table.
-
-Because tables cannot be split across pages, the best placement for
-them is typically the top of the page nearest their initial cite.  To
-ensure this proper ``floating'' placement of tables, use the
-environment \textbf{table} to enclose the table's contents and the
-table caption.  The contents of the table itself must go in the
-\textbf{tabular} environment, to be aligned properly in rows and
-columns, with the desired horizontal and vertical rules.  Again,
-detailed instructions on \textbf{tabular} material are found in the
-\textit{\LaTeX\ User's Guide}.
-
-Immediately following this sentence is the point at which
-Table~\ref{tab:freq} is included in the input file; compare the
-placement of the table here with the table in the printed output of
-this document.
-
-\begin{table}
-  \caption{Frequency of Special Characters}
-  \label{tab:freq}
-  \begin{tabular}{ccl}
-    \toprule
-    Non-English or Math&Frequency&Comments\\
-    \midrule
-    \O & 1 in 1,000& For Swedish names\\
-    $\pi$ & 1 in 5& Common in math\\
-    \$ & 4 in 5 & Used in business\\
-    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
-  \bottomrule
-\end{tabular}
-\end{table}
-
-To set a wider table, which takes up the whole width of the page's
-live area, use the environment \textbf{table*} to enclose the table's
-contents and the table caption.  As with a single-column table, this
-wide table will ``float'' to a location deemed more
-desirable. Immediately following this sentence is the point at which
-Table~\ref{tab:commands} is included in the input file; again, it is
-instructive to compare the placement of the table here with the table
-in the printed output of this document.
-
-\begin{table*}
-  \caption{Some Typical Commands}
-  \label{tab:commands}
-  \begin{tabular}{ccl}
-    \toprule
-    Command &A Number & Comments\\
-    \midrule
-    \texttt{{\char'134}author} & 100& Author \\
-    \texttt{{\char'134}table}& 300 & For tables\\
-    \texttt{{\char'134}table*}& 400& For wider tables\\
-    \bottomrule
-  \end{tabular}
-\end{table*}
-
-Always use midrule to separate table header rows from data rows, and
-use it only for this purpose. This enables assistive technologies to
-recognise table headers and support their users in navigating tables
-more easily.
-
-\section{Math Equations}
-You may want to display math equations in three distinct styles:
-inline, numbered or non-numbered display.  Each of the three are
-discussed in the next sections.
-
-\subsection{Inline (In-text) Equations}
-A formula that appears in the running text is called an inline or
-in-text formula.  It is produced by the \textbf{math} environment,
-which can be invoked with the usual
-\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
-the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
-and structures, from $\alpha$ to $\omega$, available in
-\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
-examples of in-text equations in context. Notice how this equation:
-\begin{math}
-  \lim_{n\rightarrow \infty}x=0
-\end{math},
-set here in in-line math style, looks slightly different when
-set in display style.  (See next section).
-
-\subsection{Display Equations}
-A numbered display equation---one set off by vertical space from the
-text and centered horizontally---is produced by the \textbf{equation}
-environment. An unnumbered display equation is produced by the
-\textbf{displaymath} environment.
-
-Again, in either environment, you can use any of the symbols and
-structures available in \LaTeX\@; this section will just give a couple
-of examples of display equations in context.  First, consider the
-equation, shown as an inline equation above:
-\begin{equation}
-  \lim_{n\rightarrow \infty}x=0
-\end{equation}
-Notice how it is formatted somewhat differently in
-the \textbf{displaymath}
-environment.  Now, we'll enter an unnumbered equation:
-\begin{displaymath}
-  \sum_{i=0}^{\infty} x + 1
-\end{displaymath}
-and follow it with another numbered equation:
-\begin{equation}
-  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
-\end{equation}
-just to demonstrate \LaTeX's able handling of numbering.
-
-\section{Figures}
-
-The ``\verb|figure|'' environment should be used for figures. One or
-more images can be placed within a figure. If your figure contains
-third-party material, you must clearly identify it as such, as shown
-in the example below.
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{sample-franklin}
-  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
-    Ewing, Inc. [Public domain], via Wikimedia
-    Commons. (\url{https://goo.gl/VLCRBB}).}
-  \Description{A woman and a girl in white dresses sit in an open car.}
-\end{figure}
-
-Your figures should contain a caption which describes the figure to
-the reader.
-
-Figure captions are placed {\itshape below} the figure.
-
-Every figure should also have a figure description unless it is purely
-decorative. These descriptions convey what’s in the image to someone
-who cannot see it. They are also used by search engine crawlers for
-indexing images, and when images cannot be loaded.
-
-A figure description must be unformatted plain text less than 2000
-characters long (including spaces).  {\bfseries Figure descriptions
-  should not repeat the figure caption – their purpose is to capture
-  important information that is not already provided in the caption or
-  the main text of the paper.} For figures that convey important and
-complex new information, a short text description may not be
-adequate. More complex alternative descriptions can be placed in an
-appendix and referenced in a short figure description. For example,
-provide a data table capturing the information in a bar chart, or a
-structured list representing a graph.  For additional information
-regarding how best to write figure descriptions and why doing this is
-so important, please see
-\url{https://www.acm.org/publications/taps/describing-figures/}.
-
-\subsection{The ``Teaser Figure''}
-
-A ``teaser figure'' is an image, or set of images in one figure, that
-are placed after all author and affiliation information, and before
-the body of the article, spanning the page. If you wish to have such a
-figure in your article, place the command immediately before the
-\verb|\maketitle| command:
-\begin{verbatim}
-  \begin{teaserfigure}
-    \includegraphics[width=\textwidth]{sampleteaser}
-    \caption{figure caption}
-    \Description{figure description}
-  \end{teaserfigure}
-\end{verbatim}
-
-\section{Citations and Bibliographies}
-
-The use of \BibTeX\ for the preparation and formatting of one's
-references is strongly recommended. Authors' names should be complete
---- use full first names (``Donald E. Knuth'') not initials
-(``D. E. Knuth'') --- and the salient identifying features of a
-reference should be included: title, year, volume, number, pages,
-article DOI, etc.
-
-The bibliography is included in your source document with these two
-commands, placed just before the \verb|\end{document}| command:
-\begin{verbatim}
-  \bibliographystyle{ACM-Reference-Format}
-  \bibliography{bibfile}
-\end{verbatim}
-where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
-suffix, of the \BibTeX\ file.
-
-Citations and references are numbered by default. A small number of
-ACM publications have citations and references formatted in the
-``author year'' style; for these exceptions, please include this
-command in the {\bfseries preamble} (before the command
-``\verb|\begin{document}|'') of your \LaTeX\ source:
-\begin{verbatim}
-  \citestyle{acmauthoryear}
-\end{verbatim}
-
-  Some examples.  A paginated journal article \cite{Abril07}, an
-  enumerated journal article \cite{Cohen07}, a reference to an entire
-  issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
-  monograph/whole book in a series (see 2a in spec. document)
-  \cite{Harel79}, a divisible-book such as an anthology or compilation
-  \cite{Editor00} followed by the same example, however we only output
-  the series if the volume number is given \cite{Editor00a} (so
-  Editor00a's series should NOT be present since it has no vol. no.),
-  a chapter in a divisible book \cite{Spector90}, a chapter in a
-  divisible book in a series \cite{Douglass98}, a multi-volume work as
-  book \cite{Knuth97}, a couple of articles in a proceedings (of a
-  conference, symposium, workshop for example) (paginated proceedings
-  article) \cite{Andler79, Hagerup1993}, a proceedings article with
-  all possible elements \cite{Smith10}, an example of an enumerated
-  proceedings article \cite{VanGundy07}, an informally published work
-  \cite{Harel78}, a couple of preprints \cite{Bornmann2019,
-    AnzarootPBM14}, a doctoral dissertation \cite{Clarkson85}, a
-  master's thesis: \cite{anisi03}, an online document / world wide web
-  resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game
-  (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05}
-  and (Case 3) a patent \cite{JoeScientist001}, work accepted for
-  publication \cite{rous08}, 'YYYYb'-test for prolific author
-  \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
-  contain 'duplicate' DOI and URLs (some SIAM articles)
-  \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
-  multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
-  couple of citations with DOIs:
-  \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
-  citations: \cite{TUGInstmem, Thornburg01, CTANacmart}. Artifacts:
-  \cite{R} and \cite{UMassCitations}.
-
-\section{Acknowledgments}
-
-Identification of funding sources and other support, and thanks to
-individuals and groups that assisted in the research and the
-preparation of the work should be included in an acknowledgment
-section, which is placed just before the reference section in your
-document.
-
-This section has a special environment:
-\begin{verbatim}
-  \begin{acks}
-  ...
-  \end{acks}
-\end{verbatim}
-so that the information contained therein can be more easily collected
-during the article metadata extraction phase, and to ensure
-consistency in the spelling of the section heading.
-
-Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.
-
-\section{Appendices}
-
-If your work needs an appendix, add it before the
-``\verb|\end{document}|'' command at the conclusion of your source
-document.
-
-Start the appendix with the ``\verb|appendix|'' command:
-\begin{verbatim}
-  \appendix
-\end{verbatim}
-and note that in the appendix, sections are lettered, not
-numbered. This document has two appendices, demonstrating the section
-and subsection identification method.
-
-\section{SIGCHI Extended Abstracts}
-
-The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
-not in Word) produces a landscape-orientation formatted article, with
-a wide left margin. Three environments are available for use with the
-``\verb|sigchi-a|'' template style, and produce formatted output in
-the margin:
-\begin{itemize}
-\item {\verb|sidebar|}:  Place formatted text in the margin.
-\item {\verb|marginfigure|}: Place a figure in the margin.
-\item {\verb|margintable|}: Place a table in the margin.
-\end{itemize}
 
@@
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{sample-base}
-
-%%
-%% If your work has an appendix, this is the place to put it.
-\appendix
-
-\section{Research Methods}
-
-\subsection{Part One}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
-malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
-sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
-vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
-lacinia dolor. Integer ultricies commodo sem nec semper.
-
-\subsection{Part Two}
-
-Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
-ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
-ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
-eros. Vivamus non purus placerat, scelerisque diam eu, cursus
-ante. Etiam aliquam tortor auctor efficitur mattis.
-
-\section{Online Resources}
-
-Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
-pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
-enim maximus. Vestibulum gravida massa ut felis suscipit
-congue. Quisque mattis elit a risus ultrices commodo venenatis eget
-dui. Etiam sagittis eleifend elementum.
-
-Nam interdum magna at lectus dignissim, ac dignissim lorem
-rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
-massa et mattis lacinia.
+\bibliography{paper.bib}
 
@@
 %%
-%% End of file `sample-sigplan.tex'.
+%% End of file `paper.tex'.

### 2021-09-04 paper/abstract.tex
Add paper basic structure

\begin{abstract}
The Keccak function in the SHA-3 NIST standard is responsible for the major computations of all SHA-3 hashes and Shake functions.
These functions are a key piece in modern secure hash functions. 
This work extends and improves on the methodology used in the verification of Amazon’s s2n HMAC and OpenSSL’s SHA2-256 inner hash function. 
It uses the Software Analysis Workbench (SAW) to prove out functional equivalence between OpenSSL’s SHA-3 Keccak function and a verified implementation of SHA-3 in the domain-specific language Cryptol. 
SAW creates contracts (called overrides), a functions input to output relationship, which allows for a modular approach to writing proofs. 
This work leverages SAW by creating overrides to block out computation and creating a compositional proof that remains within the capacity of an SMT solver. 
This work amends Cryptol through utility functions to handle looping which greatly simplifies the visual inspection the Cryptol implementation to provide strong assurance of correctness. 
The results of this work indicate that this methodology works with modern cryptographic specifications and that a modular approach in cryptographic algorithm design greatly simplifies the verification process.
\end{abstract}

### 2021-09-04 paper/amended.tex
Add paper basic structure

\section{Amended Cryptol}\label{sec:amended}
Parker, will you start pulling your stuff in here

### 2021-09-04 paper/conclusion.tex
Add paper basic structure

\section{Conclusion}\label{sec:conclusion}
Conclude things!

### 2021-09-04 paper/introduction.tex
Add paper basic structure

\section{Introduction}\label{sec:introduction}
  The Secure Hashing Algorithm (SHA) family of cryptography is widely used as a means of providing increased security through generating/verifying digital signatures, deriving keys, and generating pseudorandom bits. (NIST: https://csrc.nist.gov/publications/detail/fips/202/final)  
  The structure and computations of these algorithms is defined by NIST in mathematical specifications.  
  Because these specifications are just figures on a page, any assurance of correctness of an algorithm’s implementation is dependent on a visual review of the code coupled with a small sampling of test vectors.

  The reliability of this correctness check is often weakened because of the difficulty of performing that visual review.  
  Many implementations of the SHA algorithms go through major rewrites and optimizations which make visual inspection of the code nearly impossible. (Not sure if that’s too strong a claim).
  This work bridges that gap between optimized c code, and the NIST standard.

  \subsection{Related Work}
  Previously, Galois’ Inc. worked with Amazon to prove out parts of Amazon’s s2n HMAC code to the NIST specification.  
In their work, they stubbed out use of the SHA2 algorithm, because their HMAC allows for customization of the algorithm type and implementation.

The work by Decker et. al, extended on Galois' work by choosing a single SHA2 implementation to prove out.  
They chose OpenSSL’s SHA 256 implementation, as that is a commonly used library (and one that is usable in s2n).  
They were able to prove out the inner hash computations of that implementation using a methodology of compositional proofs to do so.  
They were also able to connect up their proof with the code used by Galois’ to stub out the SHA2 algorithm, which implies that the functional equivalence property established with the OpenSSL code applies to their Cryptol implementation as well.  
Their work showed the viability and limitations of creating proofs of functional equivalence for legacy cryptographic implementations.

\subsection{SHA 3}
Modern Cryptography.  Talk about what it is and it's uses.

\subsection{Results}
Brief description of our results.

### 2021-09-04 paper/iota.tex
Add paper basic structure

\section{Static Iota}\label{sec:iota}
Talk about changes we made to iota to prove out directly between precomputed and computed on the fly.

### 2021-09-04 paper/paper.tex
Add no reverse cryptol properties
Add paper basic structure

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{The Name of the Title is Hope}
+\title{Verification of OpenSSL's SHA3 Keccak function using the Software Analysis Workbench}
 
@@
 %% of authors' names for this purpose.
-\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
 
@@
 %% article.
-\begin{abstract}
-  A clear and well-documented \LaTeX\ document is presented as an
-  article formatted for publication by ACM in a conference proceedings
-  or journal publication. Based on the ``acmart'' document class, this
-  article presents and explains many of the common variations, as well
-  as many of the formatting elements an author may use in the
-  preparation of the documentation of their work.
-\end{abstract}
 
-%%
+\input{abstract}
+
+%%NEED TO COMPLETE THIS BEFORE SUBMISSION
 %% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
@@
 %% the work being presented. Separate the keywords with commas.
-\keywords{datasets, neural networks, gaze detection, text tagging}
+\keywords{verification, cryptography, symbolic execution, SHA3}
 
@@
 
-\section{Introduction}
+\input{introduction}
+\input{preliminaries}
+\input{proof}
+\input{amended}
+\input{iota}
+\input{visual}
+\input{conclusion}
 

### 2021-09-04 paper/preliminaries.tex
Add paper basic structure

\section{Preliminaries}\label{sec:preliminaries}

\subsection{Tools}
The Software Analysis Workbench (SAW) provides a verification environment for compositional proofs of functional equivalence between two pieces of supported code.  
SAW uses symbolic execution to create computational models of the code and creates problems of functional equivalence between those two models that can be pushed off to an SMT solver.  
The difficulty of these SMT problems is directly affected by the size and complexity of the computations that are being compared.  
This can lead to drastic increases in proof times and sometimes even prevents completion as computations increase.  
SAW’s solution to this is to create contracts (called overrides) of a functions input to output relationship which allows for a modular approach to writing the proof.  
By breaking up computation into smaller pieces, and creating overrides for these pieces, SAW can often simplify an otherwise complex (and thus incompletable) proof to remain within the capacity of an SMT solver.

CRYPTOL!!!!!!

\subsection{Trusted Code Base}
This work relies on the correctness of several tools, including language compilers.  
Our Trusted Code Base consists of:
\begin{itemize}
  \item LLVM bitcode compiler
  \item The Software Analysis Workbench (SAW) including the conversion from LLVM bitcode to SAWCore terms, the Cryptol compiler to SAWCore terms, and SAW's interface with SMT solvers.
  \item The z3, abc, and yices SMT solvers.
\end{itemize}

### 2021-09-04 paper/proof.tex
Add paper basic structure

\section{Proof Structure}\label{sec:proof}
Talk about our proof flow
\subsection{Modular Approach}
Talk about how NIST standard/openssl being modular simplified our proof structure.

### 2021-09-04 paper/visual.tex
Add paper basic structure

\section{Visual Limitations}\label{sec:visual}
Talk about visual Limitations
\subsection{Iota}
Have to make small concession because we do prove that it computes all 24 values using rc, but it does it in one swoop still, rather than as needed.

Assert function is not useable in the Symbolic Execution Engine.
To fix this, we replace it with an if statement that checks the same expression and only runs code if it returns true.
\subsection{Array Storage}
OpenSSL switches the ordering of their x/y's for the array.
Because we're proving the small inner parts of the code, we have to match memory structures.

Whatever we decided for the bit reversal (still might try to get a cryptol->cryptol proof running).

### 2021-09-07 paper/introduction.tex
fix typos

@@
   Previously, Galois’ Inc. worked with Amazon to prove out parts of Amazon’s s2n HMAC code to the NIST specification.  
-In their work, they stubbed out use of the SHA2 algorithm, because their HMAC allows for customization of the algorithm type and implementation.
+In their work they stubbed out use of the SHA2 algorithm because their HMAC allows for customization of the algorithm type and implementation.
 
@@
 They were able to prove out the inner hash computations of that implementation using a methodology of compositional proofs to do so.  
-They were also able to connect up their proof with the code used by Galois’ to stub out the SHA2 algorithm, which implies that the functional equivalence property established with the OpenSSL code applies to their Cryptol implementation as well.
+They were also able to connect up their proof with the code used by Galois’ to stub out the SHA2 algorithm, which implies that the functional equivalence property established with the OpenSSL code applies to Galois' Cryptol implementation as well.
 Their work showed the viability and limitations of creating proofs of functional equivalence for legacy cryptographic implementations.

### 2021-09-07 paper/introduction.tex
Add sha3/results

@@
 \subsection{SHA 3}
-Modern Cryptography.  Talk about what it is and it's uses.
+This work shows that this same methodology can be extended to modern cryptographic algorithms.
+The SHA 3 family of algorithms is defined by NIST in the FIPS 202 publication.
+It specifies a block hashing algorithm, Keccak, that is used by each of the different SHA and SHAKE algorithms.
+In comparison to the SHA 2 family of algorithms, Keccak features significantly more complex and robust computations.
+This increase in computation was of interest, because it presses the limits of SAW's capabilities.
 
 \subsection{Results}
-Brief description of our results.
+This work shows a complete proof of functional equivalence from a verified implementation of the NIST standard to OpenSSL's C code for the Keccak function.
+In doing so, it shows SAW's ability to stand up against more difficult computations.
+It also provides insights into using Cryptol's functional nature to further simplify the step of visually verifying Cryptol code to the NIST standard.

### 2021-09-08 paper/amended.tex
Update keccak_spec_no_reverse.cry to align with NIST. Update amended.tex.

@@
 \section{Amended Cryptol}\label{sec:amended}
-Parker, will you start pulling your stuff in here
+TODO:
+
+Imperative spec, functional language, Cryptol's looping syntax can be cryptic.
+
+Rational and descriptions of utility functions, examples of utility functions.
+
+for-each == javascript map, while loop == javascript reduce.
+
+Initially it appeared that visual inspection with Cryptol’s standard syntax was subpar.
+Cryptol’s functional folds and list comprehensions do not align easily with the imperative 
+and array-based NIST standard. BY leveraging Cryptol’s versatility, however, we produced a couple 
+utility functions to visually mimic the sequential nature of the specification.
+
+The phrase “for all” is quite common in the NIST document, and it implies a looping structure. 
+The following is the NIST definition of the pi function:
+\begin{verbatim}
+  1. For all triples (x,y,z) such that 
+    0 <= x < 5, 0 <= y < 5, and 0 <= z < w,
+    let A'[x, y, z] = A[(x + 3y) mod 5, x, z].
+  2. Return A'.
+\end{verbatim}
+
+Using Cryptol's list comprehensions, we find the following syntax for the pi function:
+\begin{verbatim}
+  type COL = 5 
+  type W = 64
+  LIST4 = [0..COL-1]
+  LIST63 = [0..W-1]
+  type STATE_ARR = [COL][COL][W]
+  pi : STATE_ARR -> STATE_ARR
+  pi a = [[[a @x @((x + 3y) % COL) @z 
+             | z <- LIST63] 
+           | y <- LIST4] 
+         | x <- LIST4]
+\end{verbatim}
+
+Cryptol declares list comprehension variables after they are used, and the order of the
+nested loops is not as apparent. Using a helper function, ‘for’, we were able to provide
+the following syntax:
+
+\begin{verbatim}
+  type COL = 5 
+  type W = 64
+  type STATE_ARR = [COL][COL][W]
+  LIST4 = [0..COL-1]
+  LIST63 = [0..W-1]
+  pi : STATE_ARR -> STATE_ARR
+  pi a = for LIST4 (\y ->
+           for LIST4 (\x -> 
+             for LIST4 (\z -> 
+               a @x @((x + 3*y) % `COL) @z)))
+\end{verbatim}
+
+  The ‘for’ utility function follows the same intuition as the implementation in most 
+  imperative languages. That is, it iterates over the list provided. It is similar to the 
+  Array.prototype.map method in JavaScript in that it takes in a function and a list of 
+  arguments and returns a list of results from calling the function on every element of the 
+  input array. It is implemented as so:
+
+\begin{verbatim}
+  for : {n, a, b} [n]a -> (a -> b) -> [n]b
+  for vals loop = [loop index | index <- vals]
+\end{verbatim}
+
+While simple to implement, it reorders and structures the arguments in a very elegant way.
+In constrast with standard cryptol syntax, the list to iterate over and the iterating 
+variable are visible before the loop body. List comprehensions are hidden away behind this 
+function call and visual equivalence is clearer.
+
+Another common phrasing is the “let...for...let”. This defines a state upon which a folding 
+occurs, which is most comparable to a while loop. Here is the rc function as defined in the 
+NIST spec:
+
+\begin{verbatim}
+  1. If t mod 255 = 0, return 1.
+  2. Let R = 10000000.
+  3. For i from 1 to t mod 255, let:
+    a. R = 0 || R;
+    b. R[0] = R[0] ^ R[8];
+    c. R[4] = R[4] ^ R[8];
+    d. R[5] = R[5] ^ R[8];
+    e. R[6] = R[6] ^ R[8];
+    f. R =Trunc8[R].
+  4. Return R[0].
+\end{verbatim}
+
+Cryptol list comprehensions require a list to iterate over. Our ‘for’ utility function does 
+as well. Knowing that the number of iterations is bounded by [1, 254], we can construct the 
+following ungainly Cryptol definition:
+
+\begin{verbatim}
+  rc : [64] -> Bit
+  rc t = if (t % 255) == 0 
+    then True 
+      else (rList !0).R @0 where
+        rList = [{i = 1, R = 128:[8]}] # [{
+          i = r @i + 1, 
+          R = if i <= t % 255
+            then (take`{8} ([
+              (r@0)^(r@8),
+              r@1,
+              r@2,
+              r@3, 
+              (r@4)^(r@8),
+              (r@5)^(r@8),
+              (r@6)^(r@8),
+              r@7,
+              r@8])
+                where r = [False] # state.R)
+            else (rList @i).R}
+          | i <- [0..254]]
+\end{verbatim}
+
+Not only is this computationally slower since it runs 255 loops every time rc is called,
+it differs vastly from the NIST specificiation. Using a new ‘while’ utility function, we 
+can declare an initial state, condition function, and body function that will recursively 
+be called until the condition is met. Here is the rc function with the use of this utility 
+function:
+
+\begin{verbatim}
+  rc : [64] -> Bit
+  rc t = if (t % 255) == 0 
+    then True 
+    else (while {i = 1, R = 128:[8]} //STATE
+      (\state -> state.i <= t % 255) //COND
+      (\state -> {                   //LOOP
+        i = state.i + 1, 
+        R = (take`{8} ([
+          (r@0)^(r@8),
+          r@1,
+          r@2,
+          r@3,
+          (r@4)^(r@8),
+          (r@5)^(r@8),
+          (r@6)^(r@8),
+          r@7, 
+          r@8]) 
+            where r = [False] # state.R)})
+      ).R @0
+\end{verbatim}
+
+The final state is returned by the ‘while’ function. This iterates only until the 
+condition is met, avoids the clutter of indexing within a larger array of all previous 
+states, and places the condition in a more obvious location before the loop body. Here 
+is the definition of the ‘while’ utility function:
+
+\begin{verbatim}
+  while : {a} a -> (a -> Bit) -> (a -> a) -> a
+  while state cond f = if (cond state)
+    then (while (f state) cond f)	
+    else state
+\end{verbatim}
+
+While the condition is not met by the current state, continue to iterate. Else, return the 
+current state.

### 2021-09-10 paper/introduction.tex
Add contributions to the introduction

@@
   This work bridges that gap between optimized c code, and the NIST standard.
+
+\noindent \textbf{Contributions}:
+This paper details the following contributions.
+\begin{compactitem}
+  \item  A Cryptol specification for \texttt{SHA3} that directly corresponds to the \texttt{FIPS} 202 specification and passes the test vectors provided by the same specification.
+  \item A Cryptol specification based on the OpenSSL implementation of \texttt{SHA3} that uses a reordered state vector and precomputed tables in Keccak for performance gains and passes test vectors
+  \item A \texttt{SAW} proof that the \texttt{FIPS} 202 specification is equivalent to the OpenSSL specification
+  \item A \texttt{SAW} proof that the OpenSSL Keccak specification is equivalent to the OpenSSL C-implementation of the same function for all possible state vector input
+  \item Early evidence that modular cryptographic specifications such as \texttt{SHA3} lend themselves to efficient \texttt{SAW} verification against C implementations as opposed to flattened specifications such as \texttt{SHA2}
+\end{compactitem}
 

### 2021-09-10 paper/amended.tex
Update amended.tex

@@
 \section{Amended Cryptol}\label{sec:amended}
-TODO:
-
-Imperative spec, functional language, Cryptol's looping syntax can be cryptic.
-
-Rational and descriptions of utility functions, examples of utility functions.
-
-for-each == javascript map, while loop == javascript reduce.
-
 Initially it appeared that visual inspection with Cryptol’s standard syntax was subpar.
 Cryptol’s functional folds and list comprehensions do not align easily with the imperative 
-and array-based NIST standard. BY leveraging Cryptol’s versatility, however, we produced a couple 
-utility functions to visually mimic the sequential nature of the specification.
+and array-based NIST standard. By leveraging Cryptol’s versatility, however, we produced a couple 
+utility functions to visually mimic the sequential nature of the NIST specification.
 
-The phrase “for all” is quite common in the NIST document, and it implies a looping structure. 
-The following is the NIST definition of the pi function:
+The phrase “for all” indicates a looping structure and is quite common in the NIST standard. 
+It is used in the definition of the pi function.
+
+\begin{figure}[h]
+  \centering
   \begin{verbatim}
@@
   \end{verbatim}
+  \caption{NIST Pi definition}
+  \label{fig:nistPi}
+\end{figure}
 
 Using Cryptol's list comprehensions, we find the following syntax for the pi function:
+
+\begin{figure}[h]
+  \centering
 \begin{verbatim}
@@
 type W = 64
+type STATE_ARR = [COL][COL][W]
 LIST4 = [0..COL-1]
 LIST63 = [0..W-1]
-  type STATE_ARR = [COL][COL][W]
 pi : STATE_ARR -> STATE_ARR
@@
 \end{verbatim}
+\caption{Cryptol Pi definition}
+\label{fig:cryptolPi}
+\end{figure}
 
@@
 
+\begin{figure}[h]
+  \centering
 \begin{verbatim}
@@
 \end{verbatim}
+\caption{Ammended Cryptol Pi definition}
+\label{fig:cryptolAmmendedPi}
+\end{figure}
 
@@
 \end{verbatim}
+\caption{Cryptol For utility function}
+\label{fig:cryptolFor}
+\end{figure}
 
@@
 \end{verbatim}
+\caption{NIST RC definition}
+\label{fig:nistRC}
+\end{figure}
 
@@
   then True 
-      else (rList !0).R @0 where
-        rList = [{i = 1, R = 128:[8]}] # [{
-          i = r @i + 1, 
-          R = if i <= t % 255
+  else rs !0 @0 where
+    rs = [128:[8]] # [
+      if i <= t % 255
         then (take`{8} ([
@@
           r@8])
-                where r = [False] # state.R)
-            else (rList @i).R}
-          | i <- [0..254]]
+            where r = [False] # (rs @(i - 1)))
+        else rs @(i - 1)
+      | i <- [1..254]]
 \end{verbatim}
+\caption{Cryptol RC definition}
+\label{fig:cryptolRC}
+\end{figure}
 
 Not only is this computationally slower since it runs 255 loops every time rc is called, 
-it differs vastly from the NIST specificiation. Using a new ‘while’ utility function, we 
+is stores the previous states in a list and indexes into it. This differs vastly from the 
+NIST specificiation both visually and functionally. Using a new ‘while’ utility function, we 
 can declare an initial state, condition function, and body function that will recursively 
-be called until the condition is met. Here is the rc function with the use of this utility 
+be called until the condition is met. It consolodates the information of the previous state 
+until the condition is sastified. Here is the rc function with the use of this utility 
 function:
 
+\begin{figure}[h]
+  \centering
 \begin{verbatim}
@@
 \end{verbatim}
+\caption{Ammended Cryptol RC definition}
+\label{fig:cryptolAmmendedRC}
+\end{figure}
 
@@
 \end{verbatim}
-
-While the condition is not met by the current state, continue to iterate. Else, return the 
-current state.
+\caption{Cryptol While utility function}
+\label{fig:cryptolWhile}
+\end{figure}

### 2021-09-10 paper/abstract.tex
update abstract and figure

@@
 \begin{abstract}
-The Keccak function in the SHA-3 NIST standard is responsible for the major computations of all SHA-3 hashes and Shake functions.
-These functions are a key piece in modern secure hash functions. 
-This work extends and improves on the methodology used in the verification of Amazon’s s2n HMAC and OpenSSL’s SHA2-256 inner hash function. 
-It uses the Software Analysis Workbench (SAW) to prove out functional equivalence between OpenSSL’s SHA-3 Keccak function and a verified implementation of SHA-3 in the domain-specific language Cryptol.
-SAW creates contracts (called overrides), a functions input to output relationship, which allows for a modular approach to writing proofs. 
-This work leverages SAW by creating overrides to block out computation and creating a compositional proof that remains within the capacity of an SMT solver. 
-This work amends Cryptol through utility functions to handle looping which greatly simplifies the visual inspection the Cryptol implementation to provide strong assurance of correctness. 
-The results of this work indicate that this methodology works with modern cryptographic specifications and that a modular approach in cryptographic algorithm design greatly simplifies the verification process.
+The Keccak function in the SHA-3 NIST standard is responsible for the major computations of all SHA-3 hash constructs.
+These constructs form the basis for modern secure hashing.
+This work extends and improves the methodology used to verify Amazon’s s2n HMAC and OpenSSL’s SHA2-256 inner hash function. 
+It uses the Software Analysis Workbench (SAW) to prove out functional equivalence between OpenSSL’s SHA-3 Keccak function and a verified implementation of SHA-3 in Cryptol, a domain-specific language for cryptography.
+SAW allows for a modular approach to writing proofs through the use of contracts, which define a functions input to output relationship.
+Each contract, also called an override, is an obligation in the full proof.
+This work leverages SAW by creating overrides to decompose computation and create a compositional proof that is within the capacity of an SMT solver. 
+Visual inspection between Cryptol and the NIST standard is simplified through the creation of Cryptol utility functions.
+These utility functions wrap the complex list comprehensions required for looping in Cryptol.
+The results of this work indicates that a modular approach in cryptographic algorithm design greatly simplifies the verification process, as exemplified by SHA-3.
 \end{abstract}

### 2021-09-10 paper/introduction.tex
Add limitations to the introduction
update abstract and figure

@@
   This work bridges that gap between optimized c code, and the NIST standard.
-
+  \input{proofFigure.tex}
 \noindent \textbf{Contributions}:
@@
   \item Early evidence that modular cryptographic specifications such as \texttt{SHA3} lend themselves to efficient \texttt{SAW} verification against C implementations as opposed to flattened specifications such as \texttt{SHA2}
+\end{compactitem}
+
+\noindent \textbf{Limitations}
+\begin{compactitem}
+  \item Visual inspection and limited test vectors for (1) and (2)
+  \item Only provide proof certificates for message sizes of X to Y for digest sizes of … for (3)
+  \item Visual inspection only for anything above Keccak for (3) in OpenSSLs C-code
 \end{compactitem}
@@
 It also provides insights into using Cryptol's functional nature to further simplify the step of visually verifying Cryptol code to the NIST standard.
+
+Because of how NIST defines the standard, the keccak function is used for each digest size of the SHA3 algorithm, with the only significant differences being in the padding/sponging parts of the algorithm, which falls outside this work's scope.
+This means that the proof of correctness for this function extends usefulness beyond just the SHA3-256 level which we focused on.
+In our repository, we wrote the Cryptol specification to be able to run any of the SHA3 family of algorithms using the same keccak.
+A simple extension of this work would be to take that already created and verified Cryptol specification, and prove it up against OpenSSL's slightly modified keccak versions.

### 2021-09-10 paper/amended.tex

@@
-\section{Amended Cryptol}\label{sec:amended}
+\section{Amended Cryptol}
+\label{sec:amended}
 Initially it appeared that visual inspection with Cryptol’s standard syntax was subpar.
-Cryptol’s functional folds and list comprehensions do not align easily with the imperative 
+The language's functional folds and list comprehensions do not align easily with the imperative 
 and array-based NIST standard. By leveraging Cryptol’s versatility, however, we produced a couple 
@@
 pi : STATE_ARR -> STATE_ARR
-pi a = for LIST4 (\y ->
-          for LIST4 (\x -> 
+pi a = for LIST4 (\x ->
+          for LIST4 (\y -> 
             for LIST4 (\z -> 
@@
 \end{verbatim}
-\caption{Ammended Cryptol Pi definition}
-\label{fig:cryptolAmmendedPi}
+\caption{Amended Cryptol Pi definition}
+\label{fig:cryptolAmendedPi}
 \end{figure}
@@
 The ‘for’ utility function follows the same intuition as the implementation in most 
-imperative languages. That is, it iterates over the list provided. It is similar to the 
-Array.prototype.map method in JavaScript in that it takes in a function and a list of 
-arguments and returns a list of results from calling the function on every element of the 
-input array. It is implemented as so:
+imperative languages in that it iterates and runs a loop body over defined elements. 
+It is similar to the Array.prototype.map method in JavaScript in that it takes in a 
+function and a list of arguments and returns a list of results from calling the function 
+on every element of the input array. It is implemented as so:
 
@@
         then (take`{8} ([
-          (r@0)^(r@8),
+          (r@0) ^ (r@8),
           r@1,
@@
           r@3, 
-          (r@4)^(r@8),
-          (r@5)^(r@8),
-          (r@6)^(r@8),
+          (r@4) ^ (r@8),
+          (r@5) ^ (r@8),
+          (r@6) ^ (r@8),
           r@7,
@@
       R = (take`{8} ([
-        (r@0)^(r@8),
+        (r@0) ^ (r@8),
         r@1,
@@
         r@3,
-        (r@4)^(r@8),
-        (r@5)^(r@8),
-        (r@6)^(r@8),
+        (r@4) ^ (r@8),
+        (r@5) ^ (r@8),
+        (r@6) ^ (r@8),
         r@7, 
@@
 \end{verbatim}
-\caption{Ammended Cryptol RC definition}
-\label{fig:cryptolAmmendedRC}
+\caption{Amended Cryptol RC definition}
+\label{fig:cryptolamendedRC}
 \end{figure}

### 2021-09-11 paper/amended.tex
Update paper.tex and amended.tex Add Cryptol lstlisting highlighting.

@@
 
+\subsection{For Utility Function}
 The phrase “for all” indicates a looping structure and is quite common in the NIST standard. 
-It is used in the definition of the pi function.
+It is used in the definition of the pi function shown in Figure \ref{fig:nistPi}.
 
@@
   \centering
-  \begin{verbatim}
+  \begin{lstlisting}[basewidth = {.5em}]
 1. For all triples (x,y,z) such that 
@@
 2. Return A'.
-  \end{verbatim}
+  \end{lstlisting}
   \caption{NIST Pi definition}
@@
 
-Using Cryptol's list comprehensions, we find the following syntax for the pi function:
+By nesting Cryptol's list comprehensions, we easily implemented the pi function in Figure 
+\ref{fig:cryptolPi}.
 
@@
   \centering
-\begin{verbatim}
+\begin{lstlisting}[language=Cryptol]
 type COL = 5 
@@
 type STATE_ARR = [COL][COL][W]
-LIST4 = [0..COL-1]
-LIST63 = [0..W-1]
+LIST4 = [0..COL - 1]
+LIST63 = [0..W - 1]
 pi : STATE_ARR -> STATE_ARR
-pi a = [[[a @x @((x + 3y) % COL) @z 
+pi a = [[[a @x @((x + 3*y) % 'COL) @z 
             | z <- LIST63] 
@@
         | x <- LIST4]
-\end{verbatim}
+\end{lstlisting}
 \caption{Cryptol Pi definition}
@@
 nested loops is not as apparent. Using a helper function, ‘for’, we were able to provide
-the following syntax:
+a more clear definition seen in Figure \ref{fig:cryptolAmendedPi}.
 
@@
 type STATE_ARR = [COL][COL][W]
-LIST4 = [0..COL-1]
-LIST63 = [0..W-1]
+LIST4 = [0..COL - 1]
+LIST63 = [0..W - 1]
 pi : STATE_ARR -> STATE_ARR
-pi a = for LIST4 (\x ->
-          for LIST4 (\y -> 
-            for LIST4 (\z -> 
-              a @x @((x + 3*y) % `COL) @z)))
-\end{verbatim}
+pi a = for LIST4 (&x ->
+          for LIST4 (&y -> 
+            for LIST4 (&z -> 
+              a @x @((x + 3*y) % 'COL) @z)))
+\end{lstlisting}
 \caption{Amended Cryptol Pi definition}
@@
 function and a list of arguments and returns a list of results from calling the function 
-on every element of the input array. It is implemented as so:
+on every element of the input array. Figure \ref{fig:cryptolFor} shows it's implementation.
 
@@
   \centering
-\begin{verbatim}
+\begin{lstlisting}[language=Cryptol]
 for : {n, a, b} [n]a -> (a -> b) -> [n]b
 for vals loop = [loop index | index <- vals]
-\end{verbatim}
+\end{lstlisting}
 \caption{Cryptol For utility function}
@@
 
-While simple to implement, it reorders and structures the arguments in a very elegant way.
+Though it is simple to implement, it reorders and structures the arguments in a very elegant way.
 In constrast with standard cryptol syntax, the list to iterate over and the iterating 
@@
 
+\subsection{While Utility Function}
 Another common phrasing is the “let...for...let”. This defines a state upon which a folding 
-occurs, which is most comparable to a while loop. Here is the rc function as defined in the 
-NIST spec:
+occurs, iteratively condensing the information into a final state. Imperatively, it is most 
+comparable to a while loop. As shown by Figure \ref{fig:nistRC}, is present in the rc function 
+defined by the NIST specification.
 
@@
   \centering
-\begin{verbatim}
+\begin{lstlisting}[basewidth = {.5em}]
 1. If t mod 255 = 0, return 1.
@@
 4. Return R[0].
-\end{verbatim}
+\end{lstlisting}
 \caption{NIST RC definition}
@@
 
-Cryptol list comprehensions require a list to iterate over. Our ‘for’ utility function does 
-as well. Knowing that the number of iterations is bounded by [1, 254], we can construct the 
-following ungainly Cryptol definition:
+Maintaining this state through Cryptol's list comprehensions can be difficult.
+The comprehensions require a list to iterate through.Since our ‘for’ utility 
+function is simply a wrapper for list comprehensions, it is equally unsuited for 
+visual verification on such steps. The condition to stop iterating in the RC function
+bounds the number of iterations by [1, 254], which lets us construct such a list to 
+iterate through. This leads to a rather ungainly Cryptol definition, however, as 
+shown by Figure \ref{fig:cryptolRC}.
 
@@
   \centering
-\begin{verbatim}
+\begin{lstlisting}[language=Cryptol]
 rc : [64] -> Bit
@@
       if i <= t % 255
-        then (take`{8} ([
+        then (take'{8} ([
           (r@0) ^ (r@8),
@@
       | i <- [1..254]]
-\end{verbatim}
+\end{lstlisting}
 \caption{Cryptol RC definition}
@@
 
-Not only is this computationally slower since it runs 255 loops every time rc is called, 
-is stores the previous states in a list and indexes into it. This differs vastly from the 
+Not only is this definition computationally slower as it runs 255 loops every time it is called, 
+it also awkwardly indexes into a list of previous states. This differs vastly from the 
 NIST specificiation both visually and functionally. Using a new ‘while’ utility function, we 
@@
 be called until the condition is met. It consolodates the information of the previous state 
-until the condition is sastified. Here is the rc function with the use of this utility 
-function:
+until the condition is sastified. This definition is foudn in Figure \ref{fig:cryptolamendedRC}.
 
@@
   then True 
-  else (while {i = 1, R = 128:[8]} //STATE
-    (\state -> state.i <= t % 255) //COND
-    (\state -> {                   //LOOP
+  else (while {i = 1, R = 128:[8]}
+    (&state -> state.i <= t % 255)
+    (&state -> {
       i = state.i + 1, 
-      R = (take`{8} ([
+      R = (take'{8} ([
         (r@0) ^ (r@8),
@@
     ).R @0
-\end{verbatim}
+\end{lstlisting}
 \caption{Amended Cryptol RC definition}
@@
 condition is met, avoids the clutter of indexing within a larger array of all previous 
-states, and places the condition in a more obvious location before the loop body. Here 
-is the definition of the ‘while’ utility function:
+states, and places the condition in a more obvious location before the loop body. This 
+utility function is defined in Figure \ref{fig:cryptolWhile}.
 
@@
   \centering
-\begin{verbatim}
+\begin{lstlisting}[language=Cryptol]
 while : {a} a -> (a -> Bit) -> (a -> a) -> a
@@
   else state
-\end{verbatim}
+\end{lstlisting}
 \caption{Cryptol While utility function}

### 2021-09-11 paper/paper.tex
Add proof figure
Add contributions to the introduction
Update paper.tex and amended.tex Add Cryptol lstlisting highlighting.

@@
 \documentclass[sigplan, 10pt, anonymous, review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\usepackage{subcaption}
+\usepackage{paralist}
+\usepackage{color}
+\definecolor{blue}{RGB}{42,0,255}
+\definecolor{green}{RGB}{63,127,95}
+\definecolor{purple}{RGB}{127,0,85}
+\definecolor{lime}{RGB}{114, 148, 0}
+\usepackage{listings}
+\lstdefinelanguage{Cryptol}{
+  keywordstyle=[1]\color{blue}\bfseries,
+  keywordstyle=[2]\color{purple}\bfseries,
+  keywordstyle=[3]\color{black}\bfseries,
+  basicstyle={\color{lime}\tiny},
+  literate=*
+  {*}{{{\color{black}$*$}}}{1}
+  {=}{{{\color{black}=}}}{1}
+  {+}{{{\color{black}+}}}{1}
+  {-}{{{\color{black}$-$}}}{1}
+  {\%}{{{\color{black}\%}}}{1}
+  {(}{{{\color{black}(}}}{1}
+  {)}{{{\color{black})}}}{1}
+  {\{}{{{\color{black}\{}}}{1}
+  {\}}{{{\color{black}\}}}}{1}
+  {[}{{{\color{black}[}}}{1}
+  {]}{{{\color{black}]}}}{1}
+  {|}{{{\color{black}$|$}}}{1}
+  {>}{{{\color{black}$>$}}}{1}
+  {<}{{{\color{black}$<$}}}{1}
+  {\#}{{{\color{black}\#}}}{1}
+  {@}{{{\color{black}@}}}{1}
+  {.}{{{\color{black}.}}}{1}
+  {\'}{{{\color{black}'}}}{1}
+  {:}{{{\color{black}:}}}{1}
+  {^}{{{\color{black}$\char`\^$}}}{1}
+  {&}{{{\color{black}\textbackslash}}}{1},
+  morekeywords=[1]{Bit, Integer, STATE_ARR, W, COL, type, LIST4, LIST63, True, False},
+  morekeywords=[2]{if, then, else, where},
+  morekeywords=[3]{pi, rc, cond, a, b, x, y, z, for, n, r, R, f, rs, i, index, t, 
+    vals, loop, state, while, take}
+}
 

### 2021-09-11 paper/results.tex
Add results.tex file

\section{Results and Discussion}\label{sec:results}

Talk about the cost of the verification. The amount of effort required to write the specification. Make clear the limits of the result proof.

* Visual inspection and limited test vectors for (1) and (2)
* Only provide proof certificates for message sizes of X to Y for digest sizes of … for (3)
* Visual inspection only for anything above Keccak for (3) in OpenSSLs C-code

Effort (see above and then consider below):
 - How long were we working on the Cryptol code?
 - How long were we working on the SAW proofs?
 - How long did it take to understand the OpenSSL code?
We'll want to determine how long it took us and then try to infer how long that would have taken had we been working full time (40 hours/week).

Because of the modular design of the SHA-3 specification, the Cryptol implementation and OpenSSL implementation are programmatically more similar. The only strong discrepancy is that the OpenSSL implemenation organizes the bits of the sponge construction. The affect of modular cryptographic design can be seen by the proof completing successfully in the Yices, Z3, and ABC SMT-solvers. Previous work on SHA2-256 was only successful with Yices \ref{cite:nfm-us}. Full results are show in table \ref{finaltable}.

\begin{table}[b]
\caption{Proof Runtimes for SHA-3 with SMT-solvers}\label{finaltable}
\setlength{\tabcolsep}{13.5pt}
\begin{tabular}{|l|r|r|r|}
\hline
\textbf{Function} & \textbf{Yices} & \textbf{Z3} & \textbf{ABC} \\
\hline
Pi          &   0.895 secs &    11.072 secs &    0.417 secs \\
Rho         &   4.509 secs &    15.776 secs &    5.879 secs \\
Theta       &  25.942 secs &    11.345 secs &   14.869 secs \\
Chi         &   5.059 secs &     3.930 secs &    1.327 secs \\
Iota        &  11.048 secs &     8.801 secs &  612.497 secs \\
KeccakF1600 & 745.206 secs & 15187.671 secs & 2910.549 secs \\
\hline
\end{tabular}
\end{table}

### 2021-09-13 paper/paper.tex
Update paper.tex and ammended.tex lstlisting.

--- 2021-09-04 paper/paper.tex
@@
 \documentclass[sigplan, 10pt, anonymous, review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\usepackage{subcaption}
+\usepackage{paralist}
+\usepackage{color}
+\definecolor{blue}{RGB}{42,0,255}
+\definecolor{green}{RGB}{63,127,95}
+\definecolor{purple}{RGB}{127,0,85}
+\definecolor{lime}{RGB}{114, 148, 0}
+\usepackage{listings}
+\lstdefinelanguage{Cryptol}{
+  keywordstyle=[1]\color{blue}\bfseries,
+  keywordstyle=[2]\color{purple}\bfseries,
+  keywordstyle=[3]\color{black}\bfseries,
+  basicstyle={\color{lime}\scriptsize},
+  literate=*
+  {*}{{{\color{black}$*$}}}{1}
+  {=}{{{\color{black}=}}}{1}
+  {+}{{{\color{black}+}}}{1}
+  {-}{{{\color{black}$-$}}}{1}
+  {\%}{{{\color{black}\%}}}{1}
+  {(}{{{\color{black}(}}}{1}
+  {)}{{{\color{black})}}}{1}
+  {\{}{{{\color{black}\{}}}{1}
+  {\}}{{{\color{black}\}}}}{1}
+  {[}{{{\color{black}[}}}{1}
+  {]}{{{\color{black}]}}}{1}
+  {|}{{{\color{black}$|$}}}{1}
+  {>}{{{\color{black}$>$}}}{1}
+  {<}{{{\color{black}$<$}}}{1}
+  {\#}{{{\color{black}\#}}}{1}
+  {@}{{{\color{black}@}}}{1}
+  {.}{{{\color{black}.}}}{1}
+  {\'}{{{\color{black}'}}}{1}
+  {:}{{{\color{black}:}}}{1}
+  {^}{{{\color{black}$\char`\^$}}}{1}
+  {&}{{{\color{black}\textbackslash}}}{1},
+  morekeywords=[1]{Bit, Integer, STATE_ARR, W, COL, type, LIST4, LIST63, True, False},
+  morekeywords=[2]{if, then, else, where},
+  morekeywords=[3]{pi, rc, cond, a, b, x, y, z, for, n, r, R, f, rs, i, index, t, 
+    vals, loop, state, while, take}
+}
 

### 2021-09-13 paper/iota.tex
Update iota.tex.

@@
 \section{Static Iota}\label{sec:iota}
-Talk about changes we made to iota to prove out directly between precomputed and computed on the fly.
+NIST specification invokes the rc method seven times for each iota function call.
+As the rc inputs are the same for each round, OpenSSL's optimized implementation of keccak stores 
+the output values as a constant table. Initially, we matched out Cryptol execution of the NIST standard 
+as closely as possible--including rc method calls in the iota function. We soon found that SAW was unable 
+to map our implementation of keccak to OpenSSL's without the use of a constant table.
+Our final Cryptol implementation uses the rc definition found in the NIST specification to generate a table 
+that is easiy compared to OpenSSL. This allows SAW to generate a proof and permits visual inspection of equivalence.

### 2021-09-14 paper/amended.tex
Update paper.tex and ammended.tex lstlisting.

@@
   \centering
-  \begin{lstlisting}[basewidth = {.5em}]
+  \begin{lstlisting}[basewidth = {.5em},basicstyle={\scriptsize}]
 1. For all triples (x,y,z) such that 
@@
 imperative languages in that it iterates and runs a loop body over defined elements. 
-It is similar to the Array.prototype.map method in JavaScript in that it takes in a 
-function and a list of arguments and returns a list of results from calling the function 
+It recieves a function and a list of arguments and returns a list of results from calling the function 
 on every element of the input array. Figure \ref{fig:cryptolFor} shows it's implementation.
@@
 for : {n, a, b} [n]a -> (a -> b) -> [n]b
-for vals loop = [loop index | index <- vals]
+for vals loop = [loop i | i <- vals]
 \end{lstlisting}
@@
   \centering
-\begin{lstlisting}[basewidth = {.5em}]
+\begin{lstlisting}[basewidth = {.5em},basicstyle={\scriptsize}]
 1. If t mod 255 = 0, return 1.
@@
   e. R[6] = R[6] ^ R[8];
-  f. R =Trunc8[R].
+  f. R = Trunc8[R].
 4. Return R[0].

### 2021-09-16 paper/preliminaries.tex
Restructure sections

@@
 \subsection{Tools}
-The Software Analysis Workbench (SAW) provides a verification environment for compositional proofs of functional equivalence between two pieces of supported code.  
-SAW uses symbolic execution to create computational models of the code and creates problems of functional equivalence between those two models that can be pushed off to an SMT solver.  
-The difficulty of these SMT problems is directly affected by the size and complexity of the computations that are being compared.  
-This can lead to drastic increases in proof times and sometimes even prevents completion as computations increase.  
-SAW’s solution to this is to create contracts (called overrides) of a functions input to output relationship which allows for a modular approach to writing the proof.  
-By breaking up computation into smaller pieces, and creating overrides for these pieces, SAW can often simplify an otherwise complex (and thus incompletable) proof to remain within the capacity of an SMT solver.
-
-CRYPTOL!!!!!!
-
-\subsection{Trusted Code Base}
-This work relies on the correctness of several tools, including language compilers.  
-Our Trusted Code Base consists of:
-\begin{itemize}
-  \item LLVM bitcode compiler
-  \item The Software Analysis Workbench (SAW) including the conversion from LLVM bitcode to SAWCore terms, the Cryptol compiler to SAWCore terms, and SAW's interface with SMT solvers.
-  \item The z3, abc, and yices SMT solvers.
-\end{itemize}

### 2021-09-16 paper/proof.tex
update abstract and figure
Prep abstract for submission
Restructure sections

@@
-\section{Proof Structure}\label{sec:proof}
-Talk about our proof flow
-\subsection{Modular Approach}
-Talk about how NIST standard/openssl being modular simplified our proof structure.
+\section{Proof Approach}\label{sec:proof}
+
+The Software Analysis Workbench (SAW) provides a verification environment for compositional proofs of functional equivalence between two pieces of supported code.  
+SAW uses symbolic execution to create computational models of the code and creates problems of functional equivalence between those two models that can be pushed off to an SMT solver.  
+The difficulty of these SMT problems is directly affected by the size and complexity of the computations that are being compared.  
+This can lead to drastic increases in proof times and sometimes even prevents completion as computations increase.  
+SAW’s solution to this is to create contracts (called overrides) of a functions input to output relationship which allows for a modular approach to writing the proof.  
+By breaking up computation into smaller pieces, and creating overrides for these pieces, SAW can often simplify an otherwise complex (and thus incompletable) proof to remain within the capacity of an SMT solver.
+
+CRYPTOL!!!!!!
+
+This work relies on the correctness of several tools, including language compilers.  
+Our Trusted Code Base consists of:
+\begin{itemize}
+  \item LLVM bitcode compiler
+  \item The Software Analysis Workbench (SAW) including the conversion from LLVM bitcode to SAWCore terms, the Cryptol compiler to SAWCore terms, and SAW's interface with SMT solvers.
+  \item The z3, abc, and yices SMT solvers.
+\end{itemize}

### 2021-09-16 paper/proof.tex
Add start of proof approach

@@
 
-The Software Analysis Workbench (SAW) provides a verification environment for compositional proofs of functional equivalence between two pieces of supported code.  
+The Software Analysis Workbench (SAW) provides a verification environment for compositional proofs of functional equivalence between two pieces of supported code (C, Java, or Cryptol).  
 SAW uses symbolic execution to create computational models of the code and creates problems of functional equivalence between those two models that can be pushed off to an SMT solver.  
@@
 
-CRYPTOL!!!!!!
+Cryptol is a domain-specific specification language, which has been used in the previously mentioned works by Galois' and by Decker et.
+al to write visually reviewable implementations of NIST standards.
+It has strong ties to the SAW tool and, in fact, can be compiled directly down to the SAW Terms (computational models) which are used by SAW in proving functional equivalence.
+Because of the interwoven nature of Cryptol with SAW, the Terms created from Cryptol are in most cases significantly simpler than any created through symbolic execution of C/Java code.
+
+The published NIST standards provide no compileable implementation that can be used as a starting point for proving out the algorithm.
+As a result, any proof of correctness with a NIST standard is limited in strength to how simple it is use visual review to connect code used in the proof to the NIST standard, coupled with a small sampling of test vectors.
+As shown in \ref{proofStructure}, The proof flow starts with a specification of the SHA3 algorithm in Cryptol with the goal of being as similar visually to the NIST standard as possible.
+This work discusses significant ways that Cryptol's functional nature can be leveraged to aid in this goal.
+Along with simplifying the visual inspection process, this Cryptol specification also passes all test vectors provided by NIST.
+
+The end goal of this proof is to show functional equivalence between this visually reviewable specifications Keccak function, and that of OpenSSL's C code.
+Unfortunately, a direct proof between this specification, and OpenSSL's implementation proved unfeasible, because OpenSSL varies from the NIST standard in how it stores the bits of the intermediate state array in memory.
+These minor modifications in memory storage put such a proof out of SAW's capabilities, because it relies on strict typing, and comparison of a Cryptol functions input/output to a C functions memory state.
+In order to bridge that gap, a second Cryptol specification is included in the proof flow.
+It matches in every way the first Cryptol code, except when it comes to how the state array is stored and accessed (at which point it follows OpenSSL's form).
+Because of the efficient nature of Cryptol's compilation to SAW Terms, proving equivalence between these Cryptol implementations was possible at the top level of the SHA3 algorithm (which allows for the differences in state array storage to be fully abstracted out).
+Due to the strict typing nature required for writing SAW proofs, the scope of this equivalence proof is limited to a given input message size.
+The only computations that change in the SHA3 algorithm due to message size are in the padding and sponging of a block.
+With this in mind, an exhaustive proof run over each input message byte count from 1 to the block size is sufficient to cover the differences in computation.
+The artifact repository accompanying this work contains the receipts of this exhaustive proof. (REFERENCE TO PROOF!!!!)
+
+INNER GREEK FUNCTIONS:
+
+CONTRACTS CAN BE USED TO OVERRIDE AND REPLACE C WITH Cryptol
+
+WITH THESE CONTRACTS SAW CAN COMPLETE A PROOF OF EQUIVALENCE BETWEEN TWO PIECES OF CODE.
 
 This work relies on the correctness of several tools, including language compilers.
-Our Trusted Code Base consists of:
+
+\noindent \textbf{Trusted Code Base}:
 \begin{itemize}

### 2021-09-16 paper/abstract.tex
Fix the issue with tic and apostrophe

@@
 These constructs form the basis for modern secure hashing.
-This work extends and improves the methodology used to verify Amazon’s s2n HMAC and OpenSSL’s SHA2-256 inner hash function. 
-It uses the Software Analysis Workbench (SAW) to prove out functional equivalence between OpenSSL’s SHA-3 Keccak function and a verified implementation of SHA-3 in Cryptol, a domain-specific language for cryptography.
+This work extends and improves the methodology used to verify Amazon's s2n HMAC and OpenSSL's SHA2-256 inner hash function. 
+It uses the Software Analysis Workbench (SAW) to prove out functional equivalence between OpenSSL's SHA-3 Keccak function and a verified implementation of SHA-3 in Cryptol, a domain-specific language for cryptography.
 SAW allows for a modular approach to writing proofs through the use of contracts, which define a functions input to output relationship.

### 2021-09-16 paper/introduction.tex
Fix the issue with tic and apostrophe

@@
 \section{Introduction}\label{sec:introduction}
-  The Secure Hashing Algorithm (SHA) family of cryptography is widely used as a means of providing increased security through generating/verifying digital signatures, deriving keys, and generating pseudorandom bits.
-(NIST: https://csrc.nist.gov/publications/detail/fips/202/final)
+  The Secure Hashing Algorithm (SHA) family of cryptography is widely used as a means of providing increased security through generating/verifying digital signatures, deriving keys, and generating pseudorandom bits (\href{https://csrc.nist.gov/publications/detail/fips/202/final}{NIST specification}).
   The structure and computations of these algorithms is defined by NIST in mathematical specifications.  
@@
   \item Visual inspection and limited test vectors for (1) and (2)
-  \item Only provide proof certificates for message sizes of X to Y for digest sizes of … for (3)
-  \item Visual inspection only for anything above Keccak for (3) in OpenSSLs C-code
+  \item Only provide proof certificates for message sizes of X to Y for digest sizes of \textbf{FIXME} for (3)
+  \item Visual inspection only for anything above Keccak for (3) in OpenSSL's C-code
 \end{compactitem}
@@
 \subsection{Related Work}
-  Previously, Galois’ Inc. worked with Amazon to prove out parts of Amazon’s s2n HMAC code to the NIST specification.  
+  Previously, Galois' Inc. worked with Amazon to prove out parts of Amazon's s2n HMAC code to the NIST specification.  
 In their work they stubbed out use of the SHA2 algorithm because their HMAC allows for customization of the algorithm type and implementation.
@@
 The work by Decker et. al, extended on Galois' work by choosing a single SHA2 implementation to prove out.  
-They chose OpenSSL’s SHA 256 implementation, as that is a commonly used library (and one that is usable in s2n).  
+They chose OpenSSL's SHA 256 implementation, as that is a commonly used library (and one that is usable in s2n).  
 They were able to prove out the inner hash computations of that implementation using a methodology of compositional proofs to do so.  
-They were also able to connect up their proof with the code used by Galois’ to stub out the SHA2 algorithm, which implies that the functional equivalence property established with the OpenSSL code applies to Galois' Cryptol implementation as well.
+They were also able to connect up their proof with the code used by Galois' to stub out the SHA2 algorithm, which implies that the functional equivalence property established with the OpenSSL code applies to Galois' Cryptol implementation as well.
 Their work showed the viability and limitations of creating proofs of functional equivalence for legacy cryptographic implementations.

### 2021-09-16 paper/fips.tex
Restructure sections
Fix more LaTeX apostrophe issues

--- 2021-09-14 paper/amended.tex
@@
-\section{Amended Cryptol}
-\label{sec:amended}
-Initially it appeared that visual inspection with Cryptol’s standard syntax was subpar.
+\section{SHA3 FIPS Specification in Cryptol}
+\label{sec:fips}
+Initially it appeared that visual inspection with Cryptol's standard syntax was subpar.
 The language's functional folds and list comprehensions do not align easily with the imperative 
-and array-based NIST standard. By leveraging Cryptol’s versatility, however, we produced a couple 
+and array-based NIST standard. By leveraging Cryptol's versatility, however, we produced a couple 
 utility functions to visually mimic the sequential nature of the NIST specification.
@@
 \subsection{For Utility Function}
-The phrase “for all” indicates a looping structure and is quite common in the NIST standard. 
+The phrase \emph{for all} indicates a looping structure and is quite common in the NIST standard. 
 It is used in the definition of the pi function shown in Figure \ref{fig:nistPi}.
@@
 Cryptol declares list comprehension variables after they are used, and the order of the
-nested loops is not as apparent. Using a helper function, ‘for’, we were able to provide
+nested loops is not as apparent. Using a helper function, \emph{for}, we were able to provide
 a more clear definition seen in Figure \ref{fig:cryptolAmendedPi}.
@@
 
-The ‘for’ utility function follows the same intuition as the implementation in most 
+The \emph{for} utility function follows the same intuition as the implementation in most 
 imperative languages in that it iterates and runs a loop body over defined elements. 
@@
 \subsection{While Utility Function}
-Another common phrasing is the “let...for...let”. This defines a state upon which a folding 
+Another common phrasing is the \emph{let...for...let}. This defines a state upon which a folding 
 occurs, iteratively condensing the information into a final state. Imperatively, it is most 
@@
 Maintaining this state through Cryptol's list comprehensions can be difficult.
-The comprehensions require a list to iterate through.Since our ‘for’ utility 
+The comprehensions require a list to iterate through.Since our \emph{for} utility 
 function is simply a wrapper for list comprehensions, it is equally unsuited for 
@@
 it also awkwardly indexes into a list of previous states. This differs vastly from the 
-NIST specificiation both visually and functionally. Using a new ‘while’ utility function, we 
+NIST specificiation both visually and functionally. Using a new \emph{while} utility function, we 
 can declare an initial state, condition function, and body function that will recursively 
@@
 
-The final state is returned by the ‘while’ function. This iterates only until the 
+The final state is returned by the \emph{while} function. This iterates only until the 
 condition is met, avoids the clutter of indexing within a larger array of all previous 

### 2021-09-16 paper/proof.tex
Fix more LaTeX apostrophe issues

@@
 
-The Software Analysis Workbench (SAW) provides a verification environment for compositional proofs of functional equivalence between two pieces of supported code (C, Java, or Cryptol).  
+SAW provides a verification environment for compositional proofs of functional equivalence between two pieces of supported code (C, Java, or Cryptol).  
 SAW uses symbolic execution to create computational models of the code and creates problems of functional equivalence between those two models that can be pushed off to an SMT solver.  
@@
 This can lead to drastic increases in proof times and sometimes even prevents completion as computations increase.  
-SAW’s solution to this is to create contracts (called overrides) of a functions input to output relationship which allows for a modular approach to writing the proof.  
+SAW's solution to this is to create contracts (called overrides) of a functions input to output relationship which allows for a modular approach to writing the proof.  
 By breaking up computation into smaller pieces, and creating overrides for these pieces, SAW can often simplify an otherwise complex (and thus incompletable) proof to remain within the capacity of an SMT solver.

### 2021-09-16 paper/results.tex
Restructure sections
Fix more LaTeX apostrophe issues

@@
 
-* Visual inspection and limited test vectors for (1) and (2)
-* Only provide proof certificates for message sizes of X to Y for digest sizes of … for (3)
-* Visual inspection only for anything above Keccak for (3) in OpenSSLs C-code
+\begin{compactitem}
+\item Visual inspection and limited test vectors for (1) and (2)
+\item Only provide proof certificates for message sizes of X to Y for digest sizes of \textbf{FIXME} for (3)
+\item Visual inspection only for anything above Keccak for (3) in OpenSSLs C-code
+\end{compactitem}
 
 Effort (see above and then consider below):
- - How long were we working on the Cryptol code?
- - How long were we working on the SAW proofs?
- - How long did it take to understand the OpenSSL code?
+\begin{compactitem}
+ \item How long were we working on the Cryptol code?
+ \item How long were we working on the SAW proofs?
+ \item How long did it take to understand the OpenSSL code?
+\end{compactitem}
 We'll want to determine how long it took us and then try to infer how long that would have taken had we been working full time (40 hours/week).
@@
 Because of the modular design of the SHA-3 specification, the Cryptol implementation and OpenSSL implementation are programmatically more similar. 
-The only strong discrepancy is that the OpenSSL implemenation organizes the bits of the sponge construction.
+The only strong discrepancy is that the OpenSSL implementation organizes the bits of the sponge construction. 
 The affect of modular cryptographic design can be seen by the proof completing successfully in the Yices, Z3, and ABC SMT-solvers. 
-Previous work on SHA2-256 was only successful with Yices \ref{cite:nfm-us}.
-Full results are show in table \ref{finaltable}.
+Previous work on SHA2-256 was only successful with Yices \ref{cite:nfm-us}. Full results are show in table \ref{finaltable}.
 
-\begin{table}[b]
+\begin{table*}[b]
 \caption{Proof Runtimes for SHA-3 with SMT-solvers}\label{finaltable}
@@
 \end{tabular}
-\end{table}
+\end{table*}

### 2021-09-16 paper/fips.tex
Update fips.tex and openssl.tex to comform to the rest of the paper's style

@@
 Initially it appeared that visual inspection with Cryptol's standard syntax was subpar.
-The language's functional folds and list comprehensions do not align easily with the imperative 
-and array-based NIST standard. By leveraging Cryptol's versatility, however, we produced a couple 
-utility functions to visually mimic the sequential nature of the NIST specification.
+The language's functional folds and list comprehensions do not align easily with the imperative and array-based NIST standard. 
+By leveraging Cryptol's versatility, utility functions can visually mimic the sequential nature of the NIST specification.
 
@@
 
-By nesting Cryptol's list comprehensions, we easily implemented the pi function in Figure 
-\ref{fig:cryptolPi}.
+By nesting Cryptol's list comprehensions, the pi function can be easily implemented.
+This is illustrated in Figure \ref{fig:cryptolPi}.
 
@@
 
-Cryptol declares list comprehension variables after they are used, and the order of the
-nested loops is not as apparent. Using a helper function, \emph{for}, we were able to provide
-a more clear definition seen in Figure \ref{fig:cryptolAmendedPi}.
+Cryptol declares list comprehension variables after they are used, and the order of the nested loops is not as apparent. 
+Using a helper function, \emph{for}, a clearer definition seen in Figure \ref{fig:cryptolAmendedPi}.
 
@@
 
-The \emph{for} utility function follows the same intuition as the implementation in most 
-imperative languages in that it iterates and runs a loop body over defined elements. 
-It recieves a function and a list of arguments and returns a list of results from calling the function 
-on every element of the input array. Figure \ref{fig:cryptolFor} shows it's implementation.
+The \emph{for} utility function follows the same intuition as the implementation in most imperative languages in that it iterates and runs a loop body over defined elements. 
+It recieves a function and a list of arguments and returns a list of results from calling the function on every element of the input array. 
+Figure \ref{fig:cryptolFor} shows it's implementation.
 
@@
 Though it is simple to implement, it reorders and structures the arguments in a very elegant way.
-In constrast with standard cryptol syntax, the list to iterate over and the iterating 
-variable are visible before the loop body. List comprehensions are hidden away behind this 
-function call and visual equivalence is clearer.
+In constrast with standard cryptol syntax, the list to iterate over and the iterating variable are visible before the loop body. 
+List comprehensions are hidden away behind this function call and visual equivalence is clearer.
 
 \subsection{While Utility Function}
-Another common phrasing is the \emph{let...for...let}. This defines a state upon which a folding 
-occurs, iteratively condensing the information into a final state. Imperatively, it is most 
-comparable to a while loop. As shown by Figure \ref{fig:nistRC}, is present in the rc function 
-defined by the NIST specification.
+Another common phrasing is the \emph{let...for...let}. 
+This defines a state upon which a folding occurs, iteratively condensing the information into a final state. 
+Imperatively, it is most comparable to a while loop. 
+As shown by Figure \ref{fig:nistRC}, is present in the rc function defined by the NIST specification.
 
@@
 Maintaining this state through Cryptol's list comprehensions can be difficult.
-The comprehensions require a list to iterate through.Since our \emph{for} utility 
-function is simply a wrapper for list comprehensions, it is equally unsuited for 
-visual verification on such steps. The condition to stop iterating in the RC function
-bounds the number of iterations by [1, 254], which lets us construct such a list to 
-iterate through. This leads to a rather ungainly Cryptol definition, however, as 
-shown by Figure \ref{fig:cryptolRC}.
+The comprehensions require a list to iterate through. 
+Since the \emph{for} utility function is simply a wrapper for list comprehensions, it is equally unsuited for visual verification on such steps. 
+The number of iterations in the RC function is bounded by [1, 254] by its modular condition.
+This range provides the list to iterate over as required by Cryptol list comprehensions. 
+From these observations, a rather ungainly Cryptol definition results.
+This definition is shown in Figure \ref{fig:cryptolRC}.
 
@@
 
-Not only is this definition computationally slower as it runs 255 loops every time it is called, 
-it also awkwardly indexes into a list of previous states. This differs vastly from the 
-NIST specificiation both visually and functionally. Using a new \emph{while} utility function, we 
-can declare an initial state, condition function, and body function that will recursively 
-be called until the condition is met. It consolodates the information of the previous state 
-until the condition is sastified. This definition is foudn in Figure \ref{fig:cryptolamendedRC}.
+Not only is this definition computationally slower as it runs 255 loops every time it is called, it also awkwardly indexes into a list of previous states. 
+This differs vastly from the NIST specificiation both visually and functionally. 
+The \emph{while} utility function, accepts an initial state, condition function, and body function to recursively loop. 
+It consolodates the information of the previous state until the condition is sastified. 
+This definition is found in Figure \ref{fig:cryptolamendedRC}.
 
@@
 
-The final state is returned by the \emph{while} function. This iterates only until the 
-condition is met, avoids the clutter of indexing within a larger array of all previous 
-states, and places the condition in a more obvious location before the loop body. This 
-utility function is defined in Figure \ref{fig:cryptolWhile}.
+The final state is returned by the \emph{while} function. 
+This iterates only until the condition is met, avoids the clutter of indexing within a larger array of all previous states, and places the condition in a more obvious location before the loop body. 
+This utility function is defined in Figure \ref{fig:cryptolWhile}.
 

### 2021-09-16 paper/openssl.tex
Restructure sections
Update fips.tex and openssl.tex to comform to the rest of the paper's style

--- 2021-09-04 paper/iota.tex
@@
-\section{Static Iota}\label{sec:iota}
-Talk about changes we made to iota to prove out directly between precomputed and computed on the fly.
+\section{SHA3 OpenSSL in Cryptol}\label{sec:openssl}
+The NIST specification invokes the rc method seven times for each iota function call.
+As the rc inputs are the same for each round, OpenSSL's optimized implementation of keccak stores the output values as a constant table rather than repeatedly compute them. 
+Initially, the Cryptol code to the NIST standard as closely as possible\textemdash including RC method calls in the iota function. 
+SAW was unable to map such an implementation of keccak to OpenSSL's implementation without the use of a constant table. 
+The final Cryptol implementation uses the RC definition found in the NIST specification to generate a table that is easiy compared to OpenSSL. 
+This allows SAW to generate a proof and permits visual inspection of equivalence.

### 2021-09-20 paper/openssl.tex
Add Iota section

--- 2021-09-04 paper/iota.tex
@@
-\section{Static Iota}\label{sec:iota}
-Talk about changes we made to iota to prove out directly between precomputed and computed on the fly.
+\section{SHA3 OpenSSL in Cryptol}\label{sec:openssl}
+Even with the highly modular approach of the SHA3 algorithm, a direct proof from the Cryptol specification that matched the NIST standard and the OpenSSL source code was unviable.
+OpenSSL's source code is often used in low-level software, and it is often optimized to improve performance and viability for embedded systems.
+One major instance of such optimization proved significant enough to prevent a direct proof from the NIST Cryptol specification to OpenSSL.
+
+\lstset{style=customc, firstnumber=177}
+\begin{figure}[h]
+  \centering
+\begin{lstlisting}
+  void Iota(uint64_t A[5][5], size_t i)
+  {
+    //assert(i < (sizeof(iotas) / sizeof(iotas[0])));
+    if (i < (sizeof(iotas) / sizeof(iotas[0]))) {
+      A[0][0] ^= iotas[i];
+    }
+  }
+\end{lstlisting}
+\caption{OpenSSL's Iota Function}
+\label{fig:cIota}
+\end{figure}
+
+As per the NIST specification, the Iota function invokes the rc algorithm to generate a constant value specific to the round count to use in exclusive-or operations.
+The rc function inputs are unrelated to the message to be hashed, which means that for any given message, the iteration through the rounds of keccak will produce the same constant values.
+Because these values are unchanged from run to run, OpenSSL's optimized implementation of keccak stores the output values as a constant table and use the current round number to properly index into it as is shown in Figure \ref{fig:cIota}.
+Initially, the Cryptol code followed the NIST standard as closely as possible\textemdash including rc method calls in the Iota function which computed the values every iteration.
+Attempts to prove out such a specification to the OpenSSL code failed however, because SAW was unable to show the computational equivalence between computing values as needed and looking up values in a constant table.
+
+\begin{figure}[h]
+  \centering
+\begin{lstlisting}[language=Cryptol]
+  iota : STATE_ARR -> [64] -> STATE_ARR
+  iota a i = for LIST4 (\y ->
+                for LIST4 (\x ->
+                    for LIST63 (\z ->
+                        if ((x == 0) && (y == 0))
+                            then (a @0 @0 @z) ^ (LISTIOTAS @i @z)
+	                        else a @y @x @z)))
+
+  LISTIOTAS = [reverse (
+        (while {RC = 0:[W], j = 0} //STATE
+            (\state -> state.j <= `L) //COND
+            (\state -> { //LOOP
+                RC = for LIST63 (\z ->
+                    if z == index
+                        then rc (state.j + 7*i)
+                        else state.RC @z)
+                    where index = ((1:[8]) << state.j) - 1,
+                j = state.j + 1})
+        ).RC) | i <- LISTROUNDS:[_][64]]
+
+  rc : [64] -> Bit
+  rc t = if (t % 255) == 0 
+	  then True 
+	  else (while {i = 1, R = 128:[8]} //STATE
+		  (\state -> state.i <= t % 255) //COND
+		  (\state -> { //LOOP
+			  i = state.i + 1, 
+			  R = (take`{8} ([(r@0)^(r@8), r@1, r@2, r@3, (r@4)^(r@8), (r@5)^(r@8), (r@6)^(r@8), r@7, r@8]) 
+				  where r = [False] # state.R)
+		    })
+		).R @0
+\end{lstlisting}
+\caption{OpenSSL's Iota Function}
+\label{fig:cryptolIota}
+\end{figure}
+
+The lack of any meaningful trace of computation that comes with pre-computed values, meant that a different approach would be necessary to meet in the middle.
+The solution was to modify the Cryptol specification to also use a lookup table in the Iota function.
+However, the Cryptol code still uses the rc function to generate that lookup table at the beginning of every run of SHA3, rather than strictly stating constant values in the code.
+By having both the Cryptol and C Iota functions now rely on looking up values in a table, the task of proving them became far simpler.
+SAW only had to show that the table generated by running the rc function all at once at the beginning, was identical to the one hard coded into OpenSSL's implementation.
+Because those match, and both implementations index into the table in the same way, SAW was able to generate the proof.
+
+This approach recognizably added an extra element to the visual inspection between the Cryptol specification and the C code.
+The full rc computation is still used in the Cryptol code, with the only significant difference being that rather than run rc every time Iota is called, rc is run all at once to generate all the needed values.
+The continued inclusion of the rc computation makes it only a much less significant inconvenience to accept the computed table as accurate. 
+It likewise remains simple to see visually that indexing into the table gives back the proper constant value for a given round count.
+
+As a note, one minor adjustment was also made to OpenSSL's Iota function code to allow the proof to complete properly.
+As an error check, OpenSSL uses an assert to ensure that the round count is never greater than it should be (to ensure no indexing out of the array's bounds).
+SAW cannot recognize an assert statement in its symbolic execution engine.
+As a result, that statement was removed and replaced with a comparable if statement which maintained the logic of the function.
+Because the equivalence proof for the Iota function is run with the precondition that the round count must be less than 24 (max rounds), this does not affect the comparison between the C code and the Cryptol code (which does not have that error check).

### 2021-09-20 paper/paper.tex
Prep abstract for submission
Restructure sections
Fix errors in bib file
Add Iota section

@@
     vals, loop, state, while, take}
+}
+
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{almond}{rgb}{0.94, 0.87, 0.8}
+\definecolor{amaranth}{rgb}{0.9, 0.17, 0.31}
+\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
+\definecolor{gray}{rgb}{0.5,0.5,0.5}
+
+\lstdefinestyle{customc}{
+  frame=none,
+  language=C,
+  basicstyle=\footnotesize\ttfamily, 
+  aboveskip=3mm,
+  belowskip=3mm,
+  showstringspaces=false,
+  numbers=left,
+  keywordstyle=\bfseries\color{azure},
+  commentstyle=\itshape\color{dkgreen},
+  identifierstyle=\color{black},
+  stringstyle=\color{almond},
+  numbersep=-1pt,
+  numberstyle=\color{gray},
+  breaklines=true,
+  breakatwhitespace=true,
+  tabsize=3,
 }
@@
 \input{introduction}
-\input{preliminaries}
+%%\input{preliminaries}
 \input{proof}
-\input{amended}
-\input{iota}
-\input{visual}
+\input{fips}
+\input{openssl}
+\input{results}
+\input{related}
 \input{conclusion}
@@
 \begin{acks}
-To Robert, for the bagels and explaining CMYK and color spaces.
+
 \end{acks}
@@
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{paper.bib}
+\bibliography{paper}
 

### 2021-09-27 paper/openssl.tex
Finish OpenSSL section

@@
         ).RC) | i <- LISTROUNDS:[_][64]]
-
-  rc : [64] -> Bit
-  rc t = if (t % 255) == 0 
-	  then True 
-	  else (while {i = 1, R = 128:[8]} //STATE
-		  (\state -> state.i <= t % 255) //COND
-		  (\state -> { //LOOP
-			  i = state.i + 1, 
-			  R = (take`{8} ([(r@0)^(r@8), r@1, r@2, r@3, (r@4)^(r@8), (r@5)^(r@8), (r@6)^(r@8), r@7, r@8]) 
-				  where r = [False] # state.R)
-		    })
-		).R @0
 \end{lstlisting}
@@
 The solution was to modify the Cryptol specification to also use a lookup table in the Iota function.
-However, the Cryptol code still uses the rc function to generate that lookup table at the beginning of every run of SHA3, rather than strictly stating constant values in the code.
+However, the Cryptol code still uses the rc function as defined in Figure \ref{fig:nistRC} to generate that lookup table at the beginning of every run of SHA3, rather than strictly stating constant values in the code.
 By having both the Cryptol and C Iota functions now rely on looking up values in a table, the task of proving them became far simpler.
@@
 Because the equivalence proof for the Iota function is run with the precondition that the round count must be less than 24 (max rounds), this does not affect the comparison between the C code and the Cryptol code (which does not have that error check).
+
+A second obstacle that blocked a single direct proof is how OpenSSL stores and accesses its state array in memory.  
+The NIST specification outlines the state array as being a 5 by 5 by w 3-dimensional array.
+It denotes x as the horizontal axis, y as the vertical axis, and w as the z axis, like a cartesian coordinate system.
+Because it structures the array this way, any access into the array \texttt{A[x, y, z]} is asking for the element at the x'th column, y'th row, z'th bit.
+OpenSSL defines their state array as a \texttt{uint64\textunderscore t A[5][5]}.
+As is common practice in many programming styles, their array indexes by row and then by column.
+Because of this choice, an index defined by NIST as \texttt{A[x, y, z]} would in OpenSSL's array be indexed as \texttt{bit z of A[y][x]}.
+In comparing C code to Cryptol, SAW relies on the pre and post state of memory for a given function to help determine functional equivalence.
+The difference in memory storage/access choices by NIST and OpenSSL means that SAW cannot show functional equivalence using memory comparison.
+
+OpenSSL's choice in memory storage yields another obstacle because of its array composition.
+The way that bits are ordered and stored in a \texttt{uint64\textunderscore t} in C is strictly defined by the language's semantics.
+Unfortunately, that bit ordering is reversed to the ordering NIST defines.
+This yields yet another way that the memory differs and cannot be directly compared.
+
+Between these two significant deviations by the OpenSSL implementation from the NIST standard, an intermediate step was necessary to complete a proof of equivalence.
+The solution is a second Cryptol implementation that matches OpenSSL's code in these memory differences.
+That second Cryptol code can be directly proven as equivalent to OpenSSL's code.
+
+This leaves only the step of proving the two Cryptol specifications as equivalent.
+Because of memory differences, this proof cannot be done on a low function level, as the OpenSSL proof is.
+Instead, it must be done at the top level of the SHA algorithm so that the memory compared (input message and output hash) is identical in structure, notwithstanding the difference in state array implementation.
+As SAW performs equivalence proofs based on defined static types, this proof can only be run on a specific input message size, and a specific output hash size.
+To attain full coverage this proof must be run on all input message sizes that affect the way the code is executed.
+This work determines the input set for full coverage to be all byte counts up to a full single block (200 Bytes), as this covers all padding and sponging operations performed in the creation of the state array.
+
+This approach required significant computation time and resources and cannot be easily replicated, however the artifact repository coupled with this work provides receipts of the output generated by this proof.
+However, with this step included a proof of functional equivalence between the visually reviewed Cryptol specification, and OpenSSL's Keccak function source code is complete.

### 2021-11-22 paper/proof.tex
Add outline

@@
-\section{Proof Approach}\label{sec:proof}
+SHA 3 Details
+\begin{compactitem}
+  \item Process 1600 bits at a time: $5 * 5 * 64$
+  \item State vector is $5*5*64*$
+  \item Incoming message padded out to whole number of 1600 bit bit-strings
+  \item Processed in order by \texttt{KECCAK}
+  \item Keccak turns each block into a state vector (rearranges bit layout in memory)
+  \item Keccak has 24 rounds for state vector. 
+  \item There are 5 hash functions sequentially applied in each round to iteratively update the state vector
+  \item After the last round, keccak turns the state vector back into the bit-string./\item The next unprocessed block is \emph{sponged} (e.g., absorbed) into the current bit-string from the latest keccak and the process is repeated
+  \item The max digest (hash) size is 512 bits in the standard
+  \item The final bit-string from keccak after sponging and processing every block in the original message is truncated to the desired digest size 
+\end{compactitem}
+
+SAW Details
+\begin{compactitem}
+  \item Cryptol is a specification language specifically for defining crypto-primitives
+  \item It has a compiler to turn Cryptol to SAW core terms
+  \item It has an execution environment to run test vectors on specifications
+  \item SAW is a tool to prove equivalence between SAW core terms
+  \item It uses symbolic execution and must unroll every loop to a fixed length
+  \item It can take as input C and Java code and compile those to core terms
+  \item It can use overides to replace C and Java code in the term
+\end{compactitem}
 

### 2021-11-22 paper/introduction.tex
Revise to use LLNCS formating for NFM Submission

@@
-\section{Introduction}\label{sec:introduction}
   The Secure Hashing Algorithm (SHA) family of cryptography is widely used as a means of providing increased security through generating/verifying digital signatures, deriving keys, and generating pseudorandom bits (\href{https://csrc.nist.gov/publications/detail/fips/202/final}{NIST specification}).

### 2021-11-22 paper/openssl.tex
Revise to use LLNCS formating for NFM Submission

@@
-\section{SHA3 OpenSSL in Cryptol}\label{sec:openssl}
 Even with the highly modular approach of the SHA3 algorithm, a direct proof from the Cryptol specification that matched the NIST standard and the OpenSSL source code was unviable.

### 2021-11-22 paper/paper.tex
Add outline
Revise to use LLNCS formating for NFM Submission

@@
-%%
-%% This is file `sample-sigplan.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigplan')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigplan.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%%
-%% Commands for TeXCount
-%TC:macro \cite [option:text,text]
-%TC:macro \citep [option:text,text]
-%TC:macro \citet [option:text,text]
-%TC:envir table 0 1
-%TC:envir table* 0 1
-%TC:envir tabular [ignore] word
-%TC:envir displaymath 0 word
-%TC:envir math 0 word
-%TC:envir comment 0 0
-%%
-%%
-%% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigplan, 10pt, anonymous, review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-\usepackage{subcaption}
+\documentclass[runningheads]{llncs}
+%
 \usepackage{paralist}
+\usepackage{hyperref}
+\usepackage{graphicx}
 \usepackage{color}
@@
 
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
+%
+% If you use the hyperref package, please uncomment the following line
+% to display URLs in blue roman font according to Springer's eBook style:
+\renewcommand\UrlFont{\color{blue}\rmfamily}
 
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2018}
-\acmYear{2018}
-\acmDOI{10.1145/1122445.1122456}
+\begin{document}
+%
+\title{Verification of OpenSSL's SHA3 Keccak function using the Software Analysis Workbench}
+%
+%\titlerunning{Abbreviated paper title}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
+%
+\author{
+  Parker Hanson\inst{1} \and
+  Benjamin Winters\inst{1} \and
+  Eric Mercer\inst{1}\orcidID{0000--0002-2264-2958} \and
+  Bret Decker\inst{1}
+}
+%
+\authorrunning{P. Hanson et al.}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\institute{
+  Brigham Young University, Provo UT 84602, USA
+}
+%
+\maketitle              % typeset the header of the contribution
+%
+\begin{abstract}
+\input{abstract}
+\keywords{verification \and cryptography \and symbolic execution \and SHA3.}
+\end{abstract}
 
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
-  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
-\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-  June 03--05, 2018, Woodstock, NY}
-\acmPrice{15.00}
-\acmISBN{978-1-4503-XXXX-X/18/06}
+\section{Introduction}\label{sec:introduction}
+\input{introduction}
 
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
+%%\input{preliminaries}
 
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
+\section{Proof Overview and Results}\label{sec:proof}
+\input{proof}
 
-%%
-%% end of the preamble, start of the body of the document source.
-\begin{document}
+\section{SHA3 FIPS Specification in Cryptol}\label{sec:fips}
+\input{fips}
 
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
-\title{Verification of OpenSSL's SHA3 Keccak function using the Software Analysis Workbench}
+\section{SHA3 OpenSSL in Cryptol}\label{sec:openssl}
+\input{openssl}
 
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-\author{Benjamin Winters}
-\author{Parker Hanson}
-\author{Eric Mercer}
-\email{egm@cs.byu.edu}
-\orcid{0000-0002-2264-2958}
-\author{Brett Decker}
-\email{bdecker@cs.byu.edu}
-\affiliation{%
-  \institution{Brigham Young University}
-  \streetaddress{3365 TMCB}
-  \city{Provo}
-  \state{Utah}
-  \country{USA}
-  \postcode{84602}
-}
+\section{Results and Discussion}\label{sec:results}
+\input{results}
 
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
+\section{Related Work}\label{sec:related}
+\input{related}
 
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
-
-\input{abstract}
-
-%%NEED TO COMPLETE THIS BEFORE SUBMISSION
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-\begin{CCSXML}
-<ccs2012>
- <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-  <concept_significance>500</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-\keywords{verification, cryptography, symbolic execution, SHA3}
-
-%% A "teaser" image appears between the author and affiliation
-%% information and the body of the document, and typically spans the
-%% page.
-% \begin{teaserfigure}
-%   \includegraphics[width=\textwidth]{sampleteaser}
-%   \caption{Seattle Mariners at Spring Training, 2010.}
-%   \Description{Enjoying the baseball game from the third-base
-%   seats. Ichiro Suzuki preparing to bat.}
-%   \label{fig:teaser}
-% \end{teaserfigure}
-
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
-\maketitle
-
-\input{introduction}
-%%\input{preliminaries}
-\input{proof}
-\input{fips}
-\input{openssl}
-\input{results}
-\input{related}
+\section{Conclusion}\label{sec:conclusion}
 \input{conclusion}
 
-%%
-%% The acknowledgments section is defined using the "acks" environment
-%% (and NOT an unnumbered section). This ensures the proper
-%% identification of the section in the article metadata, and the
-%% consistent spelling of the heading.
-\begin{acks}
-
-\end{acks}
-
-%%
-%% The next two lines define the bibliography style to be used, and
-%% the bibliography file.
-\bibliographystyle{ACM-Reference-Format}
+%
+%
+%
+%
+% ---- Bibliography ----
+%
+% BibTeX users should specify bibliography style 'splncs04'.
+% References will then be sorted and formatted in the correct style.
+%
+\bibliographystyle{splncs04}
 \bibliography{paper}
-
+%
 \end{document}
-\endinput
-%%
-%% End of file `paper.tex'.

### 2021-11-22 paper/paper.tex
Move imp to figs and move figs to proof section
Restructure sections and file names

@@
 \usepackage{paralist}
+\usepackage{comment}
 \usepackage{hyperref}
@@
 \renewcommand\UrlFont{\color{blue}\rmfamily}
+\newcommand\keccak{\textsc{Keccak}}
 
@@
 %
-\title{Verification of OpenSSL's SHA3 Keccak function using the Software Analysis Workbench}
+\title{Verification of OpenSSL's SHA3 \keccak\ function using the Software Analysis Workbench}
 %
@@
   Benjamin Winters\inst{1} \and
-  Eric Mercer\inst{1}\orcidID{0000--0002-2264-2958} \and
+  Eric Mercer\inst{1}\orcidID{0000-0002-2264-2958} \and
   Bret Decker\inst{1}
@@
 
-%%\input{preliminaries}
+\section{SHA-3 Overview}
+\input{sha-3}
 
-\section{Proof Overview and Results}\label{sec:proof}
+\section{Proof Outline and Results}\label{sec:proof}
 \input{proof}
 
-\section{SHA3 FIPS Specification in Cryptol}\label{sec:fips}
-\input{fips}
+\section{FIPS 202 Specification in Cryptol}\label{sec:fips}
+\input{fips202}
 
@@
 \input{openssl}
-
-\section{Results and Discussion}\label{sec:results}
-\input{results}
 

### 2021-11-22 paper/proof.tex
Move imp to figs and move figs to proof section
Restructure sections and file names

@@
 SHA 3 Details
+\begin{figure*}[ht]
+  \centering
+  \includegraphics[width=\linewidth]{figs/proof.png}
+
+  \caption{SHA3 Keccak Proof Structure}
+  \label{fig:proofStructure}
+
+\end{figure*}
+\begin{figure*}[ht]
+  \centering
+  \includegraphics[width=\linewidth]{figs/proof2.png}
+
+  \caption{Inner Function Contracts}
+  \label{fig:proofStructure2}
+
+\end{figure*}
+
 \begin{compactitem}
@@
 \end{itemize}
+
+Talk about the cost of the verification. The amount of effort required to write the specification. Make clear the limits of the result proof.
+
+\begin{compactitem}
+\item Visual inspection and limited test vectors for (1) and (2)
+\item Only provide proof certificates for message sizes of X to Y for digest sizes of \textbf{FIXME} for (3)
+\item Visual inspection only for anything above Keccak for (3) in OpenSSLs C-code
+\end{compactitem}
+
+Effort (see above and then consider below):
+\begin{compactitem}
+ \item How long were we working on the Cryptol code?
+ \item How long were we working on the SAW proofs?
+ \item How long did it take to understand the OpenSSL code?
+\end{compactitem}
+We'll want to determine how long it took us and then try to infer how long that would have taken had we been working full time (40 hours/week).
+
+Because of the modular design of the SHA-3 specification, the Cryptol implementation and OpenSSL implementation are programmatically more similar. 
+The only strong discrepancy is that the OpenSSL implementation organizes the bits of the sponge construction. 
+The affect of modular cryptographic design can be seen by the proof completing successfully in the Yices, Z3, and ABC SMT-solvers. 
+Previous work on SHA2-256 was only successful with Yices \cite{nfm-us}. Full results are show in table \ref{finaltable}.
+
+\begin{table*}[b]
+\caption{Proof Runtimes for SHA-3 with SMT-solvers}\label{finaltable}
+\setlength{\tabcolsep}{13.5pt}
+\begin{tabular}{|l|r|r|r|}
+\hline
+\textbf{Function} & \textbf{Yices} & \textbf{Z3} & \textbf{ABC} \\
+\hline
+Pi          &   0.895 secs &    11.072 secs &    0.417 secs \\
+Rho         &   4.509 secs &    15.776 secs &    5.879 secs \\
+Theta       &  25.942 secs &    11.345 secs &   14.869 secs \\
+Chi         &   5.059 secs &     3.930 secs &    1.327 secs \\
+Iota        &  11.048 secs &     8.801 secs &  612.497 secs \\
+KeccakF1600 & 745.206 secs & 15187.671 secs & 2910.549 secs \\
+\hline
+\end{tabular}
+\end{table*}

### 2021-11-22 paper/sha-3.tex
Restructure sections and file names

Overview the SHA-3 algorithm.
Add enough detail for the proof-outline and results to make sense in the next section.

### 2021-11-22 paper/sha-3.tex
Add text for sha3 overview

@@
-Overview the SHA-3 algorithm.
-Add enough detail for the proof-outline and results to make sense in the next section.
+\begin{algorithm}[t]
+  \caption{\shaThree\ sponge construction with \keccak}\label{alg:sha3}
+  \begin{algorithmic}[1]
+    \Require $n \geq 0$ and $(d,r) \in \{(224, 1152)\ (256, 1088)\ (384, 832)\ (512, 576)\}$
+    \Ensure $D =$ \shaThree(N,d,r)
+    \State $p_0, \ldots, p_m = \mathtt{pad}(N, r)$\label{line:pad}
+    \State $S = 0_0, \ldots, 0_{1599}$\label{line:init}
+    \For{$p_i = p_0, \dots, p_m$}
+      \State $S = S\ \mathbf{xor}\ \mathtt{zeroExtend}(p_i, s-r)$\label{line:absorb}
+      \State $S =$ \keccak$(S)$\label{line:keccak}
+    \EndFor
+    \State $d = \mathtt{truncate(S,d)}$\label{line:squeeze}
+  \end{algorithmic}
+\end{algorithm}
+
+\shaThree\ uses \emph{sponge construction} to generate hashes as shown in \algoref{alg:sha3} (see \nist\ \fips\ \cite{fips202}).
+Sponge construction \emph{absorbs} data into an internal state $S$ at a rate of $r$-bits per cycle (\lineref{line:absorb}), transforming the state with the absorbed data on each cycle (\lineref{line:keccak}), and once all the data is absorbed, then it \emph{squeezes} out a final $d$-bit hash (\lineref{line:squeeze}).
+The transformation function, \keccak-f[1600], simply \keccak\ in the rest of this presentation, defines the size of the state $S$ to be $s = 1600$-bits, which is initialized to 0 (\lineref{line:init}).
+
+The state itself consists of $r$-bits that are written or read plus a capacity of $c = s - r$ that is untouched by input or output.
+The \nist\ standard defines pairs of allowed hash sizes and rates as required in \algoref{alg:sha3}.
+The input $N$ is padded with $10^*1$ such that its final size is a multiple of $r$ bits, and it is partitioned into $r$-bit size blocks (\lineref{line:pad}) for the absorption phase.
+
+The \keccak\ transformation function takes place over twenty-four rounds with each round consisting of five steps: $\theta$, $\rho$, $\pi$, $\chi$, and $\iota$.
+The state $S$ itself is viewed as a $5 \times 5$ array of $64$-bit words (e.g.
+$5 * 5 * 64 = 1600$) in each of the steps.
+$\theta$ computes a parity over the columns in the state.
+$\rho$ is a bitwise rotate over the state.
+$\pi$ is a permutation on the state.
+$\chi$ is a non-linear operation that combines along the rows in the state.
+And $\iota$ mixes in a constant into a portion of the state.

### 2021-11-23 paper/sha-3.tex
Revise algorithm details and clarify what is proved

@@
     \EndFor
-    \State $d = \mathtt{truncate(S,d)}$\label{line:squeeze}
+    \State $d = \mathtt{squeeze(S,d)}$\label{line:squeeze}
   \end{algorithmic}
@@
 \shaThree\ uses \emph{sponge construction} to generate hashes as shown in \algoref{alg:sha3} (see \nist\ \fips\ \cite{fips202}).
-Sponge construction \emph{absorbs} data into an internal state $S$ at a rate of $r$-bits per cycle (\lineref{line:absorb}), transforming the state with the absorbed data on each cycle (\lineref{line:keccak}), and once all the data is absorbed, then it \emph{squeezes} out a final $d$-bit hash (\lineref{line:squeeze}).
+Sponge construction \emph{absorbs} data into an internal state $S$ at a rate of $r$-bits per cycle (\lineref{line:absorb} and \lineref{line:keccak}), transforming the state on each cycle, and once all the data is absorbed, then it \emph{squeezes} out a final $d$-bit hash (\lineref{line:squeeze}).
 The transformation function, \keccak-f[1600], simply \keccak\ in the rest of this presentation, defines the size of the state $S$ to be $s = 1600$-bits, which is initialized to 0 (\lineref{line:init}).
@@
 $\chi$ is a non-linear operation that combines along the rows in the state.
-And $\iota$ mixes in a constant into a portion of the state.
+And $\iota$ mixes in twenty-four different constants, one for each iteration, into portions of the state.

### 2021-11-30 paper/paper.tex
Add text for sha3 overview
Add background on SAW and SHA-3

@@
 \usepackage{graphicx}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 \usepackage{color}
@@
 \newcommand\keccak{\textsc{Keccak}}
+\newcommand\shaThree{\texttt{SHA-3}}
+\newcommand\nist{\texttt{NIST}}
+\newcommand\fips{\texttt{FIPS 202}}
+\newcommand\saw{\texttt{SAW}}
+\newcommand\cryptol{Cryptol}
+\newcommand\openssl{OpenSSL}
+\newcommand{\figref}[1]{Fig.~\ref{#1}}
+\newcommand{\algoref}[1]{Alg.~\ref{#1}}
+\newcommand{\lineref}[1]{Line~\ref{#1}}
 
@@
 
-\section{SHA-3 Overview}
+\section{Background}\label{sec:background}
+\cryptol\ and \saw\ are used to prove properties of \openssl's implementation of \shaThree\ relative to the \nist\ \fips\ publication \cite{fips202}. 
+What the proofs actually say, and an understanding of their limitations, relies on (1) a basic knowledge of how \saw\ works and what \cryptol\ does relative to \saw, and (2) how the \shaThree\ algorithm constructs its digests.
+Key elements relative to understanding the proof of each (\cryptol, \saw, and \shaThree) are defined in this section.
+
+\subsection{Cryptol and SAW}\label{subsec:saw}
+\input{saw}
+\subsection{SHA-3 Overview}\label{subsec:sha3}
 \input{sha-3}

### 2021-11-30 paper/proof.tex
Revise algorithm details and clarify what is proved
Add background on SAW and SHA-3

@@
   \item After the last round, keccak turns the state vector back into the bit-string./\item The next unprocessed block is \emph{sponged} (e.g., absorbed) into the current bit-string from the latest keccak and the process is repeated
-  \item The max digest (hash) size is 512 bits in the standard
+  \item The digest size is defined in pairs by the standard.
   \item The final bit-string from keccak after sponging and processing every block in the original message is truncated to the desired digest size
+  \item The brute force proof is for a digest of 256-bits.
 \end{compactitem}
@@
 SAW Details
-\begin{compactitem}
-  \item Cryptol is a specification language specifically for defining crypto-primitives
-  \item It has a compiler to turn Cryptol to SAW core terms
-  \item It has an execution environment to run test vectors on specifications
-  \item SAW is a tool to prove equivalence between SAW core terms
-  \item It uses symbolic execution and must unroll every loop to a fixed length
-  \item It can take as input C and Java code and compile those to core terms
-  \item It can use overides to replace C and Java code in the term
-\end{compactitem}
 

### 2021-11-30 paper/saw.tex
Add background on SAW and SHA-3

\cryptol\ is a domain specific language for describing cryptographic primitives, such as algorithms to compute digests, at the bit-level.
Cryptographic primitives are the building blocks for complex cryptographic protocols, such as SSL, that provide a broad range of security guarantees.
\cryptol\ is executable, being able to run test vectors to validate specifications, and more importantly, it compiles directly to \saw\ core terms for formal reasoning.

\saw\ is a tool to extract formal models from programs with support for C, Java, and \cryptol\ input.
These formal models are referred to as core terms in \saw.
\saw's function, among other things, is to prove input to output equivalence between core terms.
It accomplishes such proofs with equivalence preserving rewrites on the core terms and symbolic execution.
The rewrites simplify the terms for symbolic execution as core term size correlates with complexity. 
The rewrites also try to structurally reduce one term to the other thus avoiding symbolic execution altogether.
Symbolic execution reasons about all computation paths in the core terms and proves equivalence in computation over all inputs.
Symbolic execution must unroll loops in any core term a static number of times to accomplish the equivalence proof.
As such the number of iterations in any loop must be statically known at compile time.

\saw\ supports hierarchical reasoning with \emph{overrides}.
An override replaces one core term with another in the formal analysis.
As an example, consider a Java method that itself calls another Java method as part of its implementation.
Suppose that \saw\ is able to prove that a \cryptol\ specification for the called method and the Java implementation of the same method are equivalent.
If that is possible, then an override is able to replace the called Java method with the core term from the \cryptol\ in the model extraction from Java.
Core terms from \cryptol\ are smaller in size and simpler in structure than equivalent terms extracted from C or Java thereby reducing the size and complexity of the extracted core term in \saw.
As such, proving a property about a complex C function or Java method generally requires the use of overrides from \cryptol\ in order for symbolic execution to be successful.
Indeed, the use of overrides are key to the \saw\ proof in this paper showing equivalence between the \shaThree\ specification and the \openssl\ C implementation.

### 2021-11-30 paper/sha-3.tex
Add background on SAW and SHA-3

@@
 \shaThree\ uses \emph{sponge construction} to generate hashes as shown in \algoref{alg:sha3} (see \nist\ \fips\ \cite{fips202}).
-Sponge construction \emph{absorbs} data into an internal state $S$ at a rate of $r$-bits per cycle (\lineref{line:absorb} and \lineref{line:keccak}), transforming the state on each cycle, and once all the data is absorbed, then it \emph{squeezes} out a final $d$-bit hash (\lineref{line:squeeze}).
-The transformation function, \keccak-f[1600], simply \keccak\ in the rest of this presentation, defines the size of the state $S$ to be $s = 1600$-bits, which is initialized to 0 (\lineref{line:init}).
+Sponge construction \emph{absorbs} data into an internal state $S$ at a rate of $r$ bits per cycle (\lineref{line:absorb} and \lineref{line:keccak}), transforming the state on each cycle, and once all the data is absorbed, then it \emph{squeezes} out a final $d$-bit digest (\lineref{line:squeeze}).
+The transformation function, \keccak-f[1600], simply \keccak\ in the rest of this presentation, defines the size of the state $S$ to be $s = 1600$ bits, which is initialized to 0 (\lineref{line:init}).
 
-The state itself consists of $r$-bits that are written or read plus a capacity of $c = s - r$ that is untouched by input or output.
-The \nist\ standard defines pairs of allowed hash sizes and rates as required in \algoref{alg:sha3}.
-The input $N$ is padded with $10^*1$ such that its final size is a multiple of $r$ bits, and it is partitioned into $r$-bit size blocks (\lineref{line:pad}) for the absorption phase.
+The state itself consists of $r$ bits that are written or read plus a capacity of $c = s - r$ bits that are untouched by input or output.
+The \nist\ standard defines pairs of allowed hash sizes and rates.
+These are defined in the requires statement in \algoref{alg:sha3}. 
+The input $N$ is padded such that its final size is a multiple of $r$ bits, and it is partitioned into $r$-bit size blocks (\lineref{line:pad}) for the absorption phase.
+The padding is defined by the regular expression $10^*1$.
+An extra block is added for the padding if needed since every $N$ must be a multiple of $r$-bit size blocks matching the $10^*1$ regular expression at the end of the last block.
 

### 2021-12-01 paper/proof.tex
Add proof overview

--- 2021-11-22 paper/proof.tex
@@
-SHA 3 Details
-\begin{compactitem}
-  \item Process 1600 bits at a time: $5 * 5 * 64$
-  \item State vector is $5*5*64*$
-  \item Incoming message padded out to whole number of 1600 bit bit-strings
-  \item Processed in order by \texttt{KECCAK}
-  \item Keccak turns each block into a state vector (rearranges bit layout in memory)
-  \item Keccak has 24 rounds for state vector. 
-  \item There are 5 hash functions sequentially applied in each round to iteratively update the state vector
-  \item After the last round, keccak turns the state vector back into the bit-string./\item The next unprocessed block is \emph{sponged} (e.g., absorbed) into the current bit-string from the latest keccak and the process is repeated
-  \item The max digest (hash) size is 512 bits in the standard
-  \item The final bit-string from keccak after sponging and processing every block in the original message is truncated to the desired digest size 
-\end{compactitem}
+\begin{figure*}[t]
+  \centering
+  \includegraphics[width=\linewidth]{figs/proof.png}
 
-SAW Details
-\begin{compactitem}
-  \item Cryptol is a specification language specifically for defining crypto-primitives
-  \item It has a compiler to turn Cryptol to SAW core terms
-  \item It has an execution environment to run test vectors on specifications
-  \item SAW is a tool to prove equivalence between SAW core terms
-  \item It uses symbolic execution and must unroll every loop to a fixed length
-  \item It can take as input C and Java code and compile those to core terms
-  \item It can use overides to replace C and Java code in the term
-\end{compactitem}
+  \caption{Equivalence proof for the \shaThree\ standard and the \openssl\ C implementation.}
+  \label{fig:proofStructure}
 
-SAW provides a verification environment for compositional proofs of functional equivalence between two pieces of supported code (C, Java, or Cryptol).
-SAW uses symbolic execution to create computational models of the code and creates problems of functional equivalence between those two models that can be pushed off to an SMT solver.  
-The difficulty of these SMT problems is directly affected by the size and complexity of the computations that are being compared.
-This can lead to drastic increases in proof times and sometimes even prevents completion as computations increase.  
-SAW's solution to this is to create contracts (called overrides) of a functions input to output relationship which allows for a modular approach to writing the proof.
-By breaking up computation into smaller pieces, and creating overrides for these pieces, SAW can often simplify an otherwise complex (and thus incompletable) proof to remain within the capacity of an SMT solver.
+\end{figure*}
 
-Cryptol is a domain-specific specification language, which has been used in the previously mentioned works by Galois' and by Decker et.
-al to write visually reviewable implementations of NIST standards.
-It has strong ties to the SAW tool and, in fact, can be compiled directly down to the SAW Terms (computational models) which are used by SAW in proving functional equivalence.
-Because of the interwoven nature of Cryptol with SAW, the Terms created from Cryptol are in most cases significantly simpler than any created through symbolic execution of C/Java code.
+The goal of the work presented in this paper is to prove the \openssl\ implementation of \shaThree\ matches the \fips\ publication by \nist.
+\figref{fig:proofStructure} illustrates the proof strategy.
+The proof begins on the left of the figure with the box labeled \emph{NIST Standard}. 
+This box represents the \fips\ standard that defines the \shaThree\ algorithm with english prose and math.
+The standard also provides several test inputs with the correspondingly  expected digests.
+The first step of the proof reproduces the \fips\ \shaThree\ description in \cryptol.
+That is the box to the right of the \emph{NIST Standard} box labeled \emph{Spec}.
 
-The published NIST standards provide no compileable implementation that can be used as a starting point for proving out the algorithm.
-As a result, any proof of correctness with a NIST standard is limited in strength to how simple it is use visual review to connect code used in the proof to the NIST standard, coupled with a small sampling of test vectors.
-As shown in \ref{proofStructure}, The proof flow starts with a specification of the SHA3 algorithm in Cryptol with the goal of being as similar visually to the NIST standard as possible.
-This work discusses significant ways that Cryptol's functional nature can be leveraged to aid in this goal.
-Along with simplifying the visual inspection process, this Cryptol specification also passes all test vectors provided by NIST.
+The equivalence between the \fips\ standard and the \cryptol\ specification is argued by visual inspection along with tests over the published test vectors with their expected digests.
+\shaThree\ is described algorithmically in the standard making use of universal quantifiers and other looping structures in the definition.
+\cryptol\ is functional in that it uses list comprehensions for iteration, and it does not have quantifiers.
+Visually certifying a specification of \shaThree\ that uses list comprehensions directly for quantification and looping is difficult at best---the list comprehensions are too complex to read, and they obscure the algorithmic, and mathematic, structure in the published standard.
 
-The end goal of this proof is to show functional equivalence between this visually reviewable specifications Keccak function, and that of OpenSSL's C code.
-Unfortunately, a direct proof between this specification, and OpenSSL's implementation proved unfeasible, because OpenSSL varies from the NIST standard in how it stores the bits of the intermediate state array in memory.
-These minor modifications in memory storage put such a proof out of SAW's capabilities, because it relies on strict typing, and comparison of a Cryptol functions input/output to a C functions memory state.
-In order to bridge that gap, a second Cryptol specification is included in the proof flow.
-It matches in every way the first Cryptol code, except when it comes to how the state array is stored and accessed (at which point it follows OpenSSL's form).
-Because of the efficient nature of Cryptol's compilation to SAW Terms, proving equivalence between these Cryptol implementations was possible at the top level of the SHA3 algorithm (which allows for the differences in state array storage to be fully abstracted out).
-Due to the strict typing nature required for writing SAW proofs, the scope of this equivalence proof is limited to a given input message size.
-The only computations that change in the SHA3 algorithm due to message size are in the padding and sponging of a block.
-With this in mind, an exhaustive proof run over each input message byte count from 1 to the block size is sufficient to cover the differences in computation.
-The artifact repository accompanying this work contains the receipts of this exhaustive proof. (REFERENCE TO PROOF!!!!)
+The syntactic disconnect between the quantification and looping in the published standard and the list comprehensions in the \cryptol\ specification is bridged with a library that hides the list comprehensions with appropriately named functions.
+The final specification with the function hiding the list comprehensions has a one-to-one correspondence to the published standard that is easy to visually certify.
+The specification is further validated by tests in \cryptol\ over the published test vectors.
+The digests computed from the \cryptol\ specification exactly match those published in the standard on each given input.
+The equivalence arrow between the \emph{NIST Standard} box and the \emph{Spec} box represents the visual certification and test vector results.
+\secref{sec:fips} details the specification and library.
 
-INNER GREEK FUNCTIONS:
+The extracted core term from the \openssl\ C implementation of \shaThree\ is too big and complex for \saw\ to reason about directly.
+Adding to the complexity of the equivalence proof is that the \openssl\ implementation changes the meaning of the state $S$ in the algorithm so that it no longer directly matches the meaning defined in the published standard.
+The change effectively reorders the matrix definition of the state to be more amenable to memory and aspects of the \keccak\ implementation in C.
+The change is significant enough to make it extremely difficult to argue manually that the computation implemented in \openssl\ matches that defined in the standard.
 
-CONTRACTS CAN BE USED TO OVERRIDE AND REPLACE C WITH Cryptol
+The next step of the equivalence proof creates a \cryptol\ specification that matches the state reordering in the \openssl\ implementation in an effort to show an equivalence between it and the \cryptol\ specification for the standard.
+This new specification is the box in the right middle of \figref{fig:proofStructure} labeled \emph{OpenSSL} with the \emph{CRY} annotation in the upper right corner.
+\saw\ is not able to reason symbolically about input and digest sizes to construct a general proof of equivalence between the published standard and the \openssl\ \cryptol specifications, but it is able to prove equivalence between the two specifications given specific sizes for the input and digest.
 
-WITH THESE CONTRACTS SAW CAN COMPLETE A PROOF OF EQUIVALENCE BETWEEN TWO PIECES OF CODE.
+A series of \saw\ proofs over a range of input sizes with a digest size fixed at $256$-bits is used to prove the equivalence between the published standard and the \openssl\ \cryptol specifications.
+The series of proofs varies the input size from zero bits to $1,087$ bits to reflect the rate $1,088$ bit rate required for a 256-bit digest.
+The proofs establish the correctness of the input padding and covers the situation where the padding adds an additional $1,088$-bit block to the input 
+As such, the meaning of the equivalence arrow between the \emph{Spec} box and the \emph{OpenSSL} box in the middle of the figure is that the two specifications have equivalent computation for a $256$-bit digest for message sizes from $0$ bits to $1,087$ bits.
 
-This work relies on the correctness of several tools, including language compilers.
+\begin{figure*}[t]
+  \centering
+  \includegraphics[width=\linewidth]{figs/proof2.png}
 
-\noindent \textbf{Trusted Code Base}:
-\begin{itemize}
-  \item LLVM bitcode compiler
-  \item The Software Analysis Workbench (SAW) including the conversion from LLVM bitcode to SAWCore terms, the Cryptol compiler to SAWCore terms, and SAW's interface with SMT solvers.
-  \item The z3, abc, and yices SMT solvers.
-\end{itemize}
+  \caption{Inner Function Contracts}
+  \label{fig:proofStructure2}
+
+\end{figure*}
+
+The rest of the equivalence proof between the standard and the \openssl\ implementation centers on the \keccak\ algorithm as shown in the bottom right of \figref{fig:proofStructure}.
+The meaning of the equivalence arrow between the \keccak\ definition in the \cryptol\ specification based on the \openssl\ implementation and the actual implemented \keccak\ algorithm in the \openssl\ C implementation is that their computation exactly matches for any input state.
+That proof cannot be constructed directly by \saw\ since the extracted core term from the C implementation is too big for symbolic execution when considering all twenty-four rounds required by \keccak.
+The full proof requires overrides.
+
+The proof of equivalence between the specification of \keccak\ and the C implementation of \keccak\ in \openssl\ relies on overrides in \saw as shown in \figref{fig:proofStructure2}.
+Here, \saw\ proves equivalence between the \cryptol\ specification and the C implementation for each of the five functions that comprise one round of \keccak\ given some arbitrary input state.
+That is the meaning of the right hand side of \figref{fig:proofStructure2}.
+These are equivalent on any input.
+
+The implementations for each of the five \keccak\ functions in the C implementation are replaced by the corresponding equivalent \cryptol\ specifications in the extracted model from C.
+\saw\ is then able to prove the C implementation of \keccak equivalent to the \cryptol\ specification of the \openssl\ implementation of \keccak\ over all twenty-four rounds for any arbitrary input state.
+The equivalence is independent of the digest size, message size, or state.
+
+%% TODO: update the figure and text with results from the proof of the padding and the squeeze functions.
+
+In summary, the \cryptol\ specification of the \fips\ standard relies on visual inspection and test vectors for equivalence.
+The equivalence between the \cryptol\ specification of the standard and the \cryptol\ specification of \openssl's implementation is limited to the 256-bit digest on input sizes from $0$ bits to $1,087$ bits.
+That is a proof that considers all of \figref{fig:sha3}.
+The equivalence between \openssl\ \cryptol\ specification and the actual C implementation is limited to the \keccak algorithm only.
+That is a proof that only considers \lineref{line:keccak} in \figref{fig:sha3} but holds for any state and is independent of the digest size and rate.
+
+%% TODO: summarize the cost of the total proof in terms of time and identify the machine used for the proof.

### 2021-12-01 paper/proof.tex
Fix algorithm reference in proof

@@
 The equivalence between the \cryptol\ specification of the standard and the \cryptol\ specification of \openssl's implementation is limited to the 256-bit digest on input sizes from $0$ bits to $1,087$ bits.
-That is a proof that considers all of \figref{fig:sha3}.
+That is a proof that considers all of \algoref{alg:sha3}.
 The equivalence between \openssl\ \cryptol\ specification and the actual C implementation is limited to the \keccak algorithm only.
-That is a proof that only considers \lineref{line:keccak} in \figref{fig:sha3} but holds for any state and is independent of the digest size and rate.
+That is a proof that only considers \lineref{line:keccak} in \algoref{alg:sha3} but holds for any state and is independent of the digest size and rate.
 

### 2021-12-20 paper/paper.tex
Add proof overview
Minor revisions on background and proof

--- 2021-11-22 paper/paper.tex
@@
 % to display URLs in blue roman font according to Springer's eBook style:
-\renewcommand\UrlFont{\color{blue}\rmfamily}
-\newcommand\keccak{\textsc{Keccak}}
+\renewcommand{\UrlFont}{\color{blue}\rmfamily}
+\newcommand{\keccak}{\textsc{Keccak}}
+\newcommand{\shaThree}{\texttt{SHA-3}}
+\newcommand{\nist}{\texttt{NIST}}
+\newcommand{\fips}{\texttt{FIPS 202}}
+\newcommand{\saw}{\texttt{SAW}}
+\newcommand{\cryptol}{Cryptol}
+\newcommand{\openssl}{OpenSSL}
+\newcommand{\figref}[1]{Fig.~\ref{#1}}
+\newcommand{\algoref}[1]{Alg.~\ref{#1}}
+\newcommand{\lineref}[1]{Line~\ref{#1}}
+\newcommand{\secref}[1]{Sec.~\ref{#1}}
 
@@
 
-\section{SHA-3 Overview}
+\section{Background}\label{sec:background}
+\cryptol\ and \saw\ are used to prove properties of \openssl's implementation of \shaThree\ relative to the \nist\ \fips\ publication \cite{fips202}. 
+What the proofs actually say, and an understanding of their limitations, relies on (1) a basic knowledge of how \saw\ works and what \cryptol\ does relative to \saw, and (2) how the \shaThree\ algorithm computes a digest.
+This section provides that basic knowledge of the tools and a general overview of the \shaThree\ algorithm.
+
+\subsection{\cryptol\ and \saw}\label{subsec:saw}
+\input{saw}
+\subsection{SHA-3 Overview}\label{subsec:sha3}
 \input{sha-3}
@@
 
-\section{FIPS 202 Specification in Cryptol}\label{sec:fips}
+\section{FIPS 202 Specification in \cryptol}\label{sec:fips}
 \input{fips202}
 
-\section{SHA3 OpenSSL in Cryptol}\label{sec:openssl}
+\section{\shaThree\ from \openssl\ in \cryptol}\label{sec:openssl}
 \input{openssl}

### 2021-12-20 paper/proof.tex
Minor revisions on background and proof

@@
 This new specification is the box in the right middle of \figref{fig:proofStructure} labeled \emph{OpenSSL} with the \emph{CRY} annotation in the upper right corner.
-\saw\ is not able to reason symbolically about input and digest sizes to construct a general proof of equivalence between the published standard and the \openssl\ \cryptol specifications, but it is able to prove equivalence between the two specifications given specific sizes for the input and digest.
+\saw\ is not able to reason symbolically about input and digest sizes to construct a general proof of equivalence between the published standard and the \openssl\ \cryptol\ specifications, but it is able to prove equivalence between the two specifications given specific sizes for the input and digest.
 
-A series of \saw\ proofs over a range of input sizes with a digest size fixed at $256$-bits is used to prove the equivalence between the published standard and the \openssl\ \cryptol specifications.
-The series of proofs varies the input size from zero bits to $1,087$ bits to reflect the rate $1,088$ bit rate required for a 256-bit digest.
-The proofs establish the correctness of the input padding and covers the situation where the padding adds an additional $1,088$-bit block to the input 
-As such, the meaning of the equivalence arrow between the \emph{Spec} box and the \emph{OpenSSL} box in the middle of the figure is that the two specifications have equivalent computation for a $256$-bit digest for message sizes from $0$ bits to $1,087$ bits.
+A series of \saw\ proofs over a range of input sizes with the digest size fixed at $256$ bits is used to prove the equivalence between the published standard and the \openssl\ \cryptol\ specifications.
+The series of proofs varies the input size from zero bits to $1,087$ bits to reflect the $1,088$ bit rate required for a 256 bit digest.
+The proofs establish the correctness of the input padding and cover the situation where the padding adds an additional $1,088$ bit block to the input 
+As such, the meaning of the equivalence arrow between the \emph{Spec} box and the \emph{OpenSSL} box in the middle of the figure is that the two specifications have equivalent computation for a $256$ bit digest for message sizes from $0$ bits to $1,087$ bits.
 
@@
 
-The proof of equivalence between the specification of \keccak\ and the C implementation of \keccak\ in \openssl\ relies on overrides in \saw as shown in \figref{fig:proofStructure2}.
+The proof of equivalence between the specification of \keccak\ and the C implementation of \keccak\ in \openssl\ relies on overrides in \saw\ as shown in \figref{fig:proofStructure2}.
 Here, \saw\ proves equivalence between the \cryptol\ specification and the C implementation for each of the five functions that comprise one round of \keccak\ given some arbitrary input state.
-That is the meaning of the right hand side of \figref{fig:proofStructure2}.
+That is the meaning of the left hand side of \figref{fig:proofStructure2}.
 These are equivalent on any input.
@@
 The implementations for each of the five \keccak\ functions in the C implementation are replaced by the corresponding equivalent \cryptol\ specifications in the extracted model from C.
-\saw\ is then able to prove the C implementation of \keccak equivalent to the \cryptol\ specification of the \openssl\ implementation of \keccak\ over all twenty-four rounds for any arbitrary input state.
+\saw\ is then able to prove the C implementation of \keccak\ equivalent to the \cryptol\ specification of the \openssl\ implementation of \keccak\ over all twenty-four rounds for any arbitrary input state.
 The equivalence is independent of the digest size, message size, or state.
@@
 In summary, the \cryptol\ specification of the \fips\ standard relies on visual inspection and test vectors for equivalence.
-The equivalence between the \cryptol\ specification of the standard and the \cryptol\ specification of \openssl's implementation is limited to the 256-bit digest on input sizes from $0$ bits to $1,087$ bits.
+The equivalence between the \cryptol\ specification of the standard and the \cryptol\ specification of \openssl's implementation is limited to the 256 bit digest on input sizes from $0$ bits to $1,087$ bits.
 That is a proof that considers all of \algoref{alg:sha3}.
-The equivalence between \openssl\ \cryptol\ specification and the actual C implementation is limited to the \keccak algorithm only.
+The equivalence between the \openssl\ \cryptol\ specification and the actual C implementation is limited to the \keccak\ algorithm only.
 That is a proof that only considers \lineref{line:keccak} in \algoref{alg:sha3} but holds for any state and is independent of the digest size and rate.

### 2022-01-04 paper/fips202.tex
Restructure sections and file names
Revise the for-method text

--- 2021-09-16 paper/fips.tex
@@
-\section{SHA3 FIPS Specification in Cryptol}
-\label{sec:fips}
-Initially it appeared that visual inspection with Cryptol's standard syntax was subpar.
-The language's functional folds and list comprehensions do not align easily with the imperative and array-based NIST standard. 
-By leveraging Cryptol's versatility, utility functions can visually mimic the sequential nature of the NIST specification.
+This section illustrates with a few examples how the \fips\ specification is captured in \cryptol. 
+\cryptol's functional folds and list comprehensions do not directly align with the quantification and iteration in the specification. 
+This misalignment made the \cryptol\ specification difficult to read and to manually argue its equivalence relative to the published specification.
+The two specifications are realigned through a library of methods in \cryptol\ that hide the functional folds and list comprehensions so that the \cryptol\ specification reads more like the \fips\ specification.
+The two most common library methods are the \emph{for-method} and the \emph{while-method} discussed in this section.
 
-\subsection{For Utility Function}
-The phrase \emph{for all} indicates a looping structure and is quite common in the NIST standard. 
-It is used in the definition of the pi function shown in Figure \ref{fig:nistPi}.
+\subsection{The for-method}
 
-\begin{figure}[h]
-  \centering
-  \begin{lstlisting}[basewidth = {.5em},basicstyle={\scriptsize}]
+\newsavebox{\fipsPi}
+\begin{lrbox}{\fipsPi}
+  \begin{lstlisting}[basewidth = {.5em},basicstyle={\small}]
     1. For all triples (x,y,z) such that 
@@
   \end{lstlisting}
-  \caption{NIST Pi definition}
-  \label{fig:nistPi}
-\end{figure}
+\end{lrbox}
 
-By nesting Cryptol's list comprehensions, the pi function can be easily implemented.
-This is illustrated in Figure \ref{fig:cryptolPi}.
-
-\begin{figure}[h]
-  \centering
-\begin{lstlisting}[language=Cryptol]
-type COL = 5 
-type W = 64
-type STATE_ARR = [COL][COL][W]
-LIST4 = [0..COL - 1]
-LIST63 = [0..W - 1]
+\newsavebox{\PiCry}
+\begin{lrbox}{\PiCry}
+  \begin{lstlisting}[language=Cryptol, numbers=left, escapeinside={;}{;}]
+    type STATE_ARR = [5][5][64];\label{line:state};
+    LIST4 = [0..4];\label{line:list4};
+    LIST63 = [0..63];\label{line:list63};
     pi : STATE_ARR -> STATE_ARR
-pi a = [[[a @x @((x + 3*y) % 'COL) @z 
+    pi a = [  [  [a @x @((x + 3*y) % 5) @z;\label{line:pi};
                  | z <- LIST63] 
@@
   \end{lstlisting}
-\caption{Cryptol Pi definition}
-\label{fig:cryptolPi}
+\end{lrbox}
+
+\newsavebox{\PiCryLib}
+\begin{lrbox}{\PiCryLib}
+  \begin{lstlisting}[language=Cryptol]
+    pi a = for LIST4 (&x ->
+             for LIST4 (&y -> 
+               for LIST63 (&z -> 
+                 a @x @((x + 3*y) % 5) @z)))
+  \end{lstlisting}
+\end{lrbox}
+
+\newsavebox{\formethod}
+\begin{lrbox}{\formethod}
+  \begin{lstlisting}[language=Cryptol]
+    for : {n, a, b} [n]a -> (a -> b) -> [n]b
+    for vals f = [f i | i <- vals]
+  \end{lstlisting}
+\end{lrbox}
+
+\begin{figure}[t]
+  \begin{center}
+    \begin{tabular}{l}
+      \usebox{\fipsPi} \\ \\
+      \multicolumn{1}{c}{(a)} \\ \\
+      \usebox{\PiCry} \\ \\
+      \multicolumn{1}{c}{(b)} \\ \\
+      \usebox{\formethod} \\ \\
+      \multicolumn{1}{c}{(d)} \\ \\ 
+      \usebox{\PiCryLib} \\ \\
+      \multicolumn{1}{c}{(c)}
+    \end{tabular}
+  \end{center}
+  \caption{The $\pi$ function. (a) The \fips\ specification. (b) The \cryptol\ with comprehensions. (c) The for-method definition. (d) The \cryptol\ with the for-method.}
+  \label{fig:pi}
 \end{figure}
 
-Cryptol declares list comprehension variables after they are used, and the order of the nested loops is not as apparent. 
-Using a helper function, \emph{for}, a clearer definition seen in Figure \ref{fig:cryptolAmendedPi}.
+A common idiom in the \fips\ publication is using quantification over a finite domain defined by linear constraints to define input for some computation.
+The definition of the $\pi$ function for \keccak\ shown in \figref{fig:pi}(a) is one such example.
+Here the quantification is used to define the set of indices that are part of the $\pi$ computation that transforms the state array.
+The intent is that of a \emph{parallel-for} where order does not matter since the new state array $A^\prime$ only depends on the old state array $A$.
 
-\begin{figure}[h]
-  \centering
-\begin{lstlisting}[language=Cryptol]
-type COL = 5 
-type W = 64
-type STATE_ARR = [COL][COL][W]
-LIST4 = [0..COL - 1]
-LIST63 = [0..W - 1]
-pi : STATE_ARR -> STATE_ARR
-pi a = for LIST4 (&x ->
-          for LIST4 (&y -> 
-            for LIST4 (&z -> 
-              a @x @((x + 3*y) % 'COL) @z)))
-\end{lstlisting}
-\caption{Amended Cryptol Pi definition}
-\label{fig:cryptolAmendedPi}
-\end{figure}
+The \cryptol\ definition using list comprehensions is given in \figref{fig:pi}(b).
+\lineref{line:state}) defines the state as a $5 \times 5$ array of 64-bit words.
+\lineref{line:list4} and \lineref{line:list63} are the lists for the list comprehension.
+These simple define the indices that are used in the computation.
+\lineref{line:pi} starts the definition of $\pi$. 
+Here the comprehensions are nested with the list indices appearing in reverse order to follow the nesting.
+Although the list comprehensions are not unreadable, the \cryptol\ can be improved by creating a for-method to hide some of the details.
 
-The \emph{for} utility function follows the same intuition as the implementation in most imperative languages in that it iterates and runs a loop body over defined elements. 
-It recieves a function and a list of arguments and returns a list of results from calling the function on every element of the input array. 
-Figure \ref{fig:cryptolFor} shows it's implementation.
+\figref{fig:pi}(c) is the \cryptol\ definition of the for-method.
+It takes two arguments: a list, \emph{vals}, for the comprehension, and a function, $f$, to apply to each element of the list.
+Its definition is as expected: it applies $f$ to each element in \emph{vals}.
 
-\begin{figure}[h]
-  \centering
-\begin{lstlisting}[language=Cryptol]
-for : {n, a, b} [n]a -> (a -> b) -> [n]b
-for vals loop = [loop i | i <- vals]
-\end{lstlisting}
-\caption{Cryptol For utility function}
-\label{fig:cryptolFor}
-\end{figure}
+\figref{fig:pi}(d) is the rewritten \cryptol\ using the for-method.
+It still requires the indices list for the comprehension, but it has a more obvious correspondence to the original definition in \figref{fig:pi}(a).
+The for-method reorders and structures the arguments in a very elegant way.
+It hides the list comprehensions to simplify the task of visual inspection for equivalence.
 
-Though it is simple to implement, it reorders and structures the arguments in a very elegant way.
-In constrast with standard cryptol syntax, the list to iterate over and the iterating variable are visible before the loop body. 
-List comprehensions are hidden away behind this function call and visual equivalence is clearer.
-
-\subsection{While Utility Function}
+\subsection{The while-method}
 Another common phrasing is the \emph{let...for...let}. 

### 2022-01-04 paper/fips202.tex
Add minor revisions to for-method

@@
 \lineref{line:state}) defines the state as a $5 \times 5$ array of 64-bit words.
-\lineref{line:list4} and \lineref{line:list63} are the lists for the list comprehension.
-These simple define the indices that are used in the computation.
+\lineref{line:list4} and \lineref{line:list63} are the lists for the list comprehensions.
+These are the indices that are used in the computation and correspond to the finite domains for $x$, $y$, and $z$ defined in the quantification in \figref{fig:pi}(a).
 \lineref{line:pi} starts the definition of $\pi$. 
@@
 \figref{fig:pi}(d) is the rewritten \cryptol\ using the for-method.
-It still requires the indices list for the comprehension, but it has a more obvious correspondence to the original definition in \figref{fig:pi}(a).
+It still requires the indices lists for the comprehension, but it has a more obvious correspondence to the original definition in \figref{fig:pi}(a).
 The for-method reorders and structures the arguments in a very elegant way.

### 2022-01-04 paper/fips202.tex
Revise the while-method and text;2C

@@
     type STATE_ARR = [5][5][64];\label{line:state};
-    LIST4 = [0..4];\label{line:list4};
-    LIST63 = [0..63];\label{line:list63};
     pi : STATE_ARR -> STATE_ARR
     pi a = [  [  [a @x @((x + 3*y) % 5) @z;\label{line:pi};
-                 | z <- LIST63] 
-              | y <- LIST4] 
-           | x <- LIST4]
+                 | z <- [0..63]];\label{line:list};
+              | y <- [0..4]] 
+           | x <- [0..4]]
   \end{lstlisting}
@@
   \begin{lstlisting}[language=Cryptol]
-    pi a = for LIST4 (&x ->
-             for LIST4 (&y -> 
-               for LIST63 (&z -> 
+    pi a = for [0..4] (&x ->
+             for [0..4] (&y -> 
+               for [0..63] (&z -> 
                  a @x @((x + 3*y) % 5) @z)))
@@
       \usebox{\formethod} \\ \\
-      \multicolumn{1}{c}{(d)} \\ \\ 
+      \multicolumn{1}{c}{(c)} \\ \\ 
       \usebox{\PiCryLib} \\ \\
-      \multicolumn{1}{c}{(c)}
+      \multicolumn{1}{c}{(d)}
     \end{tabular}
@@
 \lineref{line:state}) defines the state as a $5 \times 5$ array of 64-bit words.
-\lineref{line:list4} and \lineref{line:list63} are the lists for the list comprehensions.
-These are the indices that are used in the computation and correspond to the finite domains for $x$, $y$, and $z$ defined in the quantification in \figref{fig:pi}(a).
 \lineref{line:pi} starts the definition of $\pi$. 
 Here the comprehensions are nested with the list indices appearing in reverse order to follow the nesting.
+\lineref{line:list} gives the domain for $z$ as a list with values from 0 to 63.
+The domains for $x$ and $y$ are given similarly.
+The domains for $x$, $y$, and $z$ all match the domains defined in the quantification in \figref{fig:pi}(a).
 Although the list comprehensions are not unreadable, the \cryptol\ can be improved by creating a for-method to hide some of the details.
@@
 It takes two arguments: a list, \emph{vals}, for the comprehension, and a function, $f$, to apply to each element of the list.
-Its definition is as expected: it applies $f$ to each element in \emph{vals}.
+Its definition is as expected: it creates a new list that is the result of applying $f$ to each element in \emph{vals}.
 
@@
 \subsection{The while-method}
-Another common phrasing is the \emph{let...for...let}. 
-This defines a state upon which a folding occurs, iteratively condensing the information into a final state. 
-Imperatively, it is most comparable to a while loop. 
-As shown by Figure \ref{fig:nistRC}, is present in the rc function defined by the NIST specification.
 
-\begin{figure}[h]
-  \centering
-\begin{lstlisting}[basewidth = {.5em},basicstyle={\scriptsize}]
+\newsavebox{\fipsRc}
+\begin{lrbox}{\fipsRc}
+  \begin{lstlisting}[basewidth = {.5em},basicstyle={\small}]
     1. If t mod 255 = 0, return 1.
@@
   \end{lstlisting}
-\caption{NIST RC definition}
-\label{fig:nistRC}
-\end{figure}
+\end{lrbox}
 
-Maintaining this state through Cryptol's list comprehensions can be difficult.
-The comprehensions require a list to iterate through. 
-Since the \emph{for} utility function is simply a wrapper for list comprehensions, it is equally unsuited for visual verification on such steps. 
-The number of iterations in the RC function is bounded by [1, 254] by its modular condition.
-This range provides the list to iterate over as required by Cryptol list comprehensions. 
-From these observations, a rather ungainly Cryptol definition results.
-This definition is shown in Figure \ref{fig:cryptolRC}.
-
-\begin{figure}[h]
-  \centering
-\begin{lstlisting}[language=Cryptol]
+\newsavebox{\RcCry}
+\begin{lrbox}{\RcCry}
+  \begin{lstlisting}[language=Cryptol, numbers=left, escapeinside={;}{;}]
     rc : [64] -> Bit
-rc t = if (t % 255) == 0 
-  then True 
-  else rs !0 @0 where
-    rs = [128:[8]] # [
-      if i <= t % 255
-        then (take'{8} ([
-          (r@0) ^ (r@8),
-          r@1,
-          r@2,
-          r@3, 
+    rc t =  if (t % 255) == 0 then 1
+            else rs !0 @0 where;\label{line:return};
+              rs = [0b10000000] #;\label{line:init};
+                   [if i <= t % 255 then;\label{line:compute};
+                      (take'{8} ([(r@0) ^ (r@8),
+                                   r@1, r@2, r@3, 
                                   (r@4) ^ (r@8),
@@
                                   (r@6) ^ (r@8),
-          r@7,
-          r@8])
-            where r = [False] # (rs @(i - 1)))
-        else rs @(i - 1)
-      | i <- [1..254]]
+                                   r@7,r@8])
+                        where r = [0] # (rs @(i - 1)))
+                    else rs @(i - 1);\label{line:stutter};
+                   | i <- [1..254]];\label{line:alliters};
   \end{lstlisting}
-\caption{Cryptol RC definition}
-\label{fig:cryptolRC}
-\end{figure}
+\end{lrbox}
 
-Not only is this definition computationally slower as it runs 255 loops every time it is called, it also awkwardly indexes into a list of previous states. 
-This differs vastly from the NIST specificiation both visually and functionally. 
-The \emph{while} utility function, accepts an initial state, condition function, and body function to recursively loop. 
-It consolodates the information of the previous state until the condition is sastified. 
-This definition is found in Figure \ref{fig:cryptolamendedRC}.
+\newsavebox{\whilemethod}
+\begin{lrbox}{\whilemethod}
+  \begin{lstlisting}[language=Cryptol]
+    while : {a} a -> (a -> Bit) -> (a -> a) -> a
+    while state cond f = 
+      if (cond state) then (while (f state) cond f)	
+      else state
+  \end{lstlisting}
+\end{lrbox}
 
-\begin{figure}[h]
-  \centering
-\begin{lstlisting}[language=Cryptol]
+\newsavebox{\RcCryLib}
+\begin{lrbox}{\RcCryLib}
+  \begin{lstlisting}[language=Cryptol, numbers=left, escapeinside={;}{;}]
     rc : [64] -> Bit
-rc t = if (t % 255) == 0 
-  then True 
-  else (while {i = 1, R = 128:[8]}
-    (&state -> state.i <= t % 255)
+    rc t =  if (t % 255) == 0 then 1 
+            else (while {i = 1, R = 0b10000000};\label{line:state};
+                        (&state -> state.i <= t % 255);\label{line:cond};
                         (&state -> {
-      i = state.i + 1, 
-      R = (take'{8} ([
-        (r@0) ^ (r@8),
-        r@1,
-        r@2,
-        r@3,
+                          i = state.i + 1,;\label{line:inc};
+                          R = (take'{8} ([(r@0) ^ (r@8),
+                                           r@1, r@2, r@3,
                                           (r@4) ^ (r@8),
@@
                                           (r@6) ^ (r@8),
-        r@7, 
-        r@8]) 
-          where r = [False] # state.R)})
-    ).R @0
+                                           r@7, r@8]) 
+                                where r = [0] # state.R)})
+                 ).R @0;\label{line:libreturn};
   \end{lstlisting}
-\caption{Amended Cryptol RC definition}
-\label{fig:cryptolamendedRC}
+\end{lrbox}
+
+\begin{figure}[t]
+  \begin{center}
+    \begin{tabular}{l}
+      \usebox{\fipsRc}
+    \end{tabular}
+  \end{center}
+  \caption{The \fips\ specification for \emph{rc}.}
+  \label{fig:rc}
 \end{figure}
 
-The final state is returned by the \emph{while} function. 
-This iterates only until the condition is met, avoids the clutter of indexing within a larger array of all previous states, and places the condition in a more obvious location before the loop body. 
-This utility function is defined in Figure \ref{fig:cryptolWhile}.
+Another common idiom in the \fips\ standard is to iteratively transform some state a fixed number of times to arrive at a final state.
+The definition of the \emph{rc} function in \figref{fig:rc} is one such example.
+The \emph{rc} function is a support function for the $\iota$ function in \keccak\ and is used to build a mask.
+
+The \emph{rc} function checks the value of the input integer $t$, and returns the bit $0$ if it is a multiple of $255$.
+If not, then Line 2 defines $R$ as the 8-bit value $128$, and Line 3 iteratively transforms $R$ a fixed number of times: $t\ \mathbf{mod}\ 255$.
+The transform is defined on Line 3a through Line 3f.
+Line 3a prepends a $0$ bit to $R$ so that it is now a 9-bit value.
+The array indexing refers to bit positions from left to right so $R[0]$ is the prepended left-most bit and $R[8]$ is the right-most bit.
+Each transform on $R$ prepends the 0 (Line 3a), does an exclusive-or with specific bit positions (Line 3b through Line 3e), and then takes the left-most 8 bits for the next round effectively dropping the right-most bit (Line 3f).
+It returns the $R[0]$ at the end.
+
+The \cryptol\ equivalent for \emph{rc} in \figref{fig:rccry}(a) is considerably more complex with the list comprehensions.
+Returning $1$ when the input is a multiple of $255$ is straightforward.
+It is when that is not the case that it is less obvious.
+The list comprehension builds a list of $255$ 8-bit values.
+\lineref{line:return} returns the bit at index 0 ($@0$) from the last 8-bit value in the array ($!0$).
+The computation of the list of 8-bit values follows.
+
+\lineref{line:init} is the first entry in the list and it is the initial value of $R$.
+The following entries depend on the value of the input $t$ and the value of $i$ that comes from the list used in the list comprehension on \lineref{line:alliters}.
+\lineref{line:compute} checks if $i$ is within the number of iterations, and if it is, then the new entry in the list is computed from the previous entry as defined by the transform in \figref{fig:rc}.
+\lineref{line:stutter} is the case for when $i$ is not within the number of iterations.
+Here the previous value in the list is copied to stutter the last value of $R$. 
+The \cryptol\ specification creates all $255$ entries in the list regardless of the input value $t$ except for the case when $t$ is a multiple of 255.
+As before, \figref{fig:rccry}(a) is not altogether unreadable, but it can be improved considerably.
+
+\figref{fig:rccry}(b) is a recursive definition of a \emph{while-loop}.
+It takes as input a state, a predicate on the state, \emph{cond}, that is true when it should loop and false otherwise, and a function, $f$, that computes a new state from the old state on each iteration.
+The definition makes recursive calls to the while-method until the state predicate is false which point it returns the current value of \emph{state}.
 
 \begin{figure}[h]
-  \centering
-\begin{lstlisting}[language=Cryptol]
-while : {a} a -> (a -> Bit) -> (a -> a) -> a
-while state cond f = if (cond state)
-  then (while (f state) cond f)	
-  else state
-\end{lstlisting}
-\caption{Cryptol While utility function}
-\label{fig:cryptolWhile}
+  \begin{center}
+    \begin{tabular}{l}
+      \usebox{\RcCry} \\ \\
+      \multicolumn{1}{c}{(a)} \\ \\
+      \usebox{\whilemethod} \\ \\
+      \multicolumn{1}{c}{(b)} \\ \\ 
+      \usebox{\RcCryLib} \\ \\
+      \multicolumn{1}{c}{(c)}
+    \end{tabular}
+  \end{center}
+  \caption{The \emph{rc} function in \cryptol. (b) The \cryptol\ with list comprehensions. (b) The recursive while-method definition. (c) The \cryptol\ with the while-method.}
+  \label{fig:rccry}
 \end{figure}
+
+\figref{fig:rccry}(c) is the definition of \emph{rc} using the while-method in \figref{fig:rccry}(b).
+\lineref{line:state} defines the state for the while loop to be both an integer $i$ for a counter and the 8-bit value $R$.
+\lineref{line:cond} is the looping condition defined on the state.
+The body of the loop not only computes the new value of $R$ but \lineref{line:inc} increments $i$ to track the loop iteration.
+The return value is given on \lineref{line:libreturn} as the $0^\mathrm{th}$ bit from the last value of $R$.
+Here, the only computes $R$ the number of times required by $t$ which differs significantly from the definition without the while-method.
+As with the for-method, the while-method reduces the gap between how things are defined in \fips\ and in \cryptol.

### 2022-01-05 paper/saw.tex
Minor revisions on background and proof
Revise the overview of saw

@@
-\cryptol\ is a domain specific language for describing cryptographic primitives, such as algorithms to compute digests, at the bit-level.
+\cryptol\ is a domain specific functional language for describing cryptographic primitives, such as algorithms to compute digests, at the bit-level.
 Cryptographic primitives are the building blocks for complex cryptographic protocols, such as SSL, that provide a broad range of security guarantees.
-\cryptol\ is executable, being able to run test vectors to validate specifications, and more importantly, it compiles directly to \saw\ core terms for formal reasoning.
+\cryptol\ is executable, being able to run test vectors to validate specifications, and more importantly, it compiles directly to \sawcore\ which is the language for formal models in \saw.
 
 \saw\ is a tool to extract formal models from programs with support for C, Java, and \cryptol\ input.
-These formal models are referred to as core terms in \saw.
-\saw's function, among other things, is to prove input to output equivalence between core terms.
-It accomplishes such proofs with equivalence preserving rewrites on the core terms and symbolic execution.
-The rewrites simplify the terms for symbolic execution as core term size correlates with complexity. 
-The rewrites also try to structurally reduce one term to the other thus avoiding symbolic execution altogether.
-Symbolic execution reasons about all computation paths in the core terms and proves equivalence in computation over all inputs.
-Symbolic execution must unroll loops in any core term a static number of times to accomplish the equivalence proof.
-As such the number of iterations in any loop must be statically known at compile time.
+These formal models are expressed in the \sawcore\ language that supports formal reasoning.
+As mentioned previously, \cryptol\ compiles directly to \sawcore, but that is not the case for C and Java.
+\saw\ uses \emph{symbolic execution} to extract \sawcore\ models from C and Java inputs.
+
+Symbolic execution reasons about all computation paths in the C or Java input.
+It must unroll all loops a static number of times to create the model in \sawcore.
+As such the number of iterations in any loop must be statically known at compile time or the symbolic execution fails to terminate.
+\sawcore\ models from symbolic execution tend to be more complex than those from \cryptol.
+
+\saw's function, among other things, is to prove input to output equivalence between \sawcore\ models.
+It accomplishes such proofs with equivalence preserving rewrites on the \sawcore\ and with automated reasoning.
+The rewrites simplify the \sawcore\ models as size correlates with complexity in the automated reasoning. 
+The rewrites also try to structurally reduce one model to the other thus avoiding automated reasoning altogether.
+When rewriting is not sufficient, the equivalence between two models is reduced to a satisfiability problem and dispatched to external backend solvers.
 
 \saw\ supports hierarchical reasoning with \emph{overrides}.
-An override replaces one core term with another in the formal analysis.
+An override replaces one \sawcore\ model with another in the formal analysis.
 As an example, consider a Java method that itself calls another Java method as part of its implementation.
 Suppose that \saw\ is able to prove that a \cryptol\ specification for the called method and the Java implementation of the same method are equivalent.
-If that is possible, then an override is able to replace the called Java method with the core term from the \cryptol\ in the model extraction from Java.
-Core terms from \cryptol\ are smaller in size and simpler in structure than equivalent terms extracted from C or Java thereby reducing the size and complexity of the extracted core term in \saw.
-As such, proving a property about a complex C function or Java method generally requires the use of overrides from \cryptol\ in order for symbolic execution to be successful.
+If that is possible, then an override is able to replace the called Java method with the \sawcore\ from the \cryptol\ and as a result avoid any symbolic execution of that method when creating the \sawcore\ model of the Java.
+
+As mentioned previously, \sawcore\ from \cryptol\ is less complex than equivalent models created by symbolic execution from C or Java; thereby overrides reduce the overall complexity of the final extracted \sawcore\ models from C or Java.
+As such, proving a property about a complex C function or Java method generally requires the use of overrides from \cryptol\ in order for the equivalence proof to be successful, and that is regardless of how that proof is accomplished.
+Overrides improve the likelihood of the rewrites showing equivalence and the likelihood of the backend solver being able to show equivalence if the rewrites fail.
 Indeed, the use of overrides are key to the \saw\ proof in this paper showing equivalence between the \shaThree\ specification and the \openssl\ C implementation.

### 2022-01-05 paper/proof.tex
Add minor revisions
Revise proof overview with subsections

@@
 This box represents the \fips\ standard that defines the \shaThree\ algorithm with english prose and math.
-The standard also provides several test inputs with the correspondingly  expected digests.
+The standard also provides several test inputs with the corresponding expected digests.
 The first step of the proof reproduces the \fips\ \shaThree\ description in \cryptol.
 That is the box to the right of the \emph{NIST Standard} box labeled \emph{Spec}.
+
+\subsection{Visual Inspection and Test Vectors}
 
@@
 The syntactic disconnect between the quantification and looping in the published standard and the list comprehensions in the \cryptol\ specification is bridged with a library that hides the list comprehensions with appropriately named functions.
-The final specification with the function hiding the list comprehensions has a one-to-one correspondence to the published standard that is easy to visually certify.
+The final specification with the function hiding the list comprehensions has a more direct correspondence to the published standard that is easy to visually certify.
 The specification is further validated by tests in \cryptol\ over the published test vectors.
@@
 
-The extracted core term from the \openssl\ C implementation of \shaThree\ is too big and complex for \saw\ to reason about directly.
+\subsection{Equivalence between Memory Layouts}
+
+The extracted \sawcore\ from the \openssl\ C implementation of \shaThree\ is too big and complex for \saw\ to reason about directly.
 Adding to the complexity of the equivalence proof is that the \openssl\ implementation changes the meaning of the state $S$ in the algorithm so that it no longer directly matches the meaning defined in the published standard.
@@
 
+\subsection{Overrides and Keccak}
+
 The rest of the equivalence proof between the standard and the \openssl\ implementation centers on the \keccak\ algorithm as shown in the bottom right of \figref{fig:proofStructure}.
 The meaning of the equivalence arrow between the \keccak\ definition in the \cryptol\ specification based on the \openssl\ implementation and the actual implemented \keccak\ algorithm in the \openssl\ C implementation is that their computation exactly matches for any input state.
-That proof cannot be constructed directly by \saw\ since the extracted core term from the C implementation is too big for symbolic execution when considering all twenty-four rounds required by \keccak.
+That proof cannot be constructed directly by \saw\ since the extracted \sawcore\ from the C implementation is too big for automated reasoning when considering all twenty-four rounds required by \keccak.
 The full proof requires overrides.
@@
 That is the meaning of the left hand side of \figref{fig:proofStructure2}.
-These are equivalent on any input.
+These are equivalent on any state $S$.
 

### 2022-01-06 paper/fips202.tex
Revise OpenSSL

@@
 The two specifications are realigned through a library of methods in \cryptol\ that hide the functional folds and list comprehensions so that the \cryptol\ specification reads more like the \fips\ specification.
-The two most common library methods are the \emph{for-method} and the \emph{while-method} discussed in this section.
+The two library methods are the \emph{for-method} and the \emph{while-method} discussed in this section.
 
@@
 It takes as input a state, a predicate on the state, \emph{cond}, that is true when it should loop and false otherwise, and a function, $f$, that computes a new state from the old state on each iteration.
-The definition makes recursive calls to the while-method until the state predicate is false which point it returns the current value of \emph{state}.
+The definition makes recursive calls to the while-method until the state predicate is false at which point it returns the current value of \emph{state}.
 
@@
 The return value is given on \lineref{line:libreturn} as the $0^\mathrm{th}$ bit from the last value of $R$.
-Here, the only computes $R$ the number of times required by $t$ which differs significantly from the definition without the while-method.
-As with the for-method, the while-method reduces the gap between how things are defined in \fips\ and in \cryptol.
+Here, $R$ is computed only the number of times required by $t$.
+This change differs significantly from the definition without the while-method.
+As with the for-method, the while-method reduces the gap between how things are defined in \fips\ and how they are defined in \cryptol.

### 2022-01-06 paper/openssl.tex
Revise OpenSSL

@@
-Even with the highly modular approach of the SHA3 algorithm, a direct proof from the Cryptol specification that matched the NIST standard and the OpenSSL source code was unviable.
-OpenSSL's source code is often used in low-level software, and it is often optimized to improve performance and viability for embedded systems.
-One major instance of such optimization proved significant enough to prevent a direct proof from the NIST Cryptol specification to OpenSSL.
+This section details the two differences between the \openssl\ implementation of \shaThree\ and the \fips\ specification that prevented a direct proof of equivalence between the \cryptol\ and C code.
+These differences are expressed in a second \cryptol\ model derived from the \fips\ model discussed in the previous section.
+\saw\ proves these two models equivalent for the 256-bit digest size as discussed in \secref{sec:proof}.
+This equivalence considers the whole of the sponge construction algorithm proving that the digest is the same from each of them.
+That said, and as a reminder, the equivalence between the \keccak\ description in \cryptol\ and the C implementation in \openssl\ holds for any input message and any digest size.
+That proof uses the \cryptol\ definition for the model discussed in this section that includes to two differences seen in \openssl, and the overrides used in that proof also come from the model discussed in this section.
+
+\subsection{State Array Structure}
+
+The first difference in the \openssl\ \shaThree\ implementation is in the structure of the state array.
+The difference is an artifact of how C maps structured data to memory.
+The \fips\ structures the state array as a $5 \times 5$ matrix of 64-bit words with each 64-bit word being a \emph{lane}.
+It assumes the memory layout of the data follows a normal cartesian three dimensional coordinate system with $x$ being the horizontal axis, $y$ being the vertical axis, and $z$ being the depth access an a line to get an individual bit from the lane.
+For a state $A$, $A[x,y,z]$ accesses $z^\mathrm{th}$ bit from the 64-bit word on the $x^\mathrm{th}$ column and the $y^\mathrm{th}$ row.
+That is the mapping to a cartesian coordinate system.
+
+\openssl\ declares the state as follows: \texttt{uint64\textunderscore t A[5][5]}.
+The C standard stores multidimensional arrays in contiguous memory in \emph{row-major} order meaning that each row of five 64-bit values appear consecutively in memory.
+Indexing the multidimensional array follows the standard mathematical definition for indexing matrices: $A[x,y]$ is the element at the $x^\mathrm{th}$ row and $y^\mathrm{th}$ column.
+This meaning is just opposite of that in the standard.
+
+Adding to the complexity is that the C standard does not provide array indexing to get a bit from a value.
+For the $A[x,y,z]$ example, there is no array bracket notation to get the $z^\mathrm{th}$ bit in a 64-bit word, so notation in the \fips\ standard such as that seen in the definition of the $rc$ function in \figref{fig:rc}(a) has no direct analogue in C.
+As such, the \openssl\ implementation does everything at the level of the lanes, operating on each lane as a 64-bit entity, and never refers to an individual bit in a lane.
+Finally, bit ordering in the lanes in the \fips\ standard is just opposite the ordering in C meaning that the direction of shifting in standard is just opposite the direction used in C.
+
+The \cryptol\ for the \fips\ model is rewritten to reflect these memory layout differences.
+The $x$ and $y$ indexing is swapped.
+All the inner \keccak\ functions are modified to operate on lanes in their entirety.
+And the the shifts are reversed.
+For reference, \figref{fig:piopenssl} is the rewritten $\pi$ function in \cryptol\ than matches \openssl's implementation (compare to \figref{fig:pi}(d)).
+Operating on lanes in faster and more efficient.
+That is reflected in the \cryptol\ running times.
+
+\newsavebox{\PiOpenSSL}
+\begin{lrbox}{\PiOpenSSL}
+  \begin{lstlisting}[language=Cryptol]
+    pi a = for [0..4] (&y ->
+             for [0..4] (&x -> 
+                 a @x @((x + 3*y) % 5))
+  \end{lstlisting}
+\end{lrbox}
+
+\begin{figure}
+  \begin{center}
+    \usebox{\PiOpenSSL}
+  \end{center}
+  \caption{The \openssl\ implementation of $\pi$ that reverses indexes and operates on lanes.}
+  \label{fig:piopenssl}
+\end{figure}
+
+\subsection{Lookup Tables for Computed Constants}
+
+The second difference is in using a lookup table for constants rather than computing constants.
 
 \lstset{style=customc, firstnumber=177}
-\begin{figure}[h]
+\begin{figure}[t]
   \centering
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
   \centering
@@
   iota : STATE_ARR -> [64] -> STATE_ARR
-  iota a i = for LIST4 (\y ->
-                for LIST4 (\x ->
-                    for LIST63 (\z ->
+  iota a i = for [0..4] (\y ->
+                for [0..4] (\x ->
+                    for [0..63] (\z ->
                         if ((x == 0) && (y == 0))
@@
 
-A second obstacle that blocked a single direct proof is how OpenSSL stores and accesses its state array in memory.  
-The NIST specification outlines the state array as being a 5 by 5 by w 3-dimensional array.
-It denotes x as the horizontal axis, y as the vertical axis, and w as the z axis, like a cartesian coordinate system.
-Because it structures the array this way, any access into the array \texttt{A[x, y, z]} is asking for the element at the x'th column, y'th row, z'th bit.
-OpenSSL defines their state array as a \texttt{uint64\textunderscore t A[5][5]}.
-As is common practice in many programming styles, their array indexes by row and then by column.
-Because of this choice, an index defined by NIST as \texttt{A[x, y, z]} would in OpenSSL's array be indexed as \texttt{bit z of A[y][x]}.
-In comparing C code to Cryptol, SAW relies on the pre and post state of memory for a given function to help determine functional equivalence.
-The difference in memory storage/access choices by NIST and OpenSSL means that SAW cannot show functional equivalence using memory comparison.
-
-OpenSSL's choice in memory storage yields another obstacle because of its array composition.
-The way that bits are ordered and stored in a \texttt{uint64\textunderscore t} in C is strictly defined by the language's semantics.
-Unfortunately, that bit ordering is reversed to the ordering NIST defines.
-This yields yet another way that the memory differs and cannot be directly compared.
-
 Between these two significant deviations by the OpenSSL implementation from the NIST standard, an intermediate step was necessary to complete a proof of equivalence.

### 2022-01-06 paper/fips202.tex
Minor revision and clarification

@@
 This section illustrates with a few examples how the \fips\ specification is captured in \cryptol. 
-\cryptol's functional folds and list comprehensions do not directly align with the quantification and iteration in the specification. 
-This misalignment made the \cryptol\ specification difficult to read and to manually argue its equivalence relative to the published specification.
-The two specifications are realigned through a library of methods in \cryptol\ that hide the functional folds and list comprehensions so that the \cryptol\ specification reads more like the \fips\ specification.
+\cryptol's list comprehensions do not directly align with the quantification and iteration in the specification. 
+This misalignment makes the \cryptol\ specification more difficult to read and to manually argue its equivalence relative to the published specification.
+The two specifications are better aligned through a library of two methods in \cryptol\ that hide the list comprehensions so that the \cryptol\ specification reads more like the \fips\ specification.
 The two library methods are the \emph{for-method} and the \emph{while-method} discussed in this section.
@@
   \begin{lstlisting}[language=Cryptol, numbers=left, escapeinside={;}{;}]
-    type STATE_ARR = [5][5][64];\label{line:state};
+    type STATE_ARR = [5][5][64];\label{line:pistate};
     pi : STATE_ARR -> STATE_ARR
@@
 The \cryptol\ definition using list comprehensions is given in \figref{fig:pi}(b).
-\lineref{line:state}) defines the state as a $5 \times 5$ array of 64-bit words.
+\lineref{line:pistate} defines the state as a $5 \times 5$ array of 64-bit words.
 \lineref{line:pi} starts the definition of $\pi$. 
@@
 \figref{fig:pi}(d) is the rewritten \cryptol\ using the for-method.
-It still requires the indices lists for the comprehension, but it has a more obvious correspondence to the original definition in \figref{fig:pi}(a).
+It still requires the indices lists for the list comprehensions, but it has a more obvious correspondence to the original definition in \figref{fig:pi}(a).
 The for-method reorders and structures the arguments in a very elegant way.
@@
 Each transform on $R$ prepends the 0 (Line 3a), does an exclusive-or with specific bit positions (Line 3b through Line 3e), and then takes the left-most 8 bits for the next round effectively dropping the right-most bit (Line 3f).
-It returns the $R[0]$ at the end.
+It returns the $R[0]$ bit at the end.
 

### 2022-01-06 paper/openssl.tex
Minor revision and clarification

@@
-This section details the two differences between the \openssl\ implementation of \shaThree\ and the \fips\ specification that prevented a direct proof of equivalence between the \cryptol\ and C code.
+This section details the key differences between the \openssl\ implementation of \shaThree\ and the \fips\ specification that prevented a direct proof of equivalence between the \cryptol\ and C code.
 These differences are expressed in a second \cryptol\ model derived from the \fips\ model discussed in the previous section.
@@
 That said, and as a reminder, the equivalence between the \keccak\ description in \cryptol\ and the C implementation in \openssl\ holds for any input message and any digest size.
-That proof uses the \cryptol\ definition for the model discussed in this section that includes to two differences seen in \openssl, and the overrides used in that proof also come from the model discussed in this section.
+That proof uses the \cryptol\ definition for the model discussed in this section that includes all the differences seen in \openssl, and the overrides used in that proof also come from the \cryptol\ discussed in this section.
 
-\subsection{State Array Structure}
+\subsection{State Array Structure and Computation}
 
-The first difference in the \openssl\ \shaThree\ implementation is in the structure of the state array.
-The difference is an artifact of how C maps structured data to memory.
-The \fips\ structures the state array as a $5 \times 5$ matrix of 64-bit words with each 64-bit word being a \emph{lane}.
-It assumes the memory layout of the data follows a normal cartesian three dimensional coordinate system with $x$ being the horizontal axis, $y$ being the vertical axis, and $z$ being the depth access an a line to get an individual bit from the lane.
-For a state $A$, $A[x,y,z]$ accesses $z^\mathrm{th}$ bit from the 64-bit word on the $x^\mathrm{th}$ column and the $y^\mathrm{th}$ row.
-That is the mapping to a cartesian coordinate system.
+The first set of differences in the \openssl\ \shaThree\ implementation is in the structure of the state array and how it operates on that state array.
+The difference in structure is an artifact of how C maps arrays to memory.
+The \fips\ structures the state array as a $5 \times 5$ grid of 64-bit words with each 64-bit word being a \emph{lane}.
+It assumes the layout of the data follows a normal cartesian three dimensional coordinate system with $x$ being the horizontal axis, $y$ being the vertical axis, and $z$ being the depth on a lane to access an individual bit.
+For a state $A$, $A[x,y,z]$ accesses the $z^\mathrm{th}$ bit from the 64-bit word on the $x^\mathrm{th}$ column and the $y^\mathrm{th}$ row.
 
@@
 For the $A[x,y,z]$ example, there is no array bracket notation to get the $z^\mathrm{th}$ bit in a 64-bit word, so notation in the \fips\ standard such as that seen in the definition of the $rc$ function in \figref{fig:rc}(a) has no direct analogue in C.
-As such, the \openssl\ implementation does everything at the level of the lanes, operating on each lane as a 64-bit entity, and never refers to an individual bit in a lane.
-Finally, bit ordering in the lanes in the \fips\ standard is just opposite the ordering in C meaning that the direction of shifting in standard is just opposite the direction used in C.
+The consequence is that the \openssl\ implementation does everything at the level of the lanes, operating on each lane as a 64-bit entity, and it never refers to an individual bit in a lane.
+Finally, the bit ordering in the lanes in the \fips\ standard is just opposite the ordering in C meaning that the direction of shifting in the standard is opposite the direction used in the C implementation.
 
-The \cryptol\ for the \fips\ model is rewritten to reflect these memory layout differences.
+The \cryptol\ for the \fips\ model is rewritten to reflect these memory layout and computation differences.
 The $x$ and $y$ indexing is swapped.

### 2022-01-06 paper/proof.tex
Revise OpenSSL
Minor revision and clarification

@@
 \cryptol\ is functional in that it uses list comprehensions for iteration, and it does not have quantifiers.
-Visually certifying a specification of \shaThree\ that uses list comprehensions directly for quantification and looping is difficult at best---the list comprehensions are too complex to read, and they obscure the algorithmic, and mathematic, structure in the published standard.
+Visually certifying a specification of \shaThree\ that uses list comprehensions is less direct.
+The list comprehensions can be complex, and at times, they obscure the algorithmic, and mathematic, structure in the published standard.
 
-The syntactic disconnect between the quantification and looping in the published standard and the list comprehensions in the \cryptol\ specification is bridged with a library that hides the list comprehensions with appropriately named functions.
-The final specification with the function hiding the list comprehensions has a more direct correspondence to the published standard that is easy to visually certify.
+The syntactic disconnect between the quantification and looping in the published standard and the functional nature of the \cryptol\ specification is bridged with a library that hides list comprehensions with appropriately named functions.
+The final specification using the library has a more direct correspondence to the published standard that is more directly visually certified.
 The specification is further validated by tests in \cryptol\ over the published test vectors.
@@
 
-\subsection{Equivalence between Memory Layouts}
+\subsection{Memory Layout and Optimized Computation}
 
@@
 Adding to the complexity of the equivalence proof is that the \openssl\ implementation changes the meaning of the state $S$ in the algorithm so that it no longer directly matches the meaning defined in the published standard.
-The change effectively reorders the matrix definition of the state to be more amenable to memory and aspects of the \keccak\ implementation in C.
-The change is significant enough to make it extremely difficult to argue manually that the computation implemented in \openssl\ matches that defined in the standard.
+The change effectively reorders the matrix definition of the state to be more amenable to memory, and it changes the different functions in \keccak\ to be more efficient when operating on the state.
+The changes are significant enough to make it extremely difficult to argue manually that the computation implemented in \openssl\ matches that defined in the standard.
 
@@
 
-\subsection{Overrides and Keccak}
+\subsection{Overrides and \keccak}
 

### 2022-01-06 paper/saw.tex
Minor revision and clarification

@@
 It accomplishes such proofs with equivalence preserving rewrites on the \sawcore\ and with automated reasoning.
-The rewrites simplify the \sawcore\ models as size correlates with complexity in the automated reasoning. 
+The rewrites simplify the \sawcore\ for a models to reduce the cost of automated reasoning. 
 The rewrites also try to structurally reduce one model to the other thus avoiding automated reasoning altogether.
@@
 Suppose that \saw\ is able to prove that a \cryptol\ specification for the called method and the Java implementation of the same method are equivalent.
-If that is possible, then an override is able to replace the called Java method with the \sawcore\ from the \cryptol\ and as a result avoid any symbolic execution of that method when creating the \sawcore\ model of the Java.
+If that is possible, then an override is able to replace the called Java method with the \sawcore\ from the \cryptol;
+thereby, it bypasses the symbolic execution of that called method when creating the \sawcore\ model of the Java by using the \sawcore\ from the \cryptol\ specification directly.
 
-As mentioned previously, \sawcore\ from \cryptol\ is less complex than equivalent models created by symbolic execution from C or Java; thereby overrides reduce the overall complexity of the final extracted \sawcore\ models from C or Java.
+As mentioned previously, \sawcore\ from \cryptol\ is less complex than equivalent models created by symbolic execution from C or Java;
+thereby, the overrides reduce the overall complexity of the final extracted \sawcore\ from C or Java.
 As such, proving a property about a complex C function or Java method generally requires the use of overrides from \cryptol\ in order for the equivalence proof to be successful, and that is regardless of how that proof is accomplished.

### 2022-01-06 paper/related.tex
Updates related.tex.

Our work built upon the SHA2-256 inner hash function verification (!--Citation1--).
Citation1 expanded the methodology by which Amazon's s2n HMAC was verified (!--Citation2--).
Both compared equivalent C and Cryptol specification through SAW.
While SHA3 is not implemented in s2n, Citation2 utilized both a high-level and low-level cryptol specification, similar to our work.
Citation1's verification of OpenSSL's SHA2-256 implementation served as a direct predecessor to our work.
While SHA2 is also part of NIST's Secure Hash Algorithms, it's internal structure differs vastly from the more complex and resistant SHA3.

### 2022-01-06 paper/openssl.tex

@@
 For reference, \figref{fig:piopenssl} is the rewritten $\pi$ function in \cryptol\ than matches \openssl's implementation (compare to \figref{fig:pi}(d)).
-Operating on lanes in faster and more efficient.
-That is reflected in the \cryptol\ running times.
+Operating on lanes in faster and more efficient, and that change is reflected in all of the inner \keccak\ functions.
 
@@
 
-\subsection{Lookup Tables for Computed Constants}
+\subsection{Constant Lookup Tables}
 
-The second difference is in using a lookup table for constants rather than computing constants.
-
-\lstset{style=customc, firstnumber=177}
-\begin{figure}[t]
-  \centering
+\newsavebox{\ciota}
+\begin{lrbox}{\ciota}
   \begin{lstlisting}
-  void Iota(uint64_t A[5][5], size_t i)
-  {
+    void Iota(uint64_t A[5][5], size_t i) {
       //assert(i < (sizeof(iotas) / sizeof(iotas[0])));
@@
   \end{lstlisting}    
-\caption{OpenSSL's Iota Function}
-\label{fig:cIota}
-\end{figure}
+\end{lrbox}
 
-As per the NIST specification, the Iota function invokes the rc algorithm to generate a constant value specific to the round count to use in exclusive-or operations.
-The rc function inputs are unrelated to the message to be hashed, which means that for any given message, the iteration through the rounds of keccak will produce the same constant values.
-Because these values are unchanged from run to run, OpenSSL's optimized implementation of keccak stores the output values as a constant table and use the current round number to properly index into it as is shown in Figure \ref{fig:cIota}.
-Initially, the Cryptol code followed the NIST standard as closely as possible\textemdash including rc method calls in the Iota function which computed the values every iteration.
-Attempts to prove out such a specification to the OpenSSL code failed however, because SAW was unable to show the computational equivalence between computing values as needed and looking up values in a constant table.
+% iota : STATE_ARR -> [64] -> STATE_ARR
+% iota a i = for [0..4] (\y ->
+%               for [0..4] (\x ->
+%                   for [0..63] (\z ->
+%                       if ((x == 0) && (y == 0))
+%                           then (a @0 @0 @z) ^ (LISTIOTAS @i @z)
+%                       else a @y @x @z)))
 
-\begin{figure}[t]
-  \centering
+\newsavebox{\cryciota}
+\begin{lrbox}{\cryciota}
   \begin{lstlisting}[language=Cryptol]
-  iota : STATE_ARR -> [64] -> STATE_ARR
-  iota a i = for [0..4] (\y ->
-                for [0..4] (\x ->
-                    for [0..63] (\z ->
-                        if ((x == 0) && (y == 0))
-                            then (a @0 @0 @z) ^ (LISTIOTAS @i @z)
-	                        else a @y @x @z)))
-
     LISTIOTAS = [reverse (
-        (while {RC = 0:[W], j = 0} //STATE
-            (\state -> state.j <= `L) //COND
-            (\state -> { //LOOP
-                RC = for LIST63 (\z ->
-                    if z == index
-                        then rc (state.j + 7*i)
+      (while {RC = 0:[W], j = 0} 
+              (\state -> state.j <= `L) 
+              (\state -> { 
+                RC = for [0..63] 
+                  (\z -> if z == index then rc (state.j + 7*i)
                           else state.RC @z)
@@
       ).RC) | i <- LISTROUNDS:[_][64]]
+
+    LISTIOTAS = [
+      0x0000000000000001, 0x0000000000008082, 0x800000000000808a,
+      0x8000000080008000, 0x000000000000808b, 0x0000000080000001,
+      0x8000000080008081, 0x8000000000008009, 0x000000000000008a,
+      0x0000000000000088, 0x0000000080008009, 0x000000008000000a,
+      0x000000008000808b, 0x800000000000008b, 0x8000000000008089,
+      0x8000000000008003, 0x8000000000008002, 0x8000000000000080,
+      0x000000000000800a, 0x800000008000000a, 0x8000000080008081,
+      0x8000000000008080, 0x0000000080000001, 0x8000000080008008]
+        :[_][64]
   \end{lstlisting}    
-\caption{OpenSSL's Iota Function}
-\label{fig:cryptolIota}
+\end{lrbox}
+
+\begin{figure}[t]
+  \begin{center}
+    \begin{tabular}{l}
+      \usebox{\cryciota} \\ \\
+    \end{tabular}
+  \end{center}
+  \caption{Optimized $\iota$ function. (a) \openssl's implementation. (b) Matching \cryptol\ implementation.}
+  \label{fig:iota}
 \end{figure}
 
-The lack of any meaningful trace of computation that comes with pre-computed values, meant that a different approach would be necessary to meet in the middle.
-The solution was to modify the Cryptol specification to also use a lookup table in the Iota function.
-However, the Cryptol code still uses the rc function as defined in Figure \ref{fig:nistRC} to generate that lookup table at the beginning of every run of SHA3, rather than strictly stating constant values in the code.
-By having both the Cryptol and C Iota functions now rely on looking up values in a table, the task of proving them became far simpler.
-SAW only had to show that the table generated by running the rc function all at once at the beginning, was identical to the one hard coded into OpenSSL's implementation.
-Because those match, and both implementations index into the table in the same way, SAW was able to generate the proof.
-
-This approach recognizably added an extra element to the visual inspection between the Cryptol specification and the C code.
-The full rc computation is still used in the Cryptol code, with the only significant difference being that rather than run rc every time Iota is called, rc is run all at once to generate all the needed values.
-The continued inclusion of the rc computation makes it only a much less significant inconvenience to accept the computed table as accurate. 
-It likewise remains simple to see visually that indexing into the table gives back the proper constant value for a given round count.
-
-As a note, one minor adjustment was also made to OpenSSL's Iota function code to allow the proof to complete properly.
-As an error check, OpenSSL uses an assert to ensure that the round count is never greater than it should be (to ensure no indexing out of the array's bounds).
-SAW cannot recognize an assert statement in its symbolic execution engine.
-As a result, that statement was removed and replaced with a comparable if statement which maintained the logic of the function.
-Because the equivalence proof for the Iota function is run with the precondition that the round count must be less than 24 (max rounds), this does not affect the comparison between the C code and the Cryptol code (which does not have that error check).
-
-Between these two significant deviations by the OpenSSL implementation from the NIST standard, an intermediate step was necessary to complete a proof of equivalence.
-The solution is a second Cryptol implementation that matches OpenSSL's code in these memory differences.
-That second Cryptol code can be directly proven as equivalent to OpenSSL's code.
-
-This leaves only the step of proving the two Cryptol specifications as equivalent.
-Because of memory differences, this proof cannot be done on a low function level, as the OpenSSL proof is.
-Instead, it must be done at the top level of the SHA algorithm so that the memory compared (input message and output hash) is identical in structure, notwithstanding the difference in state array implementation.
-As SAW performs equivalence proofs based on defined static types, this proof can only be run on a specific input message size, and a specific output hash size.
-To attain full coverage this proof must be run on all input message sizes that affect the way the code is executed.
-This work determines the input set for full coverage to be all byte counts up to a full single block (200 Bytes), as this covers all padding and sponging operations performed in the creation of the state array.
-
-This approach required significant computation time and resources and cannot be easily replicated, however the artifact repository coupled with this work provides receipts of the output generated by this proof.
-However, with this step included a proof of functional equivalence between the visually reviewed Cryptol specification, and OpenSSL's Keccak function source code is complete.
+The second difference is in hardcoding lookup Tables for constants rather than computing the constants along the way.
+The \fips\ \cryptol\ specification computes the lookup table once for the rounding constants, \emph{rc}, for $\iota$, and it computes the rotate constants for $\rho$ on the fly.
+The \openssl\ specification hardcodes these constants in tables without the provenance of the computation.
+\figref{fig:iota} shows the difference in the two specifications for the rounding constants with the code for the computed table coming directly from \fips.

### 2022-01-06 paper/related.tex
Revises related.tex

@@
-Our work built upon the SHA2-256 inner hash function verification (!--Citation1--).
-Citation1 expanded the methodology by which Amazon's s2n HMAC was verified (!--Citation2--).
-Both compared equivalent C and Cryptol specification through SAW.
-While SHA3 is not implemented in s2n, Citation2 utilized both a high-level and low-level cryptol specification, similar to our work.
-Citation1's verification of OpenSSL's SHA2-256 implementation served as a direct predecessor to our work.
-While SHA2 is also part of NIST's Secure Hash Algorithms, it's internal structure differs vastly from the more complex and resistant SHA3.
+The verification of OpenSSL's SHA2-256 implementation served as a direct predecessor to this work (!Citation2).
+While SHA2 is also part of NIST's Secure Hash Algorithms, its internal structure differs vastly from the more complex and resistant SHA3.
+The SHA2 proof, in turn, expanded the methodology by which Amazon's s2n HMAC was verified (!Citation2).
+While SHA3 is not implemented in s2n, the HMAC verification utilized both a high-level and low-level cryptol specification in a similar manner to this work.
+Both prior works compared equivalent C and Cryptol specification through SAW.

### 2022-01-07 paper/abstract.tex
Move imp to figs and move figs to proof section
Restructure sections and file names
Change title and abstract

--- 2021-09-04 paper/abstract.tex
@@
-\begin{abstract}
-The Keccak function in the SHA-3 NIST standard is responsible for the major computations of all SHA-3 hashes and Shake functions.
-These functions are a key piece in modern secure hash functions. 
-This work extends and improves on the methodology used in the verification of Amazon’s s2n HMAC and OpenSSL’s SHA2-256 inner hash function. 
-It uses the Software Analysis Workbench (SAW) to prove out functional equivalence between OpenSSL’s SHA-3 Keccak function and a verified implementation of SHA-3 in the domain-specific language Cryptol.
-SAW creates contracts (called overrides), a functions input to output relationship, which allows for a modular approach to writing proofs. 
-This work leverages SAW by creating overrides to block out computation and creating a compositional proof that remains within the capacity of an SMT solver. 
-This work amends Cryptol through utility functions to handle looping which greatly simplifies the visual inspection the Cryptol implementation to provide strong assurance of correctness. 
-The results of this work indicate that this methodology works with modern cryptographic specifications and that a modular approach in cryptographic algorithm design greatly simplifies the verification process.
-\end{abstract}
+\shaThree\ is the new standard from \nist\ for computing digests. 
+\openssl\ is one of the most widely used implementations of cryptographic protocols.
+\cryptol\ and the Software Analysis Workbench (\saw) are a language and verification engine to specify cryptographic primitives and prove equivalence to implementations.
+As digests are the basis for all cryptographic protocols, the research in this paper is an automated \saw\ proof of equivalence between the \shaThree\ standard and the actual \openssl\ implementation in C.
+The research contributes \cryptol\ specifications for the published standard and the \openssl\ implementation of the standard with its several changes.
+The equivalence proof shows the importance of overrides in \saw\ to replace C code with \cryptol\ to reduce the complexity of the proof obligations that must be accomplished relative to the implementation in C.
+The research establishes the viability of verifying modern cryptographic primitives and the importance of modularity in their definitions, and implementations, for overrides.
+
+% The \keccak\ function in the SHA-3 NIST standard is responsible for the major computations of all SHA-3 hash constructs.
+% These constructs form the basis for modern secure hashing.
+% This work extends and improves the methodology used to verify Amazon's s2n HMAC and OpenSSL's SHA2-256 inner hash function. 
+% It uses the Software Analysis Workbench (SAW) to prove out functional equivalence between OpenSSL's SHA-3 Keccak function and a verified implementation of SHA-3 in Cryptol, a domain-specific language for cryptography.
+% SAW allows for a modular approach to writing proofs through the use of contracts, which define a functions input to output relationship.
+% Each contract, also called an override, is an obligation in the full proof.
+% This work leverages SAW by creating overrides to decompose computation and create a compositional proof that is within the capacity of an SMT solver. 
+% Visual inspection between Cryptol and the NIST standard is simplified through the creation of Cryptol utility functions.
+% These utility functions wrap the complex list comprehensions required for looping in Cryptol.
+% The results of this work indicates that a modular approach in cryptographic algorithm design greatly simplifies the verification process, as exemplified by SHA-3.

### 2022-01-07 paper/paper.tex
Revise the overview of saw
Revise OpenSSL
Fix Brett's name
Change title and abstract

@@
 \newcommand{\saw}{\texttt{SAW}}
+\newcommand{\sawcore}{SAWCore}
 \newcommand{\cryptol}{Cryptol}
@@
 %
-\title{Verification of OpenSSL's SHA3 \keccak\ function using the Software Analysis Workbench}
+\title{Verifying the \shaThree\ Implementation from \openssl\ with the Software Analysis Workbench}
 %
@@
   Eric Mercer\inst{1}\orcidID{0000-0002-2264-2958} \and
-  Bret Decker\inst{1}
+  Brett Decker\inst{1}
 }
@@
 
-\section{FIPS 202 Specification in \cryptol}\label{sec:fips}
+\section{\fips\ Specification in \cryptol}\label{sec:fips}
 \input{fips202}
 
-\section{\shaThree\ from \openssl\ in \cryptol}\label{sec:openssl}
+\section{\openssl\ Differences}\label{sec:openssl}
 \input{openssl}

### 2022-01-07 paper/abstract.tex
Add minor revisions

@@
 The research contributes \cryptol\ specifications for the published standard and the \openssl\ implementation of the standard with its several changes.
-The equivalence proof shows the importance of overrides in \saw\ to replace C code with \cryptol\ to reduce the complexity of the proof obligations that must be accomplished relative to the implementation in C.
+The equivalence proof shows the importance of overrides in \saw\ that replace C code with \cryptol\ to reduce the complexity of the proof obligations that must be accomplished relative to the implementation in C.
 The research establishes the viability of verifying modern cryptographic primitives and the importance of modularity in their definitions, and implementations, for overrides.

### 2022-01-07 keccak.c
Clean up README

@@ -4,9 +4,10 @@
 #include <stdio.h>
 #include <assert.h>
 
-# define BIT_INTERLEAVE (0)
-
+#define BIT_INTERLEAVE (0)
 #define ROL32(a, offset) (((a) << (offset)) | ((a) >> ((32 - (offset)) & 31)))
+
+
 
 static uint64_t ROL64(uint64_t val, int offset)
 {
@@ -32,6 +33,8 @@ static uint64_t ROL64(uint64_t val, int offset)
         return ((uint64_t)hi << 32) | lo;
     }
 }
+
+
 
 void Theta(uint64_t A[5][5])
 {
@@ -68,6 +71,7 @@ void Theta(uint64_t A[5][5])
 }
 
 
+
 static const unsigned char rhotates[5][5] = {
     {  0,  1, 62, 28, 27 },
     { 36, 44,  6, 55, 20 },
@@ -75,6 +79,8 @@ static const unsigned char rhotates[5][5] = {
     { 41, 45, 15, 21,  8 },
     { 18,  2, 61, 56, 14 }
 };
+
+
 
 void Rho(uint64_t A[5][5])
 {
@@ -88,6 +94,8 @@ void Rho(uint64_t A[5][5])
         A[y][4] = ROL64(A[y][4], rhotates[y][4]);
     }
 }
+
+
 
 void Pi(uint64_t A[5][5])
 {
@@ -127,6 +135,7 @@ void Pi(uint64_t A[5][5])
 }
 
 
+
 void Chi(uint64_t A[5][5])
 {
     uint64_t C[5];
@@ -146,6 +155,8 @@ void Chi(uint64_t A[5][5])
         A[y][4] = C[4];
     }
 }
+
+
 
 static const uint64_t iotas[] = {
     BIT_INTERLEAVE ? 0x0000000000000001ULL : 0x0000000000000001ULL,
@@ -175,6 +186,7 @@ static const uint64_t iotas[] = {
 };
 
 
+
 void Iota(uint64_t A[5][5], size_t i)
 {
     //assert(i < (sizeof(iotas) / sizeof(iotas[0])));
@@ -185,7 +197,6 @@ void Iota(uint64_t A[5][5], size_t i)
         A[0][0] ^= iotas[i];
     }
 }
-
 
 
 
@@ -201,90 +212,4 @@ void KeccakF1600(uint64_t A[5][5])
         Iota(A, i);
         
     }
-}
-
-
[…]

### 2022-01-07 paper/proof.tex
Add clarification on the openSSl spec

@@
 The syntactic disconnect between the quantification and looping in the published standard and the functional nature of the \cryptol\ specification is bridged with a library that hides list comprehensions with appropriately named functions.
-The final specification using the library has a more direct correspondence to the published standard that is more directly visually certified.
+The final specification using the library has a more direct correspondence to the published standard that is more obviously visually certified.
+It size is around $200$ \emph{lines of code} (LOC) in \cryptol.
+
 The specification is further validated by tests in \cryptol\ over the published test vectors.
@@
 The equivalence arrow between the \emph{NIST Standard} box and the \emph{Spec} box represents the visual certification and test vector results.
-\secref{sec:fips} details the specification and library.
+\secref{sec:fips} details the library using examples from the specification. .
 
@@
 Adding to the complexity of the equivalence proof is that the \openssl\ implementation changes the meaning of the state $S$ in the algorithm so that it no longer directly matches the meaning defined in the published standard.
-The change effectively reorders the matrix definition of the state to be more amenable to memory, and it changes the different functions in \keccak\ to be more efficient when operating on the state.
+It also changes the computation of the inner \keccak\ functions.
+These changes effectively reorder the matrix definition of the state to be more amenable to memory, and optimize the different functions in \keccak\ to be more efficient when operating on the state.
 The changes are significant enough to make it extremely difficult to argue manually that the computation implemented in \openssl\ matches that defined in the standard.
 
-The next step of the equivalence proof creates a \cryptol\ specification that matches the state reordering in the \openssl\ implementation in an effort to show an equivalence between it and the \cryptol\ specification for the standard.
-This new specification is the box in the right middle of \figref{fig:proofStructure} labeled \emph{OpenSSL} with the \emph{CRY} annotation in the upper right corner.
-\saw\ is not able to reason symbolically about input and digest sizes to construct a general proof of equivalence between the published standard and the \openssl\ \cryptol\ specifications, but it is able to prove equivalence between the two specifications given specific sizes for the input and digest.
+The next step of the equivalence proof creates a \cryptol\ specification that matches the state reordering in the \openssl\ implementation, but still follows the bit-level updates in the \fips\ standard in an effort to show an equivalence between it and the \cryptol\ specification for the standard.
+This new specification is the box in the right middle of \figref{fig:proofStructure} labeled \emph{OpenSSL} with the \emph{CRY} annotation in the upper right corner and is around 129 LOC.
+\saw\ is not able to reason symbolically about input and digest sizes to construct a general proof of equivalence between the published standard and the \openssl\ \cryptol\ specification with the state reordering, but it is able to prove equivalence between the two specifications given specific sizes for the input and digest.
 
@@
 The series of proofs varies the input size from zero bits to $1,087$ bits to reflect the $1,088$ bit rate required for a 256 bit digest.
-The proofs establish the correctness of the input padding and cover the situation where the padding adds an additional $1,088$ bit block to the input 
+The proofs establish the correctness of the input padding and cover the situation where the padding adds an additional $1,088$ bit block to the input.
+The total running time for these proofs is around 60 hours.
 As such, the meaning of the equivalence arrow between the \emph{Spec} box and the \emph{OpenSSL} box in the middle of the figure is that the two specifications have equivalent computation for a $256$ bit digest for message sizes from $0$ bits to $1,087$ bits.
@@
 That proof cannot be constructed directly by \saw\ since the extracted \sawcore\ from the C implementation is too big for automated reasoning when considering all twenty-four rounds required by \keccak.
+The disconnect is largely a result of the C implementation operating in 64-bit words at a time where the specification is bit level.
 The full proof requires overrides.
@@
 The proof of equivalence between the specification of \keccak\ and the C implementation of \keccak\ in \openssl\ relies on overrides in \saw\ as shown in \figref{fig:proofStructure2}.
-Here, \saw\ proves equivalence between the \cryptol\ specification and the C implementation for each of the five functions that comprise one round of \keccak\ given some arbitrary input state.
+Here, \saw\ proves equivalence between the bit-level \cryptol\ specifications and the word level C implementation for each of the five functions that comprise one round of \keccak\ given some arbitrary input state.
 That is the meaning of the left hand side of \figref{fig:proofStructure2}.
@@
 The equivalence is independent of the digest size, message size, or state.
+As a minor note, the \openssl\ code for the $\iota$ function was modified in one place to remove an assert-statement safe-guarding an index into a table of constants since the assert-statement is not supported by \saw's symbolic execution engine.
+The proof takes less than 15 minutes including the proofs for the overrides.
 
@@
 
-In summary, the \cryptol\ specification of the \fips\ standard relies on visual inspection and test vectors for equivalence.
-The equivalence between the \cryptol\ specification of the standard and the \cryptol\ specification of \openssl's implementation is limited to the 256 bit digest on input sizes from $0$ bits to $1,087$ bits.
+\subsection{Results and Summary}
+
+\tableref{table:stats} is a summary of the cost of the verification in running time and LOCs. These are from fairly banal hardware seen in any laptop.
+The \cryptol\ specification of the \fips\ standard relies on visual inspection and test vectors for equivalence.
+The equivalence between the \cryptol\ specification of the standard and the \cryptol\ specification of the \openssl specification that reorders the state array is limited to the 256 bit digest on input sizes from $0$ bits to $1,087$ bits and takes around 60 hours of computation time.
 That is a proof that considers all of \algoref{alg:sha3}.
-The equivalence between the \openssl\ \cryptol\ specification and the actual C implementation is limited to the \keccak\ algorithm only.
+The equivalence between the \openssl\ \cryptol\ specification that reorders the state array and the actual C implementation is limited to the \keccak\ algorithm only.
 That is a proof that only considers \lineref{line:keccak} in \algoref{alg:sha3} but holds for any state and is independent of the digest size and rate.
+The cost of proving the equivalence for each of the overrides is trivial with the exception of $\iota$.
+$\iota$ is more costly because it uses a lookup table in the C code so every entry in the table had to be proved out (see \secref{sec:openssl}).
+The final cost of proving the equivalence in \keccak with the overrides is less than 15 minutes with the yices solver.
+
+\input{stats}
 

### 2022-01-07 paper/saw.tex
Add clarification on the openSSl spec

@@
 \cryptol\ is a domain specific functional language for describing cryptographic primitives, such as algorithms to compute digests, at the bit-level.
-Cryptographic primitives are the building blocks for complex cryptographic protocols, such as SSL, that provide a broad range of security guarantees.
+Cryptographic primitives are the building blocks for complex cryptographic protocols, such as the \emph{Secure Sockets Layer} (SSL) and \emph{Transport Layer Security} (TLS) protocols, that provide a broad range of security guarantees.
 \cryptol\ is executable, being able to run test vectors to validate specifications, and more importantly, it compiles directly to \sawcore\ which is the language for formal models in \saw.
@@
 It must unroll all loops a static number of times to create the model in \sawcore.
-As such the number of iterations in any loop must be statically known at compile time or the symbolic execution fails to terminate.
+As such, the number of iterations in any loop must be statically known at compile time or the symbolic execution fails to terminate.
 \sawcore\ models from symbolic execution tend to be more complex than those from \cryptol.
@@
 It accomplishes such proofs with equivalence preserving rewrites on the \sawcore\ and with automated reasoning.
-The rewrites simplify the \sawcore\ for a models to reduce the cost of automated reasoning. 
+The rewrites simplify the \sawcore\ for a model to reduce the cost of automated reasoning. 
 The rewrites also try to structurally reduce one model to the other thus avoiding automated reasoning altogether.
@@
 An override replaces one \sawcore\ model with another in the formal analysis.
-As an example, consider a Java method that itself calls another Java method as part of its implementation.
-Suppose that \saw\ is able to prove that a \cryptol\ specification for the called method and the Java implementation of the same method are equivalent.
-If that is possible, then an override is able to replace the called Java method with the \sawcore\ from the \cryptol;
-thereby, it bypasses the symbolic execution of that called method when creating the \sawcore\ model of the Java by using the \sawcore\ from the \cryptol\ specification directly.
+As an example, consider a C function \texttt{a}() that itself calls another C function as part of its implementation: \texttt{T a()} \{\texttt{... b() ...}\}.
+The \sawcore\ from symbolic executions for \texttt{a()} of necessity inlines and executes the code for \texttt{b()}, and as a result, yields a model in \sawcore\ that is to big and complex to accomplish the equivalence proof between \texttt{a()} and the \cryptol\ specification of \texttt{a()}.
+Now suppose that \texttt{b()} is specified in \cryptol.
+If \saw\ is able to prove that the specification for \texttt{b()} and its C implementation are equivalent, then an override is able to replace the call to \texttt{b()} in the C code with the \sawcore\ from the \cryptol;
+thereby, it bypasses the inlining and symbolic execution of \texttt{b()} when creating the \sawcore\ model for \texttt{a()}.
 
@@
 thereby, the overrides reduce the overall complexity of the final extracted \sawcore\ from C or Java.
+In the example, the equivalence of \texttt{a()} and its specification using the override for \texttt{b()} now proves out quickly.
 As such, proving a property about a complex C function or Java method generally requires the use of overrides from \cryptol\ in order for the equivalence proof to be successful, and that is regardless of how that proof is accomplished.

### 2022-01-07 paper/proof.tex
Minor tweaks

@@
 \tableref{table:stats} is a summary of the cost of the verification in running time and LOCs. These are from fairly banal hardware seen in any laptop.
-The \cryptol\ specification of the \fips\ standard relies on visual inspection and test vectors for equivalence.
+The \cryptol\ specification of the \fips\ standard relies on visual inspection and test vectors for equivalence having to look at around 200 LOC.
 The equivalence between the \cryptol\ specification of the standard and the \cryptol\ specification of the \openssl specification that reorders the state array is limited to the 256 bit digest on input sizes from $0$ bits to $1,087$ bits and takes around 60 hours of computation time.
@@
 $\iota$ is more costly because it uses a lookup table in the C code so every entry in the table had to be proved out (see \secref{sec:openssl}).
-The final cost of proving the equivalence in \keccak with the overrides is less than 15 minutes with the yices solver.
+The final cost of proving the equivalence in \keccak\ with the overrides is less than 15 minutes with the \texttt{yices} solver.
 
 \input{stats}
-
-%% TODO: summarize the cost of the total proof in terms of time and identify the machine used for the proof.

### 2022-01-07 paper/openssl.tex
Fix langague on the openssl spec

@@
 
-The \cryptol\ for the \fips\ model is rewritten to reflect these memory layout and computation differences.
-The $x$ and $y$ indexing is swapped.
-All the inner \keccak\ functions are modified to operate on lanes in their entirety.
-And the the shifts are reversed.
-For reference, \figref{fig:piopenssl} is the rewritten $\pi$ function in \cryptol\ than matches \openssl's implementation (compare to \figref{fig:pi}(d)).
-Operating on lanes in faster and more efficient, and that change is reflected in all of the inner \keccak\ functions.
+The \cryptol\ for the \fips\ model is rewritten to reflect the memory layout differences but leaves in place the bit-level operations.
+The $x$ and $y$ indexing is swapped, and the shifts are reversed.
+For reference, \figref{fig:piopenssl} is the rewritten $\pi$ function in \cryptol\ than matches swapping of the indices in \openssl (compare to \figref{fig:pi}(d)).
 
@@
              for [0..4] (&x -> 
-                 a @x @((x + 3*y) % 5))
+               for [0..63] (&z -> 
+                 a @x @((x + 3*y) % 5) @z)))  
   \end{lstlisting}
@@
 The second difference is in hardcoding lookup Tables for constants rather than computing the constants along the way.
-The \fips\ \cryptol\ specification computes the lookup table once for the rounding constants, \emph{rc}, for $\iota$, and it computes the rotate constants for $\rho$ on the fly.
-The \openssl\ specification hardcodes these constants in tables without the provenance of the computation.
-\figref{fig:iota} shows the difference in the two specifications for the rounding constants with the code for the computed table coming directly from \fips.
+The \fips\ \cryptol\ specification computes each constant on-the-fly as needed in each round.
+This redundant computation is seen with the rounding constants, \emph{rc}, for $\iota$, and the rotate constants for $\rho$. 
+The \openssl\ implementation hardcodes these constants in tables without the provenance of the computation and looks up the constants in the tables.
+The new \cryptol\ specification for the \openssl uses the same lookup tables, but it differs from the \openssl implementation in that it computes the lookup tables once so that the values in the table now have provenance.
+This provenance is sufficient for the equivalence proof with the C code.
+\figref{fig:iota} shows the difference in the two specifications for the rounding constants with the code for the computed table coming directly from \fips\ in the \openssl\ specification.

### 2022-01-07 paper/openssl.tex
Fix minor typas

@@
 The \openssl\ implementation hardcodes these constants in tables without the provenance of the computation and looks up the constants in the tables.
-The new \cryptol\ specification for the \openssl uses the same lookup tables, but it differs from the \openssl implementation in that it computes the lookup tables once so that the values in the table now have provenance.
+The new \cryptol\ specification for the \openssl\ uses the same lookup tables, but it differs from the \openssl\ implementation in that it computes the lookup tables once so that the values in the table now have provenance.
 This provenance is sufficient for the equivalence proof with the C code.

### 2022-01-08 paper/conclusion.tex
Add proof figure
Update iota.tex.
Add conclusion

@@
-\section{Conclusion}\label{sec:conclusion}
-Conclude things!
+This paper presents the proof of the \shaThree\ implementation in \openssl\ using the \saw\ tool.
+The proof is accomplished by writing a \cryptol\ specification for the \nist\ \fips\ publication that defines the \shaThree\ algorithm.
+The \cryptol\ specification is aided with two additional \cryptol\ library functions, the for-method and while-method, to hide the functional list comprehensions.
+The for-method being a parallel-for operating on each element in a list.
+The while-method being iteration over a list.
+The \cryptol\ is shown equivalent to the \fips\ standard though visual inspection and input to output matching on published test vectors.
+
+The proof uses a second \cryptol\ specification based on the actual \openssl\ C implementation of \shaThree.
+This specification reorders arrays in the specification to match the C standard following row major order.
+It also includes table lookups for constant values in the computation but it computes those tables to show the provenance of the constants.
+That povenance is needed later for the proof of equivalence with the C code.
+\saw\ proves the input to output relationship equivalent between the \fips\ specification and this revised specification for the 256-bit digest on any message of size $0$ to $1088$ bits.
+
+The final \saw\ proof shows the \keccak\ function in the revised specification is equivalent to the \keccak\ function in the actual C implementation.
+\keccak\ is the core of the \shaThree\ algorithm and the most complex part.
+The proof relies of the use of overrides.
+\saw\ proves the \cryptol\ specifications of the five inner \keccak functions equivalent to their C counterparts and replaces them in the C code with the \sawcore\ from the \cryptol.
+With the overrides, \saw\ proves the equivalence for \keccak\ in \cryptol\ and C over any input.
+
+Future work is to prove out the equivalence of the input padding and the squeeze part of the \shaThree\ computation.
+These proofs are complicated by the typing in \cryptol\ and the layout of bytes in C.
+There is also some effort to show the equivalence for the other digest sizes.
+Other work looks to accomplish a proof of a Dafny implementation of \shaThree.
+Dafny synthesizes to several different backend languages making a proven \shaThree\ implementation very interesting.

### 2022-01-08 paper/openssl.tex
Fix figures

@@
 
-\begin{figure}
+\begin{figure}[t]
   \begin{center}
@@
   \end{center}
-  \caption{The \openssl\ implementation of $\pi$ that reverses indexes and operates on lanes.}
+  \caption{The \openssl\ implementation of $\pi$ that reverses indexes.}
   \label{fig:piopenssl}
@@
   \begin{lstlisting}[language=Cryptol]
+    iota a i =  for LIST4 (\y ->
+                  for LIST4 (\x ->
+                    for LIST63 (\z ->
+                      if ((x == 0) && (y == 0))
+                        then (a @0 @0 @z) ^ (LISTIOTAS @i @z)
+	                      else a @y @x @z)))
+
     LISTIOTAS = [reverse (
@@
       ).RC) | i <- LISTROUNDS:[_][64]]
-
-    LISTIOTAS = [
-      0x0000000000000001, 0x0000000000008082, 0x800000000000808a,
-      0x8000000080008000, 0x000000000000808b, 0x0000000080000001,
-      0x8000000080008081, 0x8000000000008009, 0x000000000000008a,
-      0x0000000000000088, 0x0000000080008009, 0x000000008000000a,
-      0x000000008000808b, 0x800000000000008b, 0x8000000000008089,
-      0x8000000000008003, 0x8000000000008002, 0x8000000000000080,
-      0x000000000000800a, 0x800000008000000a, 0x8000000080008081,
-      0x8000000000008080, 0x0000000080000001, 0x8000000080008008]
-        :[_][64]
   \end{lstlisting}    
@@
   \end{center}
-  \caption{Optimized $\iota$ function. (a) \openssl's implementation. (b) Matching \cryptol\ implementation.}
+  \caption{The computed \emph{rc} table with the lookup in $\iota$.}
   \label{fig:iota}
@@
 This provenance is sufficient for the equivalence proof with the C code.
-\figref{fig:iota} shows the difference in the two specifications for the rounding constants with the code for the computed table coming directly from \fips\ in the \openssl\ specification.
+\figref{fig:iota} shows the $\iota$ function that uses the computed lookup table.

### 2022-01-08 paper/related.tex
Updates related.tex, still missing formatting for references.

@@
-The verification of OpenSSL's SHA2-256 implementation served as a direct predecessor to this work (!Citation2).
+Cryptol has been used to formally verify different implementations of the same cryptographic algorithm in the past.
+As a demonstration of Cryptol's applications, Lewis et al (CRYPTOL: HIGH ASSURANCE, RETARGETABLE CRYPTO
+DEVELOPMENT AND VALIDATION) provided an portions of the Advanced Encryption Standard implemented in the language.
+The language has also been used to verify equivalence between Skein Hash algorithm implementations (Hardware/Software Co-verification of Cryptographic Algorithms using Cryptol).
+Cryptol is not limited to hardware, as Browning and Weaver (Designing Tunable, Verifiable Cryptographic Hardware Using Cryptol) have shown that the language can be tuned to rovide an abstraction of hardware.
+They later provided a proof of equivalence between the FIPS AES specificiation and various optimized Cryptol implementations.
+For certain tables, they initially demonstrated equivalence through mathematical principles.
+These tables were also proven equivalent to less time-intensive statically defined tables used for the rest of their proofs.
+This was mirrored in our own cryptol files as the OpenSSL implementation was highly optimized with similar static values.
+The language was also used to verify equivalence between the specification and hardware of the Verilog RTL's pseudorandom number generator (Formal Verification and Analysis of a Pseudo Random Number Generator).
+
+Decker et al's verification of OpenSSL's SHA2-256 implementation served as a direct predecessor to this work (Towards Verifying SHA256 in OpenSSL with the Software Analysis Workbench).
 While SHA2 is also part of NIST's Secure Hash Algorithms, its internal structure differs vastly from the more complex and resistant SHA3.
-The SHA2 proof, in turn, expanded the methodology by which Amazon's s2n HMAC was verified (!Citation2).
-While SHA3 is not implemented in s2n, the HMAC verification utilized both a high-level and low-level cryptol specification in a similar manner to this work.
-Both prior works compared equivalent C and Cryptol specification through SAW.
+The SHA2 proof, in turn, expanded the methodology by which Amazon's s2n HMAC was verified (Verifying s2n HMAC with SAW).
+While SHA3 is not implemented in s2n, the HMAC verificatiprovided a on utilized both a high-level and low-level cryptol specification in a similar manner to this work.
+Both works compared equivalent C and Cryptol specification through SAW.
+
+Other tools have been used on a wide variety of cryptographic algorithms.
+STP(A Decision Procedure for Bit-Vectors and Arrays), a SAT-based procedure, has been used to verify implementation of AES, DES, SHA-1, and other block cipher functions (Automatic Formal Verification of Block Cipher Implementations).
+Isabelle(https://isabelle.in.tum.de/) has aided in formal verification of networking protocols(Formal Verification of Secure Forwarding Protocols).
+The Coq(https://coq.inria.fr/) Proof assistant was used to prove equivalence between the OpenSSL, mbedTLS, and NIST implementations of HMAC-DRBG (Verified Correctness and Security of mbedTLS HMAC-DRBG).
+Using Jasmine(https://github.com/jasmin-lang/jasmin), Barthe et al showed that implementations of the ChaCha20 and Poly1305 were equivalent to their specifications.
+Vale has been used to verify the SHA-256 and AES implementations (Vale: Verifying High-Performance Cryptographic Assembly Code).
+More close to home, the VeriHash framework demonstrated a bug in the RHash (https://github.com/rhash/RHash) implementation of SHA3-256 (Verification of Implementations of Cryptographic Hash Functions).
+We discovered no bugs in OpenSSL's implementation.

### 2022-01-08 paper/abstract.tex
Clarify overrides explanation

--- 2022-01-07 paper/abstract.tex
@@
 The research contributes \cryptol\ specifications for the published standard and the \openssl\ implementation of the standard with its several changes.
-The equivalence proof shows the importance of overrides in \saw\ to replace C code with \cryptol\ to reduce the complexity of the proof obligations that must be accomplished relative to the implementation in C.
+The equivalence proof shows the importance of overrides in \saw\ that replace models from C code with those from \cryptol\ to reduce the complexity of the proof obligations that must be accomplished relative to the implementation in C.
 The research establishes the viability of verifying modern cryptographic primitives and the importance of modularity in their definitions, and implementations, for overrides.

### 2022-01-08 paper/saw.tex
Clarify overrides explanation

@@
 Now suppose that \texttt{b()} is specified in \cryptol.
-If \saw\ is able to prove that the specification for \texttt{b()} and its C implementation are equivalent, then an override is able to replace the call to \texttt{b()} in the C code with the \sawcore\ from the \cryptol;
-thereby, it bypasses the inlining and symbolic execution of \texttt{b()} when creating the \sawcore\ model for \texttt{a()}.
+If \saw\ is able to prove that the specification for \texttt{b()} and its C implementation are equivalent, then after symbolic execution an override is able to replace \texttt{b()} in \texttt{a()} with the \sawcore\ from the \cryptol\ for \texttt{b()}.
 

### 2022-01-08 paper/related.tex
Updates paper.bib and related.tex

--- 2022-01-06 paper/related.tex
@@
-The verification of OpenSSL's SHA2-256 implementation served as a direct predecessor to this work (!Citation2).
+Cryptol has been used to formally verify different implementations of the same cryptographic algorithm in the past.
+As a demonstration of Cryptol's applications, Lewis et al \cite{crypt-hi} provided an portions of the Advanced Encryption Standard implemented in the language.
+The language has also been used to verify equivalence between Skein Hash algorithm implementations \cite{hard-soft}.
+Cryptol is not limited to hardware, as Browning and Weaver \cite{design-verif} have shown that the language can be tuned to rovide an abstraction of hardware.
+They later provided a proof of equivalence between the FIPS AES specificiation and various optimized Cryptol implementations.
+For certain tables, they initially demonstrated equivalence through mathematical principles.
+These tables were also proven equivalent to less time-intensive statically defined tables used for the rest of their proofs.
+As the OpenSSL implementation was highly optimized with similar static values, this paper mirrored the method Browning and Weaver outlined.
+The language was also used to verify equivalence between the specification and hardware of the Verilog RTL's pseudorandom number generator \cite{pseudorandom}.
+
+Decker et al's verification of OpenSSL's SHA2-256 implementation served as a direct predecessor to this work \cite{nfm-us}.
 While SHA2 is also part of NIST's Secure Hash Algorithms, its internal structure differs vastly from the more complex and resistant SHA3.
-The SHA2 proof, in turn, expanded the methodology by which Amazon's s2n HMAC was verified (!Citation2).
-While SHA3 is not implemented in s2n, the HMAC verification utilized both a high-level and low-level cryptol specification in a similar manner to this work.
-Both prior works compared equivalent C and Cryptol specification through SAW.
+The SHA2 proof, in turn, expanded the methodology by which Amazon's s2n HMAC was verified \cite{s2n-blog}.
+While SHA3 is not implemented in s2n, the HMAC verificatiprovided a on utilized both a high-level and low-level cryptol specification in a similar manner to this work.
+Both works compared equivalent C and Cryptol specification through SAW.
+
+Other tools have been used on a wide variety of cryptographic algorithms.
+STP \cite{stp}, a SAT-based procedure, has been used to verify implementation of AES, DES, SHA-1, and other block cipher functions \cite{auto}.
+Isabelle \cite{isabelle} has aided in formal verification of networking protocols \cite{formal-protocols}.
+The Coq \cite{coq} Proof assistant was used to prove equivalence between the OpenSSL, mbedTLS, and NIST implementations of HMAC-DRBG \cite{verified-correctness}.
+Using Jasmin \cite{jasmin}, Barthe et al \cite{hi-assurance} showed that implementations of the ChaCha20 and Poly1305 were equivalent to their specifications.
+Vale has been used to verify the SHA-256 and AES implementations \cite{vale}.
+More close to home, the VeriHash framework demonstrated a bug in the RHash \cite{rhash} implementation of SHA3-256 \cite{hash-functions}.

### 2022-01-10 paper/conclusion.tex
Add revisions from Brett

@@
 It also includes table lookups for constant values in the computation but it computes those tables to show the provenance of the constants.
-That povenance is needed later for the proof of equivalence with the C code.
+That provenance is needed later for the proof of equivalence with the C code.
 \saw\ proves the input to output relationship equivalent between the \fips\ specification and this revised specification for the 256-bit digest on any message of size $0$ to $1088$ bits.
@@
 The proof relies of the use of overrides.
-\saw\ proves the \cryptol\ specifications of the five inner \keccak functions equivalent to their C counterparts and replaces them in the C code with the \sawcore\ from the \cryptol.
+\saw\ proves the \cryptol\ specifications of the five inner \keccak\ functions equivalent to their C counterparts and replaces them in the C code with the \sawcore\ from the \cryptol.
 With the overrides, \saw\ proves the equivalence for \keccak\ in \cryptol\ and C over any input.

### 2022-01-10 paper/introduction.tex
Move imp to figs and move figs to proof section
Restructure sections and file names
Add background on SAW and SHA-3
Add clarification on the openSSl spec
Add revisions from Brett

@@
 This work bridges that gap between optimized c code, and the NIST standard.
-  \input{proofFigure.tex}
+
+Give a brief history of \shaThree. Introduce briefly SAW and Cryptol.
+
+% Contributions:
+
+% (1) A Cryptol specification for SHA3 that passes test vectors and is readily comparable to the FIPS 202 specification
+% (2) A Cpyptol specification for OpenSSL's implementation of SHA3 that passes test vectors and is readily comparable to the the OpenSSL C-code—this specification uses in the Keccak core pre-computed tables and a re-ordered state-vector
+% (3) A SAW proof that the FIPS 202 specification is equivalent to the OpenSSL specification
+% (4) A SAW proof that the OpenSSL Keccak specification is equivalent to the OpenSSL C-implementation of the same function for __**all possible state vectors**__
+
+% Limitations:
+
+% * Visual inspection and limited test vectors for (1) and (2)
+% * Only provide proof certificates for message sizes of X to Y for digest sizes of … for (3)
+% * Visual inspection only for anything above Keccak for (3) in OpenSSLs C-code
+
+% @Ben: will you clarify your second point on the visual inspection for 2?
+I think what you are saying is that it is the FIPS 202 spec with the modified state vector that is what is used in OpenSSL, so there is no value in “visual inspection” in that sense.
+It which case I would restate (2) as
+
+% (2) A Cpyptol specification based on OpenSSL implementation of SHA3 that uses a re-ordered state-vector and pre-computed tables in Keccak for performance gains that passes test vectors
+
+% To a much lesser extent, I'd say another contribution is that we've shown that when a cryptographical spec is designed in a modular way (as the modern SHA3 is compared to SHA2) that facilitates formal verification---at least if using Cryptol and SAW.
+I believe we showed this by the fact that the SHA3 proofs complete successful for Yices, Z3, and ABC, whereas with our proofs for SHA2 only Yices completed.
+
 \noindent \textbf{Contributions}:
@@
 \begin{compactitem}
-  \item  A Cryptol specification for \texttt{SHA3} that directly corresponds to the \texttt{FIPS} 202 specification and passes the test vectors provided by the same specification.
-  \item A Cryptol specification based on the OpenSSL implementation of \texttt{SHA3} that uses a reordered state vector and precomputed tables in Keccak for performance gains and passes test vectors
-  \item A \texttt{SAW} proof that the \texttt{FIPS} 202 specification is equivalent to the OpenSSL specification
-  \item A \texttt{SAW} proof that the OpenSSL Keccak specification is equivalent to the OpenSSL C-implementation of the same function for all possible state vector input
-  \item Early evidence that modular cryptographic specifications such as \texttt{SHA3} lend themselves to efficient \texttt{SAW} verification against C implementations as opposed to flattened specifications such as \texttt{SHA2}
+  \item  A Cryptol specification for \shaThree\ that directly corresponds to the \nist\ \fips\ specification and passes the test vectors provided by the same specification.
+  \item A Cryptol specification based on the \openssl\ implementation of \shaThree\ that uses a reordered state vector and precomputed tables in Keccak for performance gains and passes test vectors
+  \item A \saw\ proof that the \fips\ specification is equivalent to the OpenSSL specification
+  \item A \saw\ proof that the \openssl\ \keccak specification is equivalent to the \openssl\ C-implementation of the same function for all possible state vector input
+  \item Early evidence that modular cryptographic specifications such as \shaThree lend themselves to efficient \saw\ verification against C implementations as opposed to flattened specifications such as \shaTwo.
 \end{compactitem}
 
+\begin{comment}
 \noindent \textbf{Limitations}
@@
 A simple extension of this work would be to take that already created and verified Cryptol specification, and prove it up against OpenSSL's slightly modified keccak versions.
+\end{comment}

### 2022-01-10 paper/openssl.tex
Add revisions from Brett

@@
 \figref{fig:iota} shows the $\iota$ function that uses the computed lookup table.
+
+The change to the state array structure and the added provenance giving the computation for lookup constants are the only changes to the \cryptol\ from the \fips.
+These changes are sufficient for \saw\ to prove the input to output equivalence between the two \cryptol\ specifications for a fixed set of message sizes.
+And they are sufficient to prove the equivalence to the actual C code in \openssl\ in the \keccak function for any arbitrary state.
+The result adds more evidence to the ability to prove out implementations of cryptographic primitives when given specifications.

### 2022-01-10 paper/related.tex
Add revisions from Brett

@@
-Cryptol has been used to formally verify different implementations of the same cryptographic algorithm in the past.
-As a demonstration of Cryptol's applications, Lewis et al \cite{crypt-hi} provided an portions of the Advanced Encryption Standard implemented in the language.
+\cryptol\ has been used to formally verify different implementations of the same cryptographic algorithm in the past.
+As a demonstration of \cryptol, Lewis \emph{et al} provided portions of the \emph{Advanced Encryption Standard} (AES) implemented in the language \cite{crypt-hi}.
 The language has also been used to verify equivalence between Skein Hash algorithm implementations \cite{hard-soft}.
-Cryptol is not limited to hardware, as Browning and Weaver \cite{design-verif} have shown that the language can be tuned to rovide an abstraction of hardware.
-They later provided a proof of equivalence between the FIPS AES specificiation and various optimized Cryptol implementations.
+\cryptol\ is not limited to hardware, as Browning and Weaver \cite{design-verif} have shown that the language can be tuned to provide an abstraction of hardware.
+They later provided a proof of equivalence between the FIPS AES specification and various optimized \cryptol\ implementations.
 For certain tables, they initially demonstrated equivalence through mathematical principles.
@@
 
-Decker et al's verification of OpenSSL's SHA2-256 implementation served as a direct predecessor to this work \cite{nfm-us}.
-While SHA2 is also part of NIST's Secure Hash Algorithms, its internal structure differs vastly from the more complex and resistant SHA3.
-The SHA2 proof, in turn, expanded the methodology by which Amazon's s2n HMAC was verified \cite{s2n-blog}.
-While SHA3 is not implemented in s2n, the HMAC verificatiprovided a on utilized both a high-level and low-level cryptol specification in a similar manner to this work.
-Both works compared equivalent C and Cryptol specification through SAW.
+Decker et al's verification of OpenSSL's \shaTwo-256 implementation served as a direct predecessor to this work \cite{nfm-us}.
+While \shaTwo is also part of the \nist\ Secure Hash Algorithms, its internal structure differs vastly from the more complex and resistant \shaThree.
+The \shaTwo proof, in turn, expanded the methodology by which Amazon's s2n HMAC was verified \cite{chudov,s2n-blog}.
+While \shaThree is not implemented in s2n, the HMAC verification provided utilized both a high-level and low-level \cryptol\ specification in a similar manner to this work.
+Both works compared equivalent C and \cryptol\ specification through \saw.
+Decker's work is inspired by the \openssl verification of HMAC \cite{beringer}.
 
 Other tools have been used on a wide variety of cryptographic algorithms.
-STP \cite{stp}, a SAT-based procedure, has been used to verify implementation of AES, DES, SHA-1, and other block cipher functions \cite{auto}.
+STP \cite{stp}, a SAT-based procedure, has been used to verify implementation of AES, DES, \shaOne, and other block cipher functions \cite{auto}.
 Isabelle \cite{isabelle} has aided in formal verification of networking protocols \cite{formal-protocols}.
@@
 Using Jasmin \cite{jasmin}, Barthe et al \cite{hi-assurance} showed that implementations of the ChaCha20 and Poly1305 were equivalent to their specifications.
-Vale has been used to verify the SHA-256 and AES implementations \cite{vale}.
+Vale has been used to verify the \shaTwo-256 and AES implementations \cite{vale}.
 More close to home, the VeriHash framework demonstrated a bug in the RHash \cite{rhash} implementation of SHA3-256 \cite{hash-functions}.

### 2022-01-10 paper/abstract.tex
Revise abstract slightly

@@
+This paper discusses a proof of \shaThree\ in \openssl.
 \shaThree\ is the new standard from \nist\ for computing digests. 
@@
 \cryptol\ and the Software Analysis Workbench (\saw) are a language and verification engine to specify cryptographic primitives and prove equivalence to implementations.
-As digests are the basis for all cryptographic protocols, the research in this paper is an automated \saw\ proof of equivalence between the \shaThree\ standard and the actual \openssl\ implementation in C.
-The research contributes \cryptol\ specifications for the published standard and the \openssl\ implementation of the standard with its several changes.
+As digests are the basis for all cryptographic protocols, the research in this paper is an automated \saw\ proof of equivalence between the \shaThree\ standard and its \openssl\ implementation in C.
+The research contributes \cryptol\ specifications for the standard.
 The equivalence proof shows the importance of overrides in \saw\ that replace models from C code with those from \cryptol\ to reduce the complexity of the proof obligations that must be accomplished relative to the implementation in C.
-The research establishes the viability of verifying modern cryptographic primitives and the importance of modularity in their definitions, and implementations, for overrides.
+The research establishes the viability of verifying modern cryptographic primitives and the importance of modularity in their definitions, and implementations, for overrides to manage complexity.
 

### 2022-01-10 paper/conclusion.tex
Add revision from read

@@
 That provenance is needed later for the proof of equivalence with the C code.
-\saw\ proves the input to output relationship equivalent between the \fips\ specification and this revised specification for the 256-bit digest on any message of size $0$ to $1088$ bits.
+\saw\ proves the input to output relationship equivalent between the \fips\ specification and this revised specification for the 256-bit digest on any message of size $0$ to $1,088$ bits.
 

### 2022-01-10 paper/openssl.tex
Add revision from read

@@
 The $x$ and $y$ indexing is swapped, and the shifts are reversed.
-For reference, \figref{fig:piopenssl} is the rewritten $\pi$ function in \cryptol\ than matches swapping of the indices in \openssl (compare to \figref{fig:pi}(d)).
+For reference, \figref{fig:piopenssl} is the rewritten $\pi$ function in \cryptol\ than matches the swapping of the indices in \openssl (compare to \figref{fig:pi}(d)).
 
@@
 These changes are sufficient for \saw\ to prove the input to output equivalence between the two \cryptol\ specifications for a fixed set of message sizes.
-And they are sufficient to prove the equivalence to the actual C code in \openssl\ in the \keccak function for any arbitrary state.
+And they are sufficient to prove the equivalence to the actual C code in \openssl\ in the \keccak\ function for any arbitrary state.
 The result adds more evidence to the ability to prove out implementations of cryptographic primitives when given specifications.

### 2022-01-10 paper/paper.tex
Add minor revisions
Add clarification on the openSSl spec
Add revisions from Brett
Add revision from read

@@
 \newcommand{\shaThree}{\texttt{SHA-3}}
+\newcommand{\shaTwo}{\texttt{SHA-2}}
+\newcommand{\shaOne}{\texttt{SHA-1}}
 \newcommand{\nist}{\texttt{NIST}}
@@
 \newcommand{\secref}[1]{Sec.~\ref{#1}}
+\newcommand{\tableref}[1]{Table~\ref{#1}}
 
@@
 \input{abstract}
-\keywords{verification \and cryptography \and symbolic execution \and SHA3.}
+\keywords{Formal Verification \and Cryptography \and \shaThree \and \cryptol \and \saw.}
 \end{abstract}
@@
 \cryptol\ and \saw\ are used to prove properties of \openssl's implementation of \shaThree\ relative to the \nist\ \fips\ publication \cite{fips202}. 
-What the proofs actually say, and an understanding of their limitations, relies on (1) a basic knowledge of how \saw\ works and what \cryptol\ does relative to \saw, and (2) how the \shaThree\ algorithm computes a digest.
+What the proofs actually say, and an understanding of their limitations, relies on (1) a basic knowledge of how \saw\ works and what \cryptol\ does relative to \saw\ \cite{saw}, and (2) how the \shaThree\ algorithm computes a digest \cite{fips202}.
 This section provides that basic knowledge of the tools and a general overview of the \shaThree\ algorithm.

### 2022-01-10 paper/proof.tex
Remove extra period
Add revisions from Brett
Add revision from read

@@
 The proof begins on the left of the figure with the box labeled \emph{NIST Standard}. 
-This box represents the \fips\ standard that defines the \shaThree\ algorithm with english prose and math.
+This box represents the \fips\ standard that defines the \shaThree\ algorithm with English prose and math.
 The standard also provides several test inputs with the corresponding expected digests.
@@
 The equivalence arrow between the \emph{NIST Standard} box and the \emph{Spec} box represents the visual certification and test vector results.
-\secref{sec:fips} details the library using examples from the specification. .
+\secref{sec:fips} details the library using examples from the specification.
 
@@
 These changes effectively reorder the matrix definition of the state to be more amenable to memory, and optimize the different functions in \keccak\ to be more efficient when operating on the state.
+This includes using pre-computed constants for $\iota$.
 The changes are significant enough to make it extremely difficult to argue manually that the computation implemented in \openssl\ matches that defined in the standard.
@@
 The next step of the equivalence proof creates a \cryptol\ specification that matches the state reordering in the \openssl\ implementation, but still follows the bit-level updates in the \fips\ standard in an effort to show an equivalence between it and the \cryptol\ specification for the standard.
-This new specification is the box in the right middle of \figref{fig:proofStructure} labeled \emph{OpenSSL} with the \emph{CRY} annotation in the upper right corner and is around 129 LOC.
+It also adds the computation of the constants to the specification.
+This new specification is the box in the right middle of \figref{fig:proofStructure} labeled \emph{OpenSSL} with the \emph{CRY} annotation in the upper right corner and is around 221 LOC.
 \saw\ is not able to reason symbolically about input and digest sizes to construct a general proof of equivalence between the published standard and the \openssl\ \cryptol\ specification with the state reordering, but it is able to prove equivalence between the two specifications given specific sizes for the input and digest.
@@
 A series of \saw\ proofs over a range of input sizes with the digest size fixed at $256$ bits is used to prove the equivalence between the published standard and the \openssl\ \cryptol\ specifications.
-The series of proofs varies the input size from zero bits to $1,087$ bits to reflect the $1,088$ bit rate required for a 256 bit digest.
+The series of proofs varies the input size from zero bits to $1,088$ bits, in one byte increments, to reflect the $1,088$ bit rate required for a 256 bit digest.
 The proofs establish the correctness of the input padding and cover the situation where the padding adds an additional $1,088$ bit block to the input.
 The total running time for these proofs is around 60 hours.
-As such, the meaning of the equivalence arrow between the \emph{Spec} box and the \emph{OpenSSL} box in the middle of the figure is that the two specifications have equivalent computation for a $256$ bit digest for message sizes from $0$ bits to $1,087$ bits.
+As such, the meaning of the equivalence arrow between the \emph{Spec} box and the \emph{OpenSSL} box in the middle of the figure is that the two specifications have equivalent computation for a $256$ bit digest for message sizes from $0$ bits to $1,088$ bits.
 
@@
 The proof of equivalence between the specification of \keccak\ and the C implementation of \keccak\ in \openssl\ relies on overrides in \saw\ as shown in \figref{fig:proofStructure2}.
-Here, \saw\ proves equivalence between the bit-level \cryptol\ specifications and the word level C implementation for each of the five functions that comprise one round of \keccak\ given some arbitrary input state.
+Here, \saw\ proves equivalence between the bit-level \cryptol\ specifications and the word level C implementations for each of the five functions that comprise one round of \keccak\ given some arbitrary input state.
 That is the meaning of the left hand side of \figref{fig:proofStructure2}.
@@
 The \cryptol\ specification of the \fips\ standard relies on visual inspection and test vectors for equivalence having to look at around 200 LOC.
-The equivalence between the \cryptol\ specification of the standard and the \cryptol\ specification of the \openssl specification that reorders the state array is limited to the 256 bit digest on input sizes from $0$ bits to $1,087$ bits and takes around 60 hours of computation time.
+The equivalence between the \cryptol\ specification of the standard and the \cryptol\ specification of the \openssl\ specification that reorders the state array is limited to the 256 bit digest on input sizes from $0$ bits to $1,087$ bits and takes around 60 hours of computation time.
 That is a proof that considers all of \algoref{alg:sha3}.

### 2022-01-10 paper/related.tex
Add revision from read

@@
 
-Decker et al's verification of OpenSSL's \shaTwo-256 implementation served as a direct predecessor to this work \cite{nfm-us}.
-While \shaTwo is also part of the \nist\ Secure Hash Algorithms, its internal structure differs vastly from the more complex and resistant \shaThree.
-The \shaTwo proof, in turn, expanded the methodology by which Amazon's s2n HMAC was verified \cite{chudov,s2n-blog}.
-While \shaThree is not implemented in s2n, the HMAC verification provided utilized both a high-level and low-level \cryptol\ specification in a similar manner to this work.
+Decker et al's verification of OpenSSL's \shaTwo\ 256 bit digest implementation served as a direct predecessor to this work \cite{nfm-us}.
+While \shaTwo\ is also part of the \nist\ Secure Hash Algorithms, its internal structure differs vastly from the more complex and resistant \shaThree.
+The \shaTwo\ proof, in turn, expanded the methodology by which Amazon's s2n HMAC was verified \cite{chudnov,s2n-blog}.
+While \shaThree\ is not implemented in s2n, the HMAC verification provided utilized both a high-level and low-level \cryptol\ specification in a similar manner to this work.
 Both works compared equivalent C and \cryptol\ specification through \saw.
-Decker's work is inspired by the \openssl verification of HMAC \cite{beringer}.
+Decker's work is inspired by the \openssl verification of HMAC from Beringer \cite{beringer}.
 
@@
 STP \cite{stp}, a SAT-based procedure, has been used to verify implementation of AES, DES, \shaOne, and other block cipher functions \cite{auto}.
-Isabelle \cite{isabelle} has aided in formal verification of networking protocols \cite{formal-protocols}.
-The Coq \cite{coq} Proof assistant was used to prove equivalence between the OpenSSL, mbedTLS, and NIST implementations of HMAC-DRBG \cite{verified-correctness}.
-Using Jasmin \cite{jasmin}, Barthe et al \cite{hi-assurance} showed that implementations of the ChaCha20 and Poly1305 were equivalent to their specifications.
-Vale has been used to verify the \shaTwo-256 and AES implementations \cite{vale}.
-More close to home, the VeriHash framework demonstrated a bug in the RHash \cite{rhash} implementation of SHA3-256 \cite{hash-functions}.
+Isabelle has aided in formal verification of networking protocols \cite{isabelle,formal-protocols}.
+The Coq proof assistant was used to prove equivalence between the OpenSSL, mbedTLS, and NIST implementations of HMAC-DRBG \cite{coq,verified-correctness}.
+Using Jasmin \cite{jasmin}, Barthe \emph{et al} showed that implementations of the ChaCha20 and Poly1305 were equivalent to their specifications \cite{hi-assurance}.
+Vale has been used to verify the \shaTwo\ 256 bit digest and AES implementations \cite{vale}.
+Also of interest, the VeriHash framework demonstrated a bug in the RHash implementation of \shaThree\ with the 256 bit digest \cite{rhash,hash-functions}.

### 2022-01-10 paper/introduction.tex
Add introduction

--- 2021-11-22 paper/introduction.tex
@@
-  The Secure Hashing Algorithm (SHA) family of cryptography is widely used as a means of providing increased security through generating/verifying digital signatures, deriving keys, and generating pseudorandom bits (\href{https://csrc.nist.gov/publications/detail/fips/202/final}{NIST specification}).
-  The structure and computations of these algorithms is defined by NIST in mathematical specifications.  
-  Because these specifications are just figures on a page, any assurance of correctness of an algorithm’s implementation is dependent on a visual review of the code coupled with a small sampling of test vectors.
+This paper discusses an automated proof of the \emph{Secure Hash Algorithm 3} (\shaThree) in \openssl\ using the \emph{Software Analysis Workbench} by Galois.
+\shaThree\ is the latest algorithm for computing \emph{digests}, or hashes, from data to be published by the \emph{National Institute of Standards and Technology} (\nist).
+The algorithm is the successor to \shaTwo.
+Its intent is to address weaknesses in \shaTwo, and it departs from its predecessors in that it is based on the \keccak\ family of cryptographic primitives \cite{fips202}.
 
-  The reliability of this correctness check is often weakened because of the difficulty of performing that visual review.  
-  Many implementations of the SHA algorithms go through major rewrites and optimizations which make visual inspection of the code nearly impossible. (Not sure if that’s too strong a claim).
-  This work bridges that gap between optimized c code, and the NIST standard.
-  \input{proofFigure.tex}
-\noindent \textbf{Contributions}:
-This paper details the following contributions.
+\openssl\ is a commercial grade free cryptographic library \cite{openssl}.
+Its implementation of the \emph{Secure Sockets Layer} (SSL) and \emph{Transport Layer Security} protocols are widely deployed on internet facing applications.
+Computing digests is a critical operation in these protocols and is the basis to the security guarantees provide by these protocols.
+Widely deployed implementations, such as \openssl, are especially important to be correct because of the available attack surface to hackers if a vulnerability where discovered in such implementations.
+The seriousness of getting these implementations correct is recently seen in the Apache \texttt{log4j} vulnerability affecting virtually all internet facing servers \cite{log4j}.
+
+The \emph{Software Analysis Workbench} (\saw) is a formal verification tool to prove equivalence between \sawcore\ models.
+\sawcore\ models are defined in the \cryptol\ functional language or extracted from C or Java with \emph{symbolic execution}.
+\cryptol\ itself is designed to specify cryptographic primitives at the bit level.
+\saw\ has been shown capable in proving implementation of cryptographic primitives equivalent to hardware and software implementations \cite{crypt-hi,hard-soft,design-verif}.
+Recent work emphasizes not just the utility of \saw, but the importance of modular decomposition of cryptographic primitives in their definition as that decomposition lends itself to efficient verification in \saw \cite{nfm-us}.
+
+This paper details another example of \saw\ verification only this time it is a proof showing the \openssl\ C implementation of \shaThree\ matches a \cryptol\ specification derived from the published \fips\ standard by \nist\ \cite{fips202}.
+The \cryptol\ specifications, \saw\ scripts, and extracted \sawcore\ model from the \openssl\ C implementation are documented, and available, in a public Github repository for independent verification \cite{repo}.
+The proof relies on two specifications, both created as part of this research, of the \shaThree\ algorithm: one that closely matches the published standard and another incorporating features in the C implementation.
+\saw\ shows an input to output equivalence between these two specifications for the 256 bit digest over a fixed set of message sizes. The proof takes around 60 hours of computation time.
+
+The proof then turns to the inner \keccak\ function that computes the actual bits for the digest.
+Here it shows that the C implementation exactly matches the \cryptol\ specification on any input.
+The proof requires the use of \emph{overrides} in \saw\ to simplify the extracted model from C after symbolic execution.
+The proof of equivalence between the \keccak\ in C and \cryptol\ takes around 15 minutes of computation time.
+
+The proof, with its artifacts in the published repository, make the following contributions:
 \begin{compactitem}
-  \item  A Cryptol specification for \texttt{SHA3} that directly corresponds to the \texttt{FIPS} 202 specification and passes the test vectors provided by the same specification.
-  \item A Cryptol specification based on the OpenSSL implementation of \texttt{SHA3} that uses a reordered state vector and precomputed tables in Keccak for performance gains and passes test vectors
-  \item A \texttt{SAW} proof that the \texttt{FIPS} 202 specification is equivalent to the OpenSSL specification
-  \item A \texttt{SAW} proof that the OpenSSL Keccak specification is equivalent to the OpenSSL C-implementation of the same function for all possible state vector input
-  \item Early evidence that modular cryptographic specifications such as \texttt{SHA3} lend themselves to efficient \texttt{SAW} verification against C implementations as opposed to flattened specifications such as \texttt{SHA2}
+  \item \cryptol\ for \shaThree\ that passes the test vectors and correlates with \fips.
+  \item \cryptol\ for \shaThree\ that uses a reordered memory layout and precomputed tables that correlate with C implementations.
+  \item A \saw\ proof showing the two specifications equivalent for the 256 bit digest over a series of message sizes.
+  \item A \saw\ proof that the \keccak\ specification in \cryptol\ is equivalent to the \openssl\ C implementation of the same function for all inputs.
 \end{compactitem}
+The proof shows that \saw\ is capable of proving equivalence to C implementations of modern cryptographic primitives.
+It also shows that the modular definition of \shaThree\ naturally lends itself to formal verification in \saw.
+The work in its entirety argues the importance of using formal languages such as \cryptol\ for the specification of cryptographic primitives rather than English prose and math.
+It also argues the necessity, and viability, of proving at the intermediate representation level equivalence between cryptographic specifications and actual C implementations.
 
+\begin{comment}
 \noindent \textbf{Limitations}

### 2022-01-10 paper/sha-3.tex
Add minor revisions
Change S to A for the state to match the specification

@@
     \State $p_0, \ldots, p_m = \mathtt{pad}(N, r)$\label{line:pad}
-    \State $S = 0_0, \ldots, 0_{1599}$\label{line:init}
+    \State $A = 0_0, \ldots, 0_{1599}$\label{line:init}
     \For{$p_i = p_0, \dots, p_m$}
-      \State $S = S\ \mathbf{xor}\ \mathtt{zeroExtend}(p_i, s-r)$\label{line:absorb}
-      \State $S =$ \keccak$(S)$\label{line:keccak}
+      \State $A = A\ \mathbf{xor}\ \mathtt{zeroExtend}(p_i, a-r)$\label{line:absorb}
+      \State $A =$ \keccak$(A)$\label{line:keccak}
     \EndFor
-    \State $d = \mathtt{squeeze(S,d)}$\label{line:squeeze}
+    \State $D = \mathtt{squeeze(A,d)}$\label{line:squeeze}
   \end{algorithmic}
@@
 \shaThree\ uses \emph{sponge construction} to generate hashes as shown in \algoref{alg:sha3} (see \nist\ \fips\ \cite{fips202}).
-Sponge construction \emph{absorbs} data into an internal state $S$ at a rate of $r$ bits per cycle (\lineref{line:absorb} and \lineref{line:keccak}), transforming the state on each cycle, and once all the data is absorbed, then it \emph{squeezes} out a final $d$-bit digest (\lineref{line:squeeze}).
-The transformation function, \keccak-f[1600], simply \keccak\ in the rest of this presentation, defines the size of the state $S$ to be $s = 1600$ bits, which is initialized to 0 (\lineref{line:init}).
+Sponge construction \emph{absorbs} data into an internal state $A$ at a rate of $r$ bits per cycle (\lineref{line:absorb} and \lineref{line:keccak}), transforming the state on each cycle, and once all the data is absorbed, then it \emph{squeezes} out a final $d$-bit digest (\lineref{line:squeeze}).
+The transformation function, \keccak-f[1600], simply \keccak\ in the rest of this presentation, defines the size of the state $A$ to be $a = 1600$ bits, which is initialized to 0 (\lineref{line:init}).
 
@@
 The padding is defined by the regular expression $10^*1$.
-An extra block is added for the padding if needed since every $N$ must be a multiple of $r$-bit size blocks matching the $10^*1$ regular expression at the end of the last block.
+Padding is always added to $N$ even if adding it creates an extra block.
 
 The \keccak\ transformation function takes place over twenty-four rounds with each round consisting of five steps: $\theta$, $\rho$, $\pi$, $\chi$, and $\iota$.
-The state $S$ itself is viewed as a $5 \times 5$ array of $64$-bit words (e.g.
-$5 * 5 * 64 = 1600$) in each of the steps.
-$\theta$ computes a parity over the columns in the state.
-$\rho$ is a bitwise rotate over the state.
-$\pi$ is a permutation on the state.
-$\chi$ is a non-linear operation that combines along the rows in the state.
+The state $A$ itself is viewed as a $5 \times 5$ array of $64$-bit words (e.g. $5 * 5 * 64 = 1600$) in each of the steps.
+$\theta$ computes a parity over the columns in the state. $\rho$ is a bitwise rotate over the state.
+$\pi$ is a permutation on the state. $\chi$ is a non-linear operation that combines along the rows in the state.
 And $\iota$ mixes in twenty-four different constants, one for each iteration, into portions of the state.

### 2022-01-10 paper/introduction.tex
Fix grammar errors

@@
 \cryptol\ itself is designed to specify cryptographic primitives at the bit level.
-\saw\ has been shown capable in proving implementation of cryptographic primitives equivalent to hardware and software implementations \cite{crypt-hi,hard-soft,design-verif}.
+\saw\ has been shown capable in proving implementations of cryptographic primitives equivalent to hardware and software implementations \cite{crypt-hi,hard-soft,design-verif}.
 Recent work emphasizes not just the utility of \saw, but the importance of modular decomposition of cryptographic primitives in their definition as that decomposition lends itself to efficient verification in \saw \cite{nfm-us}.

### 2022-01-10 paper/introduction.tex
Fix GitHub typo

--- 2022-01-10 paper/introduction.tex
@@
 This paper details another example of \saw\ verification only this time it is a proof showing the \openssl\ C implementation of \shaThree\ matches a \cryptol\ specification derived from the published \fips\ standard by \nist\ \cite{fips202}.
-The \cryptol\ specifications, \saw\ scripts, and extracted \sawcore\ model from the \openssl\ C implementation are documented, and available, in a public Github repository for independent verification \cite{repo}.
+The \cryptol\ specifications, \saw\ scripts, and extracted \sawcore\ model from the \openssl\ C implementation are documented, and available, in a public GitHub repository for independent verification \cite{repo}.
 The proof relies on two specifications, both created as part of this research, of the \shaThree\ algorithm: one that closely matches the published standard and another incorporating features in the C implementation.

### 2022-01-10 paper/introduction.tex
Fix the repository reference

@@
 Its implementation of the \emph{Secure Sockets Layer} (SSL) and \emph{Transport Layer Security} protocols are widely deployed on internet facing applications.
-Computing digests is a critical operation in these protocols and is the basis to the security guarantees provide by these protocols.
-Widely deployed implementations, such as \openssl, are especially important to be correct because of the available attack surface to hackers if a vulnerability where discovered in such implementations.
+Computing digests is a critical operation in these protocols and is the basis to the security guarantees provided by these protocols.
+Widely deployed implementations, such as \openssl, are especially important to be correct because of the available attack surface to hackers if a vulnerability were discovered in such implementations.
 The seriousness of getting these implementations correct is recently seen in the Apache \texttt{log4j} vulnerability affecting virtually all internet facing servers \cite{log4j}.
@@
 This paper details another example of \saw\ verification only this time it is a proof showing the \openssl\ C implementation of \shaThree\ matches a \cryptol\ specification derived from the published \fips\ standard by \nist\ \cite{fips202}.
-The \cryptol\ specifications, \saw\ scripts, and extracted \sawcore\ model from the \openssl\ C implementation are documented, and available, in a public GitHub repository for independent verification \cite{repo}.
+The \cryptol\ specifications, \saw\ scripts, and extracted \sawcore\ model from the \openssl\ C implementation are documented, and available, in a public GitHub repository for independent verification \cite{sha3proof}.
 The proof relies on two specifications, both created as part of this research, of the \shaThree\ algorithm: one that closely matches the published standard and another incorporating features in the C implementation.

### 2022-03-24 paper/introduction.tex
fixed minor typos

@@
 \saw\ has been shown capable in proving implementations of cryptographic primitives equivalent to hardware and software implementations \cite{crypt-hi,hard-soft,design-verif}.
-Recent work emphasizes not just the utility of \saw, but the importance of modular decomposition of cryptographic primitives in their definition as that decomposition lends itself to efficient verification in \saw \cite{nfm-us}.
+Recent work emphasizes not just the utility of \saw, but the importance of modular decomposition of cryptographic primitives in their definition as that decomposition lends itself to efficient verification in \saw\ \cite{nfm-us}.
 

### 2022-03-24 paper/related.tex
Fix grammar errors
fixed minor typos

@@
 The language has also been used to verify equivalence between Skein Hash algorithm implementations \cite{hard-soft}.
-\cryptol\ is not limited to hardware, as Browning and Weaver \cite{design-verif} have shown that the language can be tuned to provide an abstraction of hardware.
+\cryptol\ is not limited to software, as Browning and Weaver \cite{design-verif} have shown that the language can be tuned to provide an abstraction of hardware.
 They later provided a proof of equivalence between the FIPS AES specification and various optimized \cryptol\ implementations.
@@
 Both works compared equivalent C and \cryptol\ specification through \saw.
-Decker's work is inspired by the \openssl verification of HMAC from Beringer \cite{beringer}.
+Decker's work is inspired by the \openssl\ verification of HMAC from Beringer \cite{beringer}.
 

### 2022-03-24 paper/saw.tex
fixed minor typos

@@
 Cryptographic primitives are the building blocks for complex cryptographic protocols, such as the \emph{Secure Sockets Layer} (SSL) and \emph{Transport Layer Security} (TLS) protocols, that provide a broad range of security guarantees.
-\cryptol\ is executable, being able to run test vectors to validate specifications, and more importantly, it compiles directly to \sawcore\ which is the language for formal models in \saw.
+\cryptol\ is executable, being able to run test vectors to validate specifications, and more importantly, it compiles directly to \sawcore, which is the language for formal models in \saw.
 

### 2022-03-24 paper/saw.tex
fix last review typos

@@
 As an example, consider a C function \texttt{a}() that itself calls another C function as part of its implementation: \texttt{T a()} \{\texttt{... b() ...}\}.
-The \sawcore\ from symbolic executions for \texttt{a()} of necessity inlines and executes the code for \texttt{b()}, and as a result, yields a model in \sawcore\ that is to big and complex to accomplish the equivalence proof between \texttt{a()} and the \cryptol\ specification of \texttt{a()}.
+The \sawcore\ generated from symbolic executions for \texttt{a()} of necessity inlines and executes the code for \texttt{b()}, and as a result, yields a model in \sawcore\ that is to big and complex to accomplish the equivalence proof between \texttt{a()} and the \cryptol\ specification of \texttt{a()}.
 Now suppose that \texttt{b()} is specified in \cryptol.

### 2022-03-25 paper/related.tex
Add related work

@@
+Almeida et al give a formal specification of \shaThree\ in Jasmin \cite{10.1145/3319535.3363211}.
+They then prove that \shaThree\ is resistant to collisions, correctly implemented by Jasmin in vectorized x86, and that implementation is resistant to side-channel attacks.
+The proofs are mechanized and carried out by the EasyCrypt proof assistant \cite{easycrypt}.
+The vectorized implementation is performant and efficient.
+
+EverCrypt from Protzenko et al is a library of verified cryptographic primitives including several state of art algorithms for computing digests \cite{9152808}.
+Here the algorithms are specified in F* \cite{fstar}.
+Automated proofs of properties of the crypto-primitives are done with a weakest-precondition calculus.
+As subset of F* is compilable to C the library is compiled to different target backends.
+The resulting library is performant and efficient.
+
+The work is this paper differs from the these related efforts in that this work does not prove properties of \shaThree\ itself but rather proves an existing implementation of \shaThree\ is correct relative to the published standard.
+Here the verification effort is to prove \openssl\ correct in its implementation of \shaThree.
+\openssl\ being a widely deployed implementation written for performance not thinking of verification makes it an interesting target for formal verification and an interesting case study for \saw.
+
 \cryptol\ has been used to formally verify different implementations of the same cryptographic algorithm in the past.

### 2022-05-12 paper/conclusion.tex
Add minor grammar and sentence structure changes

@@
 The \cryptol\ specification is aided with two additional \cryptol\ library functions, the for-method and while-method, to hide the functional list comprehensions.
-The for-method being a parallel-for operating on each element in a list.
-The while-method being iteration over a list.
-The \cryptol\ is shown equivalent to the \fips\ standard though visual inspection and input to output matching on published test vectors.
+The for-method is a parallel-for operating on each element in a list, and the while-method is a list iteration.
+The \cryptol\ is shown equivalent to the \fips\ standard through visual inspection and input to output matching on published test vectors.
 

### 2022-05-12 paper/saw.tex
Add minor grammar and sentence structure changes

--- 2022-03-24 paper/saw.tex
@@
 As an example, consider a C function \texttt{a}() that itself calls another C function as part of its implementation: \texttt{T a()} \{\texttt{... b() ...}\}.
-The \sawcore\ from symbolic executions for \texttt{a()} of necessity inlines and executes the code for \texttt{b()}, and as a result, yields a model in \sawcore\ that is to big and complex to accomplish the equivalence proof between \texttt{a()} and the \cryptol\ specification of \texttt{a()}.
+The \sawcore\ generated from symbolic executions for \texttt{a()} of necessity inlines and executes the code for \texttt{b()}, and as a result, yields a model in \sawcore\ that is too big and complex to accomplish the equivalence proof between \texttt{a()} and the \cryptol\ specification of \texttt{a()}.
 Now suppose that \texttt{b()} is specified in \cryptol.

### 2022-05-12 paper/fips202.tex
Add revisions for oppenssl text
Fix figures
Abstract while body, scale, and caption for figure 5.

@@
                    [if i <= t % 255 then;\label{line:compute};
-                      (take'{8} ([(r@0) ^ (r@8),
-                                   r@1, r@2, r@3, 
-                                  (r@4) ^ (r@8),
-                                  (r@5) ^ (r@8),
-                                  (r@6) ^ (r@8),
-                                   r@7,r@8])
-                        where r = [0] # (rs @(i - 1)))
+                      ((nextR r) where r = [0] # (rs @(i - 1)))
                     else rs @(i - 1);\label{line:stutter};
@@
                           i = state.i + 1,;\label{line:inc};
-                          R = (take'{8} ([(r@0) ^ (r@8),
-                                           r@1, r@2, r@3,
-                                          (r@4) ^ (r@8),
-                                          (r@5) ^ (r@8),
-                                          (r@6) ^ (r@8),
-                                           r@7, r@8]) 
-                                where r = [0] # state.R)})
+                          R = (nextR r) where r = [0] # state.R})
                  ).R @0;\label{line:libreturn};
@@
 Another common idiom in the \fips\ standard is to iteratively transform some state a fixed number of times to arrive at a final state.
-The definition of the \emph{rc} function in \figref{fig:rc} is one such example.
+The definition of the \emph{round constant}, \emph{rc}, function in \figref{fig:rc} is one such example.
 The \emph{rc} function is a support function for the $\iota$ function in \keccak\ and is used to build a mask.
@@
 
+\begin{figure}[t]
+  \begin{center}
+    \begin{tabular}{l}
+      \usebox{\RcCry} \\ \\
+      \multicolumn{1}{c}{(a)} \\ \\
+      \usebox{\whilemethod} \\ \\
+      \multicolumn{1}{c}{(b)} \\ \\ 
+      \usebox{\RcCryLib} \\ \\
+      \multicolumn{1}{c}{(c)}
+    \end{tabular}
+  \end{center}
+  \caption{The \emph{rc} function in \cryptol. (a) The \cryptol\ with list comprehensions. (b) The recursive while-method definition. (c) The \cryptol\ with the while-method.}
+  \label{fig:rccry}
+\end{figure}
+
 \lineref{line:init} is the first entry in the list and it is the initial value of $R$.
@@
 
-\begin{figure}[h]
-  \begin{center}
-    \begin{tabular}{l}
-      \usebox{\RcCry} \\ \\
-      \multicolumn{1}{c}{(a)} \\ \\
-      \usebox{\whilemethod} \\ \\
-      \multicolumn{1}{c}{(b)} \\ \\ 
-      \usebox{\RcCryLib} \\ \\
-      \multicolumn{1}{c}{(c)}
-    \end{tabular}
-  \end{center}
-  \caption{The \emph{rc} function in \cryptol. (b) The \cryptol\ with list comprehensions. (b) The recursive while-method definition. (c) The \cryptol\ with the while-method.}
-  \label{fig:rccry}
-\end{figure}
-
 \figref{fig:rccry}(c) is the definition of \emph{rc} using the while-method in \figref{fig:rccry}(b).

### 2022-05-13 paper/introduction.tex
Fix minor typos and make word changes

@@
-This paper discusses an automated proof of the \emph{Secure Hash Algorithm 3} (\shaThree) in \openssl\ using the \emph{Software Analysis Workbench} by Galois.
+This paper discusses an automated proof of the \emph{Secure Hash Algorithm 3} (\shaThree) in \openssl\ using the \emph{Software Analysis Workbench} (\saw) by Galois.
 \shaThree\ is the latest algorithm for computing \emph{digests}, or hashes, from data to be published by the \emph{National Institute of Standards and Technology} (\nist).
@@
 
-The \emph{Software Analysis Workbench} (\saw) is a formal verification tool to prove equivalence between \sawcore\ models.
+\saw\ is a formal verification tool to prove equivalence between \sawcore\ models.
 \sawcore\ models are defined in the \cryptol\ functional language or extracted from C or Java with \emph{symbolic execution}.

### 2022-05-13 paper/proof.tex
fixed minor typos
Fix minor typos and make word changes

@@
 
-The goal of the work presented in this paper is to prove the \openssl\ implementation of \shaThree\ matches the \fips\ publication by \nist.
+The goal of the work presented in this paper is to prove the \openssl\ implementation of \shaThree\ matches the \fips\ specification by \nist.
 \figref{fig:proofStructure} illustrates the proof strategy.
 The proof begins on the left of the figure with the box labeled \emph{NIST Standard}. 
-This box represents the \fips\ standard that defines the \shaThree\ algorithm with English prose and math.
-The standard also provides several test inputs with the corresponding expected digests.
+This box represents the \fips\ specification that defines the \shaThree\ algorithm with English prose and math.
+The specification also provides several test inputs with the corresponding expected digests.
 The first step of the proof reproduces the \fips\ \shaThree\ description in \cryptol.
@@
 The final specification using the library has a more direct correspondence to the published standard that is more obviously visually certified.
-It size is around $200$ \emph{lines of code} (LOC) in \cryptol.
+Its size is around $200$ \emph{lines of code} (LOC) in \cryptol.
 
@@
 It also changes the computation of the inner \keccak\ functions.
-These changes effectively reorder the matrix definition of the state to be more amenable to memory, and optimize the different functions in \keccak\ to be more efficient when operating on the state.
+These changes effectively reorder the matrix definition of the state to be more amenable to memory and optimize the different functions in \keccak\ to be more efficient when operating on the state.
 This includes using pre-computed constants for $\iota$.
@@
 The \cryptol\ specification of the \fips\ standard relies on visual inspection and test vectors for equivalence having to look at around 200 LOC.
-The equivalence between the \cryptol\ specification of the standard and the \cryptol\ specification of the \openssl\ specification that reorders the state array is limited to the 256 bit digest on input sizes from $0$ bits to $1,087$ bits and takes around 60 hours of computation time.
+The equivalence between the \cryptol\ specification of the standard and the \cryptol\ specification of the \openssl\ specification that reorders the state array is limited to the 256 bit digest on input sizes from $0$ bits to $1,088$ bits and takes around 60 hours of computation time.
 That is a proof that considers all of \algoref{alg:sha3}.

### 2022-05-13 paper/sha-3.tex
Fix minor typos and make word changes

@@
     \State $p_0, \ldots, p_m = \mathtt{pad}(N, r)$\label{line:pad}
-    \State $A = 0_0, \ldots, 0_{1599}$\label{line:init}
+    \State $A = 0_0, \ldots, 0_{1599}$\label{line:initkeccak}
     \For{$p_i = p_0, \dots, p_m$}
@@
 Sponge construction \emph{absorbs} data into an internal state $A$ at a rate of $r$ bits per cycle (\lineref{line:absorb} and \lineref{line:keccak}), transforming the state on each cycle, and once all the data is absorbed, then it \emph{squeezes} out a final $d$-bit digest (\lineref{line:squeeze}).
-The transformation function, \keccak-f[1600], simply \keccak\ in the rest of this presentation, defines the size of the state $A$ to be $a = 1600$ bits, which is initialized to 0 (\lineref{line:init}).
+The transformation function, \keccak-f[1600], simply \keccak\ in the rest of this presentation, defines the size of the state $A$ to be $a = 1600$ bits, which is initialized to 0 (\lineref{line:initkeccak}).
 

### 2022-05-13 paper/paper.tex
Style and caption fig 5

@@
   morekeywords=[3]{pi, rc, cond, a, b, x, y, z, for, n, r, R, f, rs, i, index, t, 
-    vals, loop, state, while, take}
+    vals, loop, state, while, take, nextR}
 }

### 2022-05-13 paper/related.tex
Add minor grammar and sentence structure changes
Fix et al in rw

@@
 
-EverCrypt from Protzenko et al is a library of verified cryptographic primitives including several state of art algorithms for computing digests \cite{9152808}.
+EverCrypt from Protzenko \emph{et al.} is a library of verified cryptographic primitives including several state of art algorithms for computing digests \cite{9152808}.
 Here the algorithms are specified in F* \cite{fstar}.
@@
 
-\cryptol\ has been used to formally verify different implementations of the same cryptographic algorithm in the past.
+\cryptol\ has been used to formally verify several cryptographic algorithms in the past.
 As a demonstration of \cryptol, Lewis \emph{et al} provided portions of the \emph{Advanced Encryption Standard} (AES) implemented in the language \cite{crypt-hi}.

### 2022-06-14 paper/paper.tex
Improved cryptol color consitency

--- 2022-01-10 paper/paper.tex
@@
   {.}{{{\color{black}.}}}{1}  
-  {\'}{{{\color{black}'}}}{1}
+  {\'}{{{\color{black}‘}}}{1}
   {:}{{{\color{black}:}}}{1}
@@
   morekeywords=[2]{if, then, else, where},
-  morekeywords=[3]{pi, rc, cond, a, b, x, y, z, for, n, r, R, f, rs, i, index, t, 
-    vals, loop, state, while, take}
+  morekeywords=[3]{pi, iota, LISTIOTAS, LISTROUNDS, reverse, L, rc, RC, cond, a, b, x, y, z, for, n, r, R, f, rs, i, j, index, t, 
+    vals, loop, state, while, take, nextR}
 }

### 2022-06-14 paper/paper.tex
Add Parker's orcid

@@
 \author{
-  Parker Hanson\inst{1} \and
+  Parker Hanson\inst{1}\orcidID{0000-0003-0893-0096} \and
   Benjamin Winters\inst{1} \and
